### 2019-02-13 paper/main.tex
Init tex paper

\documentclass[a4paper]{book}

% TODO: Add bib functionality

% Typography
\usepackage{palatino}
\usepackage{fontenc}

% Links
\usepackage{hyperref}

\begin{document}

\begin{titlepage}
    \begin{center}
        \huge\textbf{Thesis Title}
        \normalsize

        \vspace{1cm}

        \textbf{Tudor Paisa}\\
        SNR: 2019551\\
        ANR: 315146\\

        \vspace{1cm}
        \textsc{Thesis submitted in partial fulfillment\\
        of the requirements for the degree of\\
        Master of Science in Data Science and Society,\\
        at the School of Humanities and Digital Sciences\\
        of Tilburg University\\}

        \vspace{2cm}
        Thesis Committee:\\
        Dr. Menno van Zaanen\\
        Dr. SECOND READER\\
        \vfill

        Tilburg University\\
        School of Humanities and Digital Sciences\\
        Department of Cognitive Science \& Artificial Intelligence\\
        Tilburg, The Netherlands\\
        \today

    \end{center}
\end{titlepage}

\chapter*{Abstract}
    Lorem Ipsum

\tableofcontents

\chapter{Introduction}

\chapter{Related Work}

\chapter{Methods}

\chapter{Results}

\chapter{Discussion}

\chapter{Conclusion}

\end{document}

### 2019-03-12 paper/images/colorization_colorization_latex/fig.tex
Added proposal, early baseline, and early reptile

\begin{figure}[h]
\begin{center}
\bgroup 
 \def\arraystretch{0.2} 
 \setlength\tabcolsep{0.2pt}
\begin{tabular}{cccc}
 & Classification & Ours & \\
 L2 \cite{zhang2016colorful} & (rebal.) \cite{zhang2016colorful} & (L1 + cGAN) & Ground truth \\
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L2_zhang2016_33094.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/classrebal_zhang2016_33094.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L1cGAN_33094.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/gt_33094.png} \\ 
% \includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L2_zhang2016_42366.png} &
% \includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/classrebal_zhang2016_42366.png} &
% \includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L1cGAN_42366.png} &
% \includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/gt_42366.png} \\ 
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L2_zhang2016_43989.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/classrebal_zhang2016_43989.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L1cGAN_43989.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/gt_43989.png} \\ 
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L2_zhang2016_44956.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/classrebal_zhang2016_44956.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L1cGAN_44956.png} &
\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/gt_44956.png} %\\ 
%\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L2_zhang2016_45617.png} &
%\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/classrebal_zhang2016_45617.png} &
%\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/L1cGAN_45617.png} &
%\includegraphics[width=0.25\linewidth]{figs/colorization_colorization_latex/gt_45617.png}

\end{tabular} \egroup 
\end{center}
\vspace{-0.1in}
\caption{Colorization results of conditional GANs versus the L2 regression from \cite{zhang2016colorful} and the full method (classification with rebalancing) from \cite{zhou2016learning}. The cGANs can produce compelling colorizations (first two rows), but have a common failure mode of producing a grayscale or desaturated result (last row).}
\vspace{-0.1in}
\label{colorization_res}
\end{figure}

### 2019-03-12 paper/images/handbags_edges_lotsofresults_latex/fig.tex
Added proposal, early baseline, and early reptile

\begin{figure*}
\begin{center}
\bgroup 
 \def\arraystretch{0.2} 
 \setlength\tabcolsep{0.2pt}
\begin{tabular}{cccccc}
Input & Ground truth & Output & Input & Ground truth & Output \\ 
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_128_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_128_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_128_AB.jpg} \hspace{0.025in} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_146_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_146_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_146_AB.jpg} \\ 
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_12_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_12_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_12_AB.jpg} \hspace{0.025in} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_157_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_157_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_157_AB.jpg} \\ 
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_130_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_130_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_130_AB.jpg} \hspace{0.025in} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_119_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_119_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_119_AB.jpg} \\ 
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_62_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_62_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_62_AB.jpg} \hspace{0.025in} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_106_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_106_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_106_AB.jpg}

\end{tabular} \egroup 
\end{center}
\caption{Example results of our method on automatically detected edges$\rightarrow$handbags, compared to ground truth.}
\label{handbags_edges_lotsofresults}
\end{figure*}

### 2019-03-12 paper/main.tex
Added proposal, early baseline, and early reptile

@@
 
-% TODO: Add bib functionality
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
+\usepackage{titlesec, blindtext, color}
+\definecolor{gray75}{gray}{0.75}
+\newcommand{\hsp}{\hspace{10pt}}
+\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{gray75}{\thechapter \hsp \vline}\hsp}{0pt}{\Huge\bfseries}
+
+\usepackage[american]{babel}
+\usepackage{csquotes}
+\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
+\DeclareLanguageMapping{american}{american-apa}
+\addbibresource{references.bib}
+
+% Essentials
+\usepackage{tabularx}
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage[ruled]{algorithm2e}
 
@@
 \usepackage{hyperref}
+
+% Math stuff
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amstext}
+\usepackage{bm}
+
+% Commands
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 
@@
         \textbf{Tudor Paisa}\\
-        SNR: 2019551\\
-        ANR: 315146\\
+        SNR: 2019551 | ANR: 315146\\
+        t.paisa@tilburguniversity.edu\\
 
@@
 
+The reemergence of neural networks (NN) has brought new ways to tackle Machine Learning problems (such as classification; see \cite{lecun_deep_2015} for several examples), as well as new possibilities to further artificial intelligence research (such as chatbots, or general-purpose AI systems; e.g., \cite{dale_return_2016} and \cite{vinyals_starcraft_2017} respectively).
+Currently, two exciting areas of research in the field of Deep Learning (DL) are in designing generative models \parencite{goodfellow_nips_2016} and training NNs on small datasets \parencite{finn_model-agnostic_2017, nichol_first-order_2018}.
+This research will combine both of these tasks by exploring the extent to which generative models can create novel and qualitative samples of music under the constraints posed by the few-shot learning problem.
+In other words, when data is scarce, can a generative model create samples of music that are new (i.e., not in the training set), consistent with common scales, and diverse (tone sequences are not repeated over and over, combines chord structures with single-note sequences, is not restricted to - say - one octave).
+This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under the Reptile learning framework \parencite{nichol_first-order_2018} on small amounts of MIDI data.
+
+DL as a field, promises to discover rich hierarchical models \parencite{lecun_deep_2015}, over various artificial intelligence applications, such as speech recognition, image classification, or text generation.
+The majority of developments usually involved discriminative models \parencite{goodfellow_generative_2014}, where high-dimensional input is assigned to a class label \parencite{zhang_character-level_2015, krizhevsky_imagenet_2012}.
+In contrast, successes in generative models have been scarce, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable for the longest run, followed years later by Bidirectional Long Short Term Memory \parencite{graves_generating_2013} and Variational Autoencoders \parencite{kingma_auto-encoding_2013}.
+
+However, studies on generative architectures were fairly uncommon until the discovery of the Generative Adversarial Network (GAN) by \textcite{goodfellow_generative_2014}.
+GANs have had significant successes in image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, image-to-image translation such as a satellite image into a map, or a sketch into a photorealistic image \parencite[;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}{.24\linewidth}
+        \includegraphics[width=\linewidth]{images/comic_HR.jpg}
+        \caption{original}
+        \label{fig:srgan_orig}
+    \end{subfigure}
+    \begin{subfigure}{.24\linewidth}
+        \includegraphics[width=\linewidth]{images/comic_SRF_4_bicubic.jpg}
+        \caption{bicubic}
+        \label{fig:srgan_bicubic}
+    \end{subfigure}
+    \begin{subfigure}{.24\linewidth}
+        \includegraphics[width=\linewidth]{images/comic_SRResNet-MSE.jpg}
+        \caption{SRResNet}
+        \label{fig:srgan_srresnet}
+    \end{subfigure}
+    \begin{subfigure}{.24\linewidth}
+        \includegraphics[width=\linewidth]{images/comic_SRGAN-VGG54.jpg}
+        \caption{SRGAN}
+        \label{fig:srgan_srgan}
+    \end{subfigure}
+
+    \caption{Example of single-image super-resolution results that highlight the advantages of a GAN.
+The original high-resolution image (\ref{fig:srgan_orig}) has been downsampled to make a low-resolution image.
+The results of bicubic interpolation (\ref{fig:srgan_bicubic}) can be seen in the second image, followed by the results of SRResNet (\ref{fig:srgan_srresnet}), a neural network trained on mean squared error.
+Lastly, the results of SRGAN can be seen in the final image (\ref{fig:srgan_srgan}).
+Figure adapted from \textcite{ledig_photo-realistic_2016}.}
+    \label{fig:srgan}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \begin{tabular}{ccc}
+    Input & Ground Truth & Output \\
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_106_AB.jpg} &
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_106_AB.jpg} &
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_106_AB.jpg} \\
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_12_AB.jpg} &
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_12_AB.jpg} &
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_12_AB.jpg} \\
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_130_AB.jpg} &
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_130_AB.jpg} &
+    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_130_AB.jpg} \\
+    \end{tabular}
+    \caption{\textcite{isola_image--image_2016} coined the term image-to-image translation where an input image is transformed into a revisualization of the original input.
+This figures illustrates how an outline of a purse (leftmost column) is transformed into a colored photorealistic version of it (rightmost column).
+In the middle we have the ground truth: the real image of the purse.
+Figure adapted from \textcite{isola_image--image_2016}.}
+    \label{fig:pix2pix}
+\end{figure}
+
+These developments in generative models lie in the field of Computer Vision, but that is not to say that we can generate only images.
+GANs (and others) have been applied - with varying degrees of success - in Natural Language Processing where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}.
+Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}.
+This paper will focus on the problem of generating music with a NN-based model.
+Respectively, this paper will evaluate the samples generated by C-RNN-GAN \parencite{mogren_c-rnn-gan_2016} and Performance-RNN \parencite{oore_this_2018} when trained under the few-shot learning problem (detailed below).
+
+It is fairly well known that DL models tend to be very "data-hungry", a caveat imposed by its optimization algorithm \parencite{chen_closer_2018, ravi_optimization_2016}.
+DL usually relies on a procedure called stochastic gradient descent (SGC) which implies computing for a given input vector the output and error, followed by calculating the input's average gradient and consequently adjusting the network's weights \parencite{lecun_deep_2015}.
+\textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates: when faced with non-convex optimization problems, DL algorithms do not guarantee convergence within a reasonable amount of time.
+Finally, they add by saying that for each new problem (e.g., new dataset), the model needs to reinitialize with a new set of random parameters which hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}.
+
+Besides the promise of making DL algorithms easier to train, primary reasons why this issue needs to be addressed are as follows.
+Firstly, humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}, something which DL algorithms do not.
+Secondly, this would alleviate the constant need for data collection in order to provide reasonable performances.
+Finally, there are many fields where the data exhibits a large number of classes, with few examples per class and Bridging the gap between human-type learning and current learning architectures would allow models to properly capture this sort of data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}.
+This is known as the few-shot problem \parencite{chen_closer_2018, vinyals_matching_2016, larochelle_few-shot_2017}.
+
+In spite of its proposed benefits, developments and experiments in few-shot learning are still scarce \parencite{larochelle_few-shot_2017} and evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}.
+Few-shot experiments with generative networks are even less frequent \parencite{zhang_metagan_2018, clouatre_figr_2019}.
+This research aims to explore the viability of this learning approach in the context of generating music data.
+More concretely, in light of recent developments, this research will evaluate two generative networks modelled on MIDI data under the constraints of few-shot learning, and comparing its performance to the baseline proposed by \textcite{larochelle_few-shot_2017} and properties extracted from real music performances.
+
 \chapter{Related Work}
 
+\textcite{lake_human-level_2015} initially popularized the idea of learning from a few class examples using a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
+\textcite{rezende_one-shot_2016} were among the first to provide a solution to the one-shot problem in the context of a neural network (NN) by implementing Bayesian reasoning into a deep NN embedded within hierarchical latent variable models.
+The Reptile algorithm \parencite{nichol_first-order_2018} is a recent approach to this problem.
+It counteracts the shortcomings of gradient-optimization algorithms by learning the model's initial parameters in order to maximize its performance on novel tasks.
+Furthermore, it allows for any type of network to be used and does not place any restrictions on the type of loss function that can be used \parencite{nichol_first-order_2018}.
+The algorithm for Reptile is as follows:
+
+\IncMargin{1em}
+\begin{algorithm}
+Initialize $\phi$, the vector of initial parameters
+
+\For{iteration = $1, 2, \dots$}{
+    Sample task $\tau$, corresponding to loss $L_{\tau}$ on weight vectors $\tilde \phi$
+
+    Compute $\tilde \phi = U_{\tau}^{k} (\phi)$, denoting $k$ steps of SGD or Adam
+
+    Update $\phi \leftarrow \phi + \epsilon (\tilde{\phi} - \phi)$
+}
+
+\caption{Reptile (serial version)}
+\label{alg:reptile}
+\end{algorithm}
+\DecMargin{1em}
+
+Although simplistic, this algorithm reaches similar results to another popular few-shot algorithm called MAML \parencite{finn_model-agnostic_2017}.
+Both algorithms are initialization-based learning methods however, a major distinction between the two is that the vanilla MAML implementation requires calculating a second-order derivative, which can significantly slowdown computation \parencite{nichol_first-order_2018}.
+This forms a compelling reason to use Reptile as the few-shot learning method for this research.
+
+Turning over to the generative side of things, nowadays, the most popular form of generating data is through the use of a Generative Adversarial Network \parencite[GAN;][]{goodfellow_generative_2014}.
+Here a generative model $G$ captures the data distribution and creates samples from it, whereas a discriminative model $D$ estimates the probability that the input is real (i.e., comes from the training data) rather than fake (i.e., comes from $G$).
+Formally, given $x$ sampled from real data, and $z$ being input noise, $D$ tries to keep $D(x)$ near $1$, while making $D(G(z))$ near $0$.
+Conversely, $G$ tries to make $D(G(z))$ near $1$ \parencite{goodfellow_nips_2016}.
+A notable implementation is C-RNN-GAN \parencite{mogren_c-rnn-gan_2016}, a continuous recurrent neural network with adversarial training.
+It consists of a network with two LSTM layers and 350 hidden units per layer, in both $G$ and $D$.
+However, $D$ has a bidirectional layout which allows is to take context from both past and future for its decisions, while $G$ is unidirectional \parencite{mogren_c-rnn-gan_2016}.
+As is the standard practice for GANs, $G$ will create fake samples of music, whereas $D$ will try to classify correctly which sample is real.
+
+Notwithstanding GANs, other generative successes have made use of adaptation on the Long Short-Term Memory (LSTM) network \parencite{hochreiter_long_1997}.
+Here, the cells of a network are designed to allow for information to persist by implementing a recursive state unit, an input gate, a forget gate, and an output gate.
+The input of the network is allowed into the state unit if it is allowed to pass through by the input gate.
+Next, the forget gate controls the self-loop of the state cell.
+Finally, the output gate has the power to shut off the output cell \parencite{goodfellow_deep_2016}.
+Here, a notable implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 layers each.
+Compared to C-RNN-GAN, here the network is completely unidirectional.
+Probably more important than the network architecture, a central piece of the model's performance is due to the final representation of the MIDI data: each time slice is represented as a one-hot encoded vector containing Note-On (128 values), Note-Off (128 values), Time-Shift (125 values), and Velocity (32 values) events \parencite{oore_this_2018}.
+
+C-RNN-GAN and Performance-RNN have not been tested and adapted to the few-shot learning problem, but more general GANs and LSTM networks have been \parencite[see][]{zhang_metagan_2018, vinyals_matching_2016}.
+However, as mentioned previously, no evaluation has been made in the musical domain.
+Therefore, this study will seek to answer the following research questions:
+
+\begin{itemize}
+    \item To what extent is the music created by a few-shot generative model comparable to the music of a generative model that is trained on the entire dataset?
+    \item To what extent is the music created by a few-shot generative model comparable to real music?
+
+\end{itemize}
+
 \chapter{Methods}
+
+To provide an answer to the above research questions, this study will develop (in Keras) two generative models based on C-RNN-GAN and Performance-RNN, where the learning procedure will incorporate the Reptile algorithm.
+Other than that, the models will not differ from their original implementation.
+These models will firstly be compared to the baseline proposed by \textcite{larochelle_few-shot_2017} (a LSTM network trained on the entire train set) and then with the music from the dataset.
+The evaluation between the models with the baseline will be based on the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018} and domain specific measurements (polyphony, scale consistency, repetitions, tone span).
+The evaluation between the model's generated samples and real music in the dataset will be based on the domain specific metric.
+Refer to the evaluation section for more details on these metrics.
+
+\section{Dataset}
+
+The dataset for this research will be similar to that of \parencite{oore_this_2018} namely, the MAESTRO Dataset \parencite{hawthorne_enabling_2018}.
+It consists of recorded MIDI data collected from each installment of the International Piano-e-Competition.
+All performances were done by piano experts on a Yamaha Disklavier, instrument which integrates a highly precise MIDI capture and playback system.
+Recorded MIDI events are of sufficient fidelity to allow judges to remotely listen to the contestant's performance (also on a Disklavier).
+
+The dataset contains MIDI recordings from nine years of the International Piano-e-Competition.
+This amounts to 1,184 piano performances, approximately 430 compositions, 6.18 million notes played, and approximately 172 hours of playback.
+There is also a recommended train/validation/test split created on the following criteria \parencite{hawthorne_enabling_2018}:
+
+\begin{itemize}
+    \item No composition should appear in more than one split
+    \item Training set is approx 80\% of the dataset (in time), and the remaining is split equally on between the validation and test sets.
+Where possible, these proportions are true also within each composer.
+    \item Popular compositions are in the training set
+    \item The validation and test splits should maintain a variety of compositions
+\end{itemize}
+
+Moreover, each performance comes with additional metadata namely, the name of the composer, the title of the performance, the suggested train, validation, test splits, year of the performance, name of the file, and duration in seconds of the performance.
+
+\section{Evaluation}
+
+Generally, DL models work by the principle of likelihood maximization, which simply says to choose the parameters that maximize the probability that the model assigns to the training data \parencite{goodfellow_nips_2016}.
+Formally, this means selecting the parameters that maximize $\prod_{i=1}^N p_{\text{model}}(\bm{x}^{(i)}; \bm{\theta})$:
+
+\begin{equation}
+    \bm{\theta^*} = \arg \max \prod_{i=1}^N p_{\text{model}}(\bm{x}^{(i)}; \bm{\theta}) \label{eq:og_mle}.
+\end{equation}
+
+However, calculating the product over many probabilities is prone to numerical problems such as underflow \parencite{goodfellow_nips_2016}.
+This is alleviated by calculating $\bm{\theta^*}$ in $\log$ space where the product is transformed into a sum:
+
+\begin{equation}
+    \bm{\theta^*} = \arg \max \sum_{i=1}^N \log p_{\text{model}}(\bm{x}^{(i)}; \bm{\theta}) \label{eg:log_mle}.
+\end{equation}
+
+That being said, calculating maximum likelihood can be thought of minimizing the Kullback-Leibler (KL) divergence: minimizing the dissimilarity between the data generating distribution $p_{\text{data}}$ and the model $p_{\text{model}}$.
+Thus, $\bm{\theta^*} = \argmin_{\bm{\theta}} D_{KL}(p_{\text{data}}(\bm{x}) \| p_{\text{model}} (\bm{x}; \bm{\theta}))$.
+The KL divergence is given by:
+
+\begin{equation}
+    DL_{KL} = \mathbb{E}_{\mathbf{x} \sim p_{\text{data}}} [\log p_{\text{data}} (\bm{x}) - \log p_{\text{model}} (\bm{x})] \label{eq:kl_div}.
+\end{equation}
+
+Having all these in mind, $\log p_{\text{data}}$ is a result of the data-generating process, and not the model.
+Therefore, a final simplification is applied where the maximum likelihood estimate would be calculated as:
+
+\begin{equation}
+    \bm{\theta^*} = - \mathbb{E}_{\mathbf{x} \sim p_{\text{data}}} [\log p_{\text{model}} (\bm{x})] \label{eq:nll}.
+\end{equation}
+
+Note that Eq.
+\ref{eg:log_mle} is the same as Eq.
+\ref{eq:nll}.
+The reason why this is brought up, is because negative log-likelihood (NLL; Eq.
+\ref{eq:nll}) is used both as a loss function and as an evaluation metric \parencite[also for generative models;][]{yu_seqgan_2016, borji_pros_2018}.
+
+However, as \textcite{borji_pros_2018} notes, NLL is uninformative about the quality of the samples generated and it does not allow to answer whether the generative network is simply memorizing training examples.
+As such, this study will also rely on other evaluation metrics when assessing generative models: the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018}, polyphony, scale consistency, repetitions, and tone span \parencite{mogren_c-rnn-gan_2016}.
+
+NDB is an evaluation metric specifically designed for generative models to measue the diversity of the generated samples.
+It follows the intuition that given two samples from the same distribution, the number of samples that fall into a bin (read as: class or cluster) should be the same \parencite{richardson_gans_2018}.
+Let $I_B(\mathbf{x})$ be the indicator function for bin $B$.
+Then $I_B(\mathbf{x}) = 1$ if the sample falls into $B$, and zero otherwise.
+Let also $\{\mathbf{x}_i^p\}$ define $N_p$ samples from a $p$ distribution (e.g., test set samples) and $\{\mathbf{x}_i^q\}$ be the $N_q$ samples from a $q$ distribution (e.g., generated samples).
+If $p = q$ then the followings is also true:
+
+\begin{equation}
+    \frac{1}{N_p} \sum_i I_B(\mathbf{x}_i^p) \approx \frac{1}{N_q} \sum_i I_B(\mathbf{x}_i^q)
+\end{equation}
+
+The pooled sample proportion $P_B$ is the proportion of samples (from the joined sets) that fall into $B$. The standard error for bin $B$ is given by:
+
+\begin{equation}
+    SE_B = \sqrt{P_B (1 - P_B)[1 / N_p + 1 / N_q]}
+\end{equation}
+
+The test statistic is a $z$-score $z = \frac{P_B^p - P_B^q}{SE_B}$ where $P_B^p$ and $P_B^q$ are the proportions of each sample that fall into $B$.
+If $z$ is smaller than a threshold (a significance level such as $\alpha = 0.05$), then the samples within that bin are statistically different.
+This whole process is repeated for each bin, and the number of statistically different bins is reported.
+The selection of bins is done through a $K$-means clustering performed on the $\mathbf{x}^p$ samples.
+Each of the generated samples $\mathbf{x}^q$ is by assigned to the nearest $L_2$ $K$ centroid \parencite{richardson_gans_2018}.
+
+The more domain-specific evaluation measures are adapted from \textcite{mogren_c-rnn-gan_2016}.
+Polyphony measures how often a minimum of two tones are played simultaneously (when the start time is the same).
+It should be noted that this is a restrictive metric: it can give a low score to music where the two notes start at different times.
+Scale consistency is the fraction of notes that are part of a standard scale (e.g., major, minor, lydian, etc.), and reporting the results for the best matching scale.
+Repetitions counts consecutive subsequences of notes, and tone span is the difference between the lowest note and highest note (counted in semi-tones) \parencite{mogren_c-rnn-gan_2016}.
+
+The choice of NDB comes from the fact that it is good at detecting overfitting \parencite{borji_pros_2018}, whereas the domain-specific metrics would favor models that generate high fidelity samples.
+In addition, these evaluation metrics have the benefit of being model-agnostic; they do not require a specific type of generative model.
 
@@
 
+\printbibliography
+
 \end{document}

### 2019-03-12 paper/proposal.tex
Added proposal, early baseline, and early reptile

--- 2019-02-13 paper/main.tex
@@
-\documentclass[a4paper]{book}
+\documentclass[11pt, a4paper, man, floatsintext]{apa6}
 
-% TODO: Add bib functionality
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\usepackage[ruled]{algorithm2e}
+
+\usepackage[american]{babel}
+\usepackage{csquotes}
+\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
+\DeclareLanguageMapping{american}{american-apa}
+\addbibresource{references.bib}
+
+\usepackage{tabularx}
+\usepackage{graphicx}
+\usepackage{subcaption}
+
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
+
+\title{Composing like a human: Adapting generative networks to few-shot learning in the musical domain}
+\shorttitle{Composing like a human}
+
+\author{Tudor Paisa\\
+        SNR: 2019551 | ANR: 315146\\
+        t.paisa@tilburguniversity.edu
+    }
+
+\note{Thesis Supervisor:\\ Dr. Menno van Zaanen}
+
+\affiliation{Tilburg University\\
+        School of Humanities and Digital Sciences\\
+        Department of Cognitive Science \& Artificial Intelligence\\
+        Tilburg, The Netherlands}
+
+\authornote{Thesis proposal submitted in partial fulfillment of the requirements for the degree of Master of Science in Data Science and Society, at the School of Humanities and Digital Sciences of Tilburg University}
 
 % Typography
-\usepackage{palatino}
-\usepackage{fontenc}
+\usepackage{times}
 
 % Links
-\usepackage{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 
@@
 
-\begin{titlepage}
-    \begin{center}
-        \huge\textbf{Thesis Title}
-        \normalsize
+\maketitle
 
-        \vspace{1cm}
+\section{Introduction}
 
-        \textbf{Tudor Paisa}\\
-        SNR: 2019551\\
-        ANR: 315146\\
+The reemergence of neural networks (NN) has brought new ways to tackle Machine Learning problems (see \cite{lecun_deep_2015} for examples), as well as new possibilities to further artificial intelligence research (such as general-purpose AI systems that play video-games; \cite{vinyals_starcraft_2017} respectively).
+Currently, two exciting areas of research in the field of Deep Learning (DL) are in designing generative models \parencite{goodfellow_nips_2016} and training NNs on small datasets \parencite{finn_model-agnostic_2017, nichol_first-order_2018}.
+This study will combine both tasks by exploring the extent to which generative models can create novel and qualitative samples of music under the constraints posed by the few-shot learning problem.
+In other words, when data is scarce, can a generative model create samples of music that are new (i.e., not in the training set), consistent with common scales, and diverse (tone sequences are not repeated over and over, combines chord structures with single-note sequences, is not restricted to - say - one octave).
+This research will evaluate the generated samples of two generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under the Reptile learning framework \parencite{nichol_first-order_2018} on small amounts of MIDI data.
 
-        \vspace{1cm}
-        \textsc{Thesis submitted in partial fulfillment\\
-        of the requirements for the degree of\\
-        Master of Science in Data Science and Society,\\
-        at the School of Humanities and Digital Sciences\\
-        of Tilburg University\\}
+DL as a field, promises to discover rich hierarchical models \parencite{lecun_deep_2015}, over various artificial intelligence applications, such as speech recognition, image classification, or text generation.
+The majority of developments usually involved discriminative models \parencite{goodfellow_generative_2014}, where high-dimensional input is assigned to a class label \parencite[e.g.,][]{krizhevsky_imagenet_2012}.
+In contrast, successes in generative models have been scarce until the discovery of the Generative Adversarial Network (GAN) by \textcite{goodfellow_generative_2014}.
+GANs have had significant successes in image super-resolution \parencite{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as a sketch into a photorealistic image;][]{isola_image--image_2016}.
+These developments lie in the field of Computer Vision, but that is not to say that we can generate only images.
+GANs (and others) have been applied in Natural Language Processing where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}.
+Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}.
+This paper will focus on the problem of generating music with the models proposed by \textcite{mogren_c-rnn-gan_2016} and \textcite{oore_this_2018}, trained under the few-shot learning problem (detailed below).
 
-        \vspace{2cm}
-        Thesis Committee:\\
-        Dr. Menno van Zaanen\\
-        Dr. SECOND READER\\
-        \vfill
+It is fairly well known that DL models tend to be "data-hungry", a caveat imposed by its optimization algorithm \parencite{chen_closer_2018, ravi_optimization_2016}.
+DL usually relies on a procedure called stochastic gradient descent (SGC) which implies computing for a given input vector the output and error, followed by calculating the input's average gradient and consequently adjusting the network's weights \parencite{lecun_deep_2015}.
+\textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates.
+Finally, they add by saying that for each new problem (e.g., new dataset), the model needs to reinitialize with a new set of random parameters which hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}.
 
-        Tilburg University\\
-        School of Humanities and Digital Sciences\\
-        Department of Cognitive Science \& Artificial Intelligence\\
-        Tilburg, The Netherlands\\
-        \today
+Besides the promise of making DL algorithms easier to train, primary reasons why this issue needs to be addressed are as follows.
+Firstly, humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, ravi_optimization_2016}, something which DL algorithms do not.
+Secondly, this would alleviate the constant need for data collection to ensure reasonable performances.
+Finally, there are many fields where the data exhibits a large number of classes and few examples per class; approaching human-type learning could allow models to properly capture this sort of data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}.
+This is known as the few-shot problem \parencite{vinyals_matching_2016}.
 
-    \end{center}
-\end{titlepage}
+In spite of its proposed benefits, developments and experiments in few-shot learning are still scarce \parencite{larochelle_few-shot_2017} and evaluations are largely concentrated on image data \parencite[e.g.,][]{lake_omniglot_2019, clouatre_figr_2019, chen_closer_2018}.
+Few-shot experiments with generative networks are even less frequent \parencite{zhang_metagan_2018, clouatre_figr_2019}.
+This research aims to explore the viability of this learning approach in the context of generating music data.
+More concretely, in light of recent developments, I will evaluate two generative networks trained on MIDI data under the constraints of few-shot learning, and compare their performance to the baseline proposed by \textcite{larochelle_few-shot_2017} and properties extracted from real music performances.
 
-\chapter*{Abstract}
-    Lorem Ipsum
+\section{Related Work}
 
-\tableofcontents
+\textcite{lake_human-level_2015} popularized the idea of learning from a few class examples using a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
+\textcite{rezende_one-shot_2016} were among the first to provide a solution to the one-shot problem in the context of a neural network (NN) by implementing Bayesian reasoning into a deep NN embedded within hierarchical latent variable models.
+The Reptile algorithm \parencite{nichol_first-order_2018} is a recent approach to this problem.
+It counteracts the shortcomings of gradient-optimization algorithms by learning the model's initial parameters in order to maximize its performance on novel tasks.
+Furthermore, it allows for any type of network to be used and does not place any restrictions on the type of loss function that can be used \parencite{nichol_first-order_2018}.
+Moreover, it avoids the computational slowdowns of its counterpart, MAML \parencite{finn_model-agnostic_2017}.
+The algorithm for Reptile is as follows:
 
-\chapter{Introduction}
+\IncMargin{1em}
+\begin{algorithm}
+Initialize $\phi$, the vector of initial parameters
 
-\chapter{Related Work}
+\For{iteration = $1, 2, \dots$}{
+    Sample task $\tau$, corresponding to loss $L_{\tau}$ on weight vectors $\tilde \phi$
 
-\chapter{Methods}
+    Compute $\tilde \phi = U_{\tau}^{k} (\phi)$, denoting $k$ steps of SGD or Adam
 
-\chapter{Results}
+    Update $\phi \leftarrow \phi + \epsilon (\tilde{\phi} - \phi)$
+}
 
-\chapter{Discussion}
+\caption{Reptile (serial version)}
+\label{alg:reptile}
+\end{algorithm}
+\DecMargin{1em}
 
-\chapter{Conclusion}
+Turning over to the generative side of things, a popular form of generating data is through the use of a Generative Adversarial Network \parencite[GAN;][]{goodfellow_generative_2014}.
+Here a generative model $G$ captures the data distribution and creates samples from it, whereas a discriminative model $D$ estimates the probability that the input is real (comes from the training data) rather than fake (comes from $G$).
+Formally, given $x$ sampled from real data, and $z$ being input noise, $D$ tries to keep $D(x)$ near $1$, while making $D(G(z))$ near $0$.
+Conversely, $G$ tries to make $D(G(z))$ near $1$ \parencite{goodfellow_nips_2016}.
+A notable implementation is C-RNN-GAN \parencite{mogren_c-rnn-gan_2016}, a continuous recurrent neural network with adversarial training.
+It consists of a network with two LSTM layers and 350 hidden units per layer, in both $G$ and $D$.
+However, $D$ has a bidirectional layout which allows it to take context from both past and future for its decisions, while $G$ is unidirectional \parencite{mogren_c-rnn-gan_2016}.
 
+Notwithstanding GANs, other generative successes have made use of adaptations on the Long Short-Term Memory (LSTM) network \parencite{hochreiter_long_1997}.
+Here, the cells of a network are designed to allow for information to persist by implementing a recursive state unit, an input gate, a forget gate, and an output gate.
+The input of the network is allowed into the state unit if it is allowed to pass through by the input gate.
+Next, the forget gate controls the self-loop of the state cell.
+Finally, the output gate has the power to shut off the output cell \parencite{goodfellow_deep_2016}.
+Here, a notable implementation is Performance-RNN \parencite{oore_this_2018}; a three layer unidirectional LSTM network with 512 units each.
+
+GANs and LSTM networks have been tested and adapted to the few-shot learning problem \parencite[see][]{zhang_metagan_2018, vinyals_matching_2016} however, no such evaluation has been made in the musical domain.
+Therefore, this study will seek to answer the following research questions:
+
+\begin{itemize}
+    \item To what extent is the music created by a few-shot generative model comparable to the music of a generative model that is trained on the entire dataset?
+    \item To what extent is the music created by a few-shot generative model comparable to real music?
+\end{itemize}
+
+\section{Methods}
+
+To provide an answer to the above research questions, this study will develop (in Keras) two generative models based on C-RNN-GAN and Performance-RNN, where the learning procedure will incorporate the Reptile algorithm.
+Other than that, the models will not differ from their original implementation.
+These models will firstly be compared to a baseline (a LSTM network of depth 1 and 200 units trained on the entire train set) and then with the music from the dataset.
+The evaluation between the models with the baseline and with the music in the dataset will be based on the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018} and domain specific measurements (polyphony, scale consistency, repetitions, tone span).
+Refer to the evaluation section for more details on these metrics.
+The data for the models will be one-hot encoded as per \textcite{oore_this_2018}; a vector for each time slice containing Note-On (128 values), Note-Off (128 values), Time-Shift (125 values), and Velocity (32 values) events.
+
+\subsection{Dataset}
+
+The dataset for this research will be similar to that of \textcite{oore_this_2018} namely, the MAESTRO Dataset \parencite{hawthorne_enabling_2018}.
+The dataset contains MIDI recordings from nine years of the International Piano-e-Competition.
+This amounts to 1,184 piano performances, approximately 430 compositions, 6.18 million notes played, and approximately 172 hours of playback.
+The dataset also features a recommended train/validation/test split.
+Moreover, each performance comes with additional metadata namely, the name of the composer, the title of the performance, the suggested train/validation/test split, year of the performance, name of the file, and duration in seconds of the performance.
+The name of the performer is not included \parencite{hawthorne_enabling_2018}.
+
+\subsection{Evaluation}
+
+Most DL models work by the principle of likelihood maximization, which simply says to choose the parameters that maximize the probability that the model assigns to the training data \parencite{goodfellow_nips_2016}.
+Formally, this means selecting the parameters that maximize $\sum_{i=1}^N \log p_{\text{model}}(\bm{x}^{(i)}; \bm{\theta})$:
+
+\begin{equation}
+    \bm{\theta^*} = \arg \max \sum_{i=1}^N \log p_{\text{model}}(\bm{x}^{(i)}; \bm{\theta}) \label{eg:log_mle}.
+\end{equation}
+
+That being said, calculating maximum likelihood can be thought of minimizing the Kullback-Leibler (KL) divergence.
+Thus, $\bm{\theta^*} = \argmin_{\bm{\theta}} \mathbb{E}_{\mathbf{x} \sim p_{\text{data}}} [\log p_{\text{data}} (\bm{x}) - \log p_{\text{model}} (\bm{x})]$.
+With this in mind, $\log p_{\text{data}}$ is a result of the data-generating process, and not the model.
+Therefore, a final simplification is applied where the maximum likelihood estimate would be calculated as:
+
+\begin{equation}
+    \bm{\theta^*} = - \mathbb{E}_{\mathbf{x} \sim p_{\text{data}}} [\log p_{\text{model}} (\bm{x})] \label{eq:nll}.
+\end{equation}
+
+Note that Eq.
+\ref{eg:log_mle} is the same as Eq.
+\ref{eq:nll}.
+The reason why this is brought up, is because negative log-likelihood (NLL; Eq.
+\ref{eq:nll}) is used both as a loss function and as an evaluation metric \parencite[][]{yu_seqgan_2016, borji_pros_2018}.
+However, as \textcite{borji_pros_2018} notes, NLL is uninformative about the quality of the samples generated and it does not allow to answer whether the generative network is simply memorizing training examples.
+As such, this study will also include other evaluation metrics when assessing generative models: the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018}, polyphony, scale consistency, repetitions, and tone span \parencite{mogren_c-rnn-gan_2016}.
+
+NDB is an evaluation metric specifically designed for generative models.
+It follows the intuition that given two samples from the same distribution, the number of samples that fall into a bin (read as: cluster) should be the same \parencite{richardson_gans_2018}.
+Let $I_B(\mathbf{x})$ be the indicator function for bin $B$.
+Then $I_B(\mathbf{x}) = 1$ if the sample falls into $B$, and zero otherwise.
+Let also $\{\mathbf{x}_i^p\}$ define $N_p$ samples from a $p$ distribution (e.g., test set samples) and $\{\mathbf{x}_i^q\}$ be the $N_q$ samples from a $q$ distribution (e.g., generated samples).
+If $p = q$ then the followings is also true:
+
+\begin{equation}
+    \frac{1}{N_p} \sum_i I_B(\mathbf{x}_i^p) \approx \frac{1}{N_q} \sum_i I_B(\mathbf{x}_i^q)
+\end{equation}
+
+The pooled sample proportion $P_B$ is the proportion of samples (from the joined sets) that fall into $B$. The standard error for bin $B$ is given by:
+
+\begin{equation}
+    SE_B = \sqrt{P_B (1 - P_B)[1 / N_p + 1 / N_q]}
+\end{equation}
+
+The test statistic is a $z$-score $z = \frac{P_B^p - P_B^q}{SE_B}$ where $P_B^p$ and $P_B^q$ are the proportions of each sample that fall into $B$.
+If $z$ is smaller than a significance level, then the samples within that bin are statistically different.
+This process is repeated for each bin, and the number of statistically different bins is reported.
+The selection of bins is done through $K$-means clustering ($K \ll N_p, N_q$) on the $\mathbf{x}^p$ samples.
+Each of the generated samples $\mathbf{x}^q$ is by assigned to the nearest ($L_2$) $K$ centroid \parencite{richardson_gans_2018}.
+
+The more domain-specific evaluation measures are adapted from \textcite{mogren_c-rnn-gan_2016}.
+Polyphony measures how often a minimum of two tones are played simultaneously (when the start time is the same).
+Scale consistency is the fraction of notes that are part of a standard scale (e.g., major, minor, lydian, etc.).
+Repetitions counts consecutive subsequences of notes, and tone span is the difference in semitones between the lowest note and highest note \parencite{mogren_c-rnn-gan_2016}.
+
+The choice of NDB comes from the fact that it is good at detecting overfitting \parencite{borji_pros_2018}, whereas the domain-specific metrics would favor models that generate high fidelity samples.
+These evaluation metrics also have the benefit of being model-agnostic; they do not require a specific generative model.
+
+\section{Milestones}
+
+\begin{table}
+    \centering
+    \begin{tabular}{l|r}
+        Task    & Date \\ \hline
+        Develop one-hot encoder for MIDI data & March 15th \\
+        Develop domain-specific evaluation tools & March 22nd \\
+        Create a Keras implementation of Performance-RNN & March 24th \\
+        Create baseline & March 24th \\
+        Create a Keras implementation of C-RNN-GAN & March 29th \\
+        Augment models with Reptile & April 5th \\
+        Train models & April 5th - April 19th \\
+        Evaluate models & April 26th \\
+        Write thesis & April 26th - May 13th \\
+    \end{tabular}
+\end{table}
+
+\printbibliography
 \end{document}

### 2019-03-12 download_data.sh

@@ -1,0 +1,7 @@
+# /bin/bash
+
+curl -O "https://storage.googleapis.com/magentadata/datasets/maestro/v1.0.0/maestro-v1.0.0-midi.zip"
+
+unzip maestro-v1.0.0-midi.zip
+
+mv "maestro-v1.0.0" data

### 2019-03-12 src/data.py

@@ -1,0 +1,300 @@
+import os
+from time import strftime, time
+
+import numpy as np
+import pandas as pd
+import pretty_midi
+import mido
+import utils
+from flags import FLAGS
+from sklearn.externals import joblib
+from keras.utils import to_categorical
+
+
+def to_seq(path):
+    midi = mido.MidiFile(path)
+    tpb = midi.ticks_per_beat  # Ticks/beat. between 384-480
+    tempo = midi.tracks[0][0].tempo  # Tempo in ms. All are 500000
+    # max_tick = mido.second2tick(second=1, ticks_per_beat=tpb, tempo=tempo)
+
+    melody = midi.tracks[1]
+
+    sequence = []
+    for msg in melody:
+        if hasattr(msg, 'note'):
+            # move_by = int(round(msg.time/8, 0)*8)
+            # Convert from ticks to miliseconds
+            move_by = int(mido.tick2second(msg.time, tpb, tempo)*1000)
+            # SORT OF quantizing by 8ms
+            move_by = int(round(move_by/8, 0)*8)
+
+            # If the time between the previous note and
+            #  the current note is greater than 1 second,
+            #  force time difference to be 1 second.
+            #  1 second at 120 BPM (default) == 2 bars
+            #  Therefore, @var:move_by will still be in time
+            if move_by > 1000:
+                move_by = 1000
+            # Append time since last message
+            sequence.append('time_move:{}'.format(move_by))
+            # Append velocity; we limit the number of possible
+            #  velocity values to 32 instead of 128
+            #  When decoding velocity must be multiplied by 32
+            #  get back the appropriate values
+            sequence.append('velocity_set:{}'.format(msg.velocity//32))
+            # Append `note_on` or `note_off`
+            sequence.append(msg.type + ':{}'.format(msg.note))
+    return sequence
+
+
+def build_sequences(data_paths):
+    sequences = []
+    for i in data_paths:
+        if data_paths.index(i) % 50 == 0:
+            progress = str(round(data_paths.index(i)/len(data_paths),
+                                 4)*100)+'%'
+            print('[ ] {} Encoding data: {}'.format(utils.now(), progress))
+
+        sequences += to_seq(i)
+    return sequences
+
+
+def build_vocab():
+    time_move = ['time_move:'+str(i) for i in range(0, 1001, 8)]  # [0:1000]
[…]

### 2019-03-12 src/flags.py

@@ -1,0 +1,8 @@
+
+FLAGS = {
+         'raw_datadir': '/home/spacewhisky/projects/thesis/data/',
+         'proc_datadir': '/home/spacewhisky/projects/thesis/data/processed/',
+         'savedir': '/home/spacewhisky/projects/thesis/generated/',
+         'fs': 100,
+         'epochs': 1000
+        }

### 2019-03-12 src/lstm_baseline.py

@@ -1,0 +1,17 @@
+import utils
+
+from keras.models import Sequential, Model
+from keras.layers import Input, Flatten, Dropout, LSTM, ReLU, Activation, Dense
+from keras.activations import relu
+
+
+def build_baseline(shape_1, shape_2, n_vocab):
+    model = Sequential()
+    model.add(LSTM(256, input_shape=(shape_1, shape_2),
+                   return_sequences=True))
+    model.add(Dropout(0.3))
+    model.add(Flatten())
+    model.add(Dense(n_vocab))
+    model.add(Activation('softmax'))
+    model.compile(loss='kullback_leibler_divergence', optimizer='rmsprop')
+    return model

### 2019-03-12 src/utils.py

@@ -1,0 +1,42 @@
+import time
+import os
+from glob import glob
+
+
+def get_file_paths(folder, extension='midi'):
+    """
+    Return full path for files ending wit extension (default: midi)
+
+    Parameters
+    ----------
+    folder : str
+        Path for where to initiate the search
+    extension : str
+        The file-type to search for
+        Default value is 'midi'
+
+    Returns
+    -------
+    paths : list
+        List of paths that contain the queried file-type (e.g., midi)
+    """
+    src = '*.' + extension
+    return [y for x in os.walk(folder)
+            for y in glob(os.path.join(x[0], src))]
+
+
+def make_folder(folder):
+    """
+    Create a folder if it doesn't exist.
+
+    Parameters
+    ----------
+    folder: str
+        Path of folder to create (if it doesn't exist)
+    """
+    if not os.path.exists(folder):
+        os.makedirs(folder)
+
+
+def now():
+    return time.strftime("%H:%M:%S")

### 2019-03-13 paper/presentation.tex
Drafted some ideas for next week's presentation

\documentclass{beamer}
\usetheme{Rochester}

\title[Composing like a human]{Composing like a human}
\subtitle{Adapting generative networks to few-shot learning in the musical domain}
\author{Tudor Paisa  \texorpdfstring{\\ (SNR: 2019551, ANR: 315146)} \and \texttt{t.paisa@tilburguniversity.edu} }
\institute{Tilburg University}
\date{\today}

\begin{document}
\begin{frame}
    \titlepage
\end{frame}

\begin{frame}
    \frametitle{Outline}
    \tableofcontents
\end{frame}

\begin{frame}
    \frametitle{Introduction}
    \section{Introduction}
    \subsection{Deep Learning}
    \begin{itemize}
        \item Deep Learning promises to discover rich hierarchical models over various applications
    \end{itemize}

\end{frame}

\end{document}

### 2019-04-19 paper/main.tex
updated with many

@@
 \documentclass[a4paper]{book}
+
+\usepackage{draftwatermark}
+\SetWatermarkText{Draft}
+\SetWatermarkColor[gray]{0.9}
 
@@
 
-\usepackage{titlesec, blindtext, color}
+\usepackage{titlesec, blindtext, xcolor}
 \definecolor{gray75}{gray}{0.75}
+\definecolor{base00}{HTML}{fafafa}
+\definecolor{base01}{HTML}{d4d4d4}
+\definecolor{base02}{HTML}{abacae}
+\definecolor{base03}{HTML}{7e8087}
+\definecolor{base04}{HTML}{555761}
+\definecolor{base05}{HTML}{333333}
+\definecolor{base06}{HTML}{4d4d4d}
+\definecolor{base07}{HTML}{666666}
+\definecolor{base08}{HTML}{a10705}
+\definecolor{base09}{HTML}{0d52bf}
+\definecolor{base0A}{HTML}{cc3b02}
+\definecolor{base0B}{HTML}{57392d}
+\definecolor{base0C}{HTML}{d48e15}
+\definecolor{base0D}{HTML}{3a9104}
+\definecolor{base0E}{HTML}{7239b3}
+\definecolor{base0F}{HTML}{667885}
+
 \newcommand{\hsp}{\hspace{10pt}}
-\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{gray75}{\thechapter \hsp \vline}\hsp}{0pt}{\Huge\bfseries}
+\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{base02}{\thechapter \hsp \vline}\hsp}{0pt}{\Huge\bfseries}
+\titleformat{\section}[hang]{\Large\bfseries}{\textcolor{base01}{\thesection \hsp \vline}\hsp}{0pt}{\Large\bfseries}
+\titleformat{\subsection}[hang]{\large\bfseries}{\textcolor{base01}{\thesubsection \hsp \vline}\hsp}{0pt}{\large\bfseries}
 
@@
 % Links
-\usepackage{hyperref}
+\usepackage[colorlinks=true, citecolor=base08, linkcolor=base08, urlcolor=base08]{hyperref}
 
@@
     \begin{center}
-        \huge\textbf{Thesis Title}
+        \huge\textbf{Composing like a human:} \\
+        \Large\textbf{Adapting generative networks to few-shot learning in the musical domain}
         \normalsize
@@
         \textbf{Tudor Paisa}\\
-        SNR: 2019551 | ANR: 315146\\
+        Student Number: 2019551 \\
+        Administration Number: 315146\\
         t.paisa@tilburguniversity.edu\\
@@
 
-\chapter{Introduction}
+\chapter{The return of the Neural Net}
 
-The reemergence of neural networks (NN) has brought new ways to tackle Machine Learning problems (such as classification; see \cite{lecun_deep_2015} for several examples), as well as new possibilities to further artificial intelligence research (such as chatbots, or general-purpose AI systems; e.g., \cite{dale_return_2016} and \cite{vinyals_starcraft_2017} respectively).
-Currently, two exciting areas of research in the field of Deep Learning (DL) are in designing generative models \parencite{goodfellow_nips_2016} and training NNs on small datasets \parencite{finn_model-agnostic_2017, nichol_first-order_2018}.
+Over the past sixty years, Artificial Neural Networks (ANNs) have experienced a number of cycles in popularity noted by extensive publicity of over-inflated expectations on the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs [CITATION NEEDED]): that a system could learn to do anything it was programmed to do [CITATION NEEDED] \parencite{minsky_perceptrons_1988}.
+However, each wave of praise was followed by a mass of public disappointment, criticism, and funding cuts [CITATION NEEDED].
+These upturns and downturns of the field of connectionism - or more generally A.I.
+- are now known in the literature as the A.I.
+winters [CITATION NEEDED].
+
+The recent revival of ANNs under the scope of Deep Learning (DL) has brought new ways to tackle Machine Learning problems, most notably in the context of classification (but also regression, although less popular; see \cite{lecun_deep_2015} for several examples of DL applications).
+In addition, with the advent of chatbots \parencite{dale_return_2016}, voice-activated personal assistants \parencite{xiong_microsoft_2018}, and general-purpose A.I.
+systems \parencite{vinyals_starcraft_2017}, DL introduced new frontiers in artificial intelligence research.
+
+% NOTE: USE ME LATER
+% Currently, two exciting areas of research in the field of Deep Learning (DL) are in designing generative models \parencite{goodfellow_nips_2016} and training NNs on small datasets \parencite{finn_model-agnostic_2017, nichol_first-order_2018}.
 This research will combine both of these tasks by exploring the extent to which generative models can create novel and qualitative samples of music under the constraints posed by the few-shot learning problem.
@@
 
-DL as a field, promises to discover rich hierarchical models \parencite{lecun_deep_2015}, over various artificial intelligence applications, such as speech recognition, image classification, or text generation.
-The majority of developments usually involved discriminative models \parencite{goodfellow_generative_2014}, where high-dimensional input is assigned to a class label \parencite{zhang_character-level_2015, krizhevsky_imagenet_2012}.
-In contrast, successes in generative models have been scarce, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable for the longest run, followed years later by Bidirectional Long Short Term Memory \parencite{graves_generating_2013} and Variational Autoencoders \parencite{kingma_auto-encoding_2013}.
+Broadly speaking, DL as a field, promises to solve tasks which are easily (intuitively) performed by people, but hard to formalize \parencite[e.g., recognizing faces or spoken word;][]{goodfellow_deep_2016}.
+The solution comprises of enabling computers to learn form experience and understand the world through the discovery of hierarchical relationships between concepts \parencite{lecun_deep_2015}.
+This way, there is no need for manual input on the sort of knowledge that the computer needs \parencite{goodfellow_deep_2016}.
+A more simplistic and concise interpretation of DL would be to see it as a statistical technique for identifying patterns in sample data using large (deep) ANNs \parencite{marcus_deep_2018}.
 
-However, studies on generative architectures were fairly uncommon until the discovery of the Generative Adversarial Network (GAN) by \textcite{goodfellow_generative_2014}.
-GANs have had significant successes in image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, image-to-image translation such as a satellite image into a map, or a sketch into a photorealistic image \parencite[;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.
+The reasons for presenting these facts will be clarified in the following sections however, a terse argumentation is that progress towards general A.I.
+systems is marked by scant explorations of ANNs outside of the classification task, and even more so when trying to overcome some of their shortcomings, such as being highly reliant on massive amounts of data [CITATION NEEDED].
+% NOTE: Maybe I was too harsh?
+
+\section{We do not discriminate}\label{sec:no_disc}
+
+Typically, in a neural network, the data samples go through a set of input units (that might represent pixels, word embeddings, etc.), then through multiple hidden layers, each with a given number of nodes, and reaching the output units where the answer is given \parencite{marcus_deep_2018}.
+An overwhelming majority of experiments with ANNs involve discriminative models \parencite{goodfellow_generative_2014}, where the goal is to assign the high-dimensional inputs to a class label \parencite[such as an animal or piece of furniture;][]{zhang_character-level_2015, krizhevsky_imagenet_2012}.
+However, other possible applications of DL models include (but are not limited to): regression [CITATION NEEDED], data compression \parencite{cheng_deep_2018}, and generating new data points \parencite{graves_generating_2013}.
+
+For the purposes of this paper (see Section \ref{sec:thesis_goal}), generative model are central to our experiments (Chapter \ref{chap:methods}).
+Borrowing the analogy from \textcite{goodfellow_nips_2016}, one might wonder what is the value of studying generative models.
+After all, in the image domain, such a framework would merely generate more images (something which the Internet as no shortage of).
+Some of the more eloquent answers include using generative models for simulating possible futures to be later used for planning [CITATION NEEDED], providing predictions to missing data, or enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.
+
+However, successes in generative models have been scarce, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run.
+Yet, this method requires generating each sample one at a time (ramping up the cost of the process to $O(n)$).
+Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}.
+Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task.
+Their main drawback is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.
+
+The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this field.
+Besides countering the drawbacks outlined above \parencite{goodfellow_nips_2016}, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.
+
+Although not as popular as the abovementioned, Recurrent Neural Networks (RNNs) have been successfully implemented for various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}.
+More specifically, a variation of these, Long Short-Term Memory Networks \parencite[LSTMs;][]{hochreiter_long_1997}, are of particular interest to this task.
+RNNs (and by extension LSTMs) contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}.
+This makes them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc \parencite{fan_tts_2014}.
 
@@
 
+\section{"Just... One more bite"}\label{sec:hungry}
+
 It is fairly well known that DL models tend to be very "data-hungry", a caveat imposed by its optimization algorithm \parencite{chen_closer_2018, ravi_optimization_2016}.
@@
 This is known as the few-shot problem \parencite{chen_closer_2018, vinyals_matching_2016, larochelle_few-shot_2017}.
+
+\section{Making music like a hooman} \label{sec:thesis_goal}
 
@@
 
-\chapter{Methods}
+\chapter{Methods}\label{chap:methods}
 

### 2019-04-19 paper/presentation.tex
updated with many

@@
 \subtitle{Adapting generative networks to few-shot learning in the musical domain}
-\author{Tudor Paisa  \texorpdfstring{\\ (SNR: 2019551, ANR: 315146)} \and \texttt{t.paisa@tilburguniversity.edu} }
+\author{Tudor Paisa  \texorpdfstring{\\ (SNR: 2019551, ANR: 315146)}{} \texorpdfstring{\\ \texttt{t.paisa@tilburguniversity.edu}}{} }
 \institute{Tilburg University}
@@
 
+%\begin{frame}
+    %\frametitle{Outline}
+    %\tableofcontents
+%\end{frame}
+
 \begin{frame}
-    \frametitle{Outline}
-    \tableofcontents
+    \frametitle{Introduction}
+    \section{Introduction}
+    \begin{block}{Deep Learning}
+    \begin{itemize}
+        \item Deep Learning promises to discover rich hierarchical models over various applications
+        \item The majority of developments usually involve discriminative models
+        \item In recent years...
+            \begin{itemize}
+            \item Generative Models
+            \item Few-Shot Learning
+            \end{itemize}
+    \end{itemize}
+\end{block}
 \end{frame}
@@
     \frametitle{Introduction}
-    \section{Introduction}
-    \subsection{Deep Learning}
+    \begin{block}{Generative Models}
         \begin{itemize}
-        \item Deep Learning promises to discover rich hierarchical models over various applications
+            \item A Neural Network that is able to generate data
+            \item Simple concept...
+                \begin{itemize}
+                    \item ... slow developments...
                 \end{itemize}
+            \item ... until frameworks like GANs \& LSTM
+            \item Widely researched in Computer Vision, Natural Language Processing, not so much in arts (music especially)
+        \end{itemize}
+    \end{block}
+\end{frame}
 
+\begin{frame}
+    \frametitle{Introduction}
+    \begin{block}{Few-Shot Learning}
+        \begin{itemize}
+            \item Neural Networks tend to be "data-hungry"
+            \item SGD does not allow convergence in a limited number of steps
+            \item For AI applications this sucks
+                \begin{itemize}
+                    \item Humans generalize after one/few examples
+                    \item Vanilla neural nets do not
+                \end{itemize}
+            \item Few-Shot Learning: making neural networks generalize using by showing it few instances of each class
+            \item Solutions? Yes! More on that later
+        \end{itemize}
+    \end{block}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Introduction}
+    \begin{block}{Getting There...}
+        \begin{itemize}
+            \item Overwhelming majority of few-shot generative developments have been in Computer Vision
+            \item Then some in Natural Language Processing
+            \item Musical domain largely unexplored
+        \end{itemize}
+    \end{block}
+
+    \begin{block}{The research!}
+        \begin{itemize}
+            \item Adapt two generative models to few-shot learning
+            \item Evaluate their ability to create music
+        \end{itemize}
+    \end{block}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Theoretical Framework}
+    \begin{block}{Music Generators}
+        There are two candidates:
+    \end{block}
+        \begin{block}{C-RNN-GAN}
+            \begin{minipage}[b]{0.48\textwidth}
+                $G$enerator \\
+                2 $\times$ \textbf{unidirectional} LSTM \\
+                350 hidden units
+            \end{minipage}
+            \hfill
+            \begin{minipage}[b]{0.48\textwidth}
+                $D$iscriminator \\
+                2 $\times$ \textbf{bidirectional} LSTM \\
+                350 hidden units
+            \end{minipage}
+        \end{block}
+        \begin{block}{PerformanceRNN}
+            3 $\times$ \textbf{unidirectional} LSTM \\
+            512 hidden units
+        \end{block}
 \end{frame}

### 2019-04-19 paper/thesis.tex
updated with many

--- 2019-03-12 paper/main.tex
@@
-\documentclass[a4paper]{book}
+%\documentclass[a4paper]{book}
 
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+%\usepackage[utf8]{inputenc}
+%\usepackage[T1]{fontenc}
 
-\usepackage{titlesec, blindtext, color}
-\definecolor{gray75}{gray}{0.75}
-\newcommand{\hsp}{\hspace{10pt}}
-\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{gray75}{\thechapter \hsp \vline}\hsp}{0pt}{\Huge\bfseries}
+%\usepackage{titlesec, blindtext, color}
+%\definecolor{gray75}{gray}{0.75}
+%\newcommand{\hsp}{\hspace{10pt}}
+%\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{gray75}{\thechapter \hsp \vline}\hsp}{0pt}{\Huge\bfseries}
 
-\usepackage[american]{babel}
-\usepackage{csquotes}
-\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
-\DeclareLanguageMapping{american}{american-apa}
-\addbibresource{references.bib}
+%\usepackage[american]{babel}
+%\usepackage{csquotes}
+%\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
+%\DeclareLanguageMapping{american}{american-apa}
+%\addbibresource{references.bib}
 
-% Essentials
-\usepackage{tabularx}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage[ruled]{algorithm2e}
+%% Essentials
+%\usepackage{tabularx}
+%\usepackage{graphicx}
+%\usepackage{subcaption}
+%\usepackage[ruled]{algorithm2e}
 
-% Typography
-\usepackage{palatino}
-\usepackage{fontenc}
+%% Typography
+%\usepackage{palatino}
+%\usepackage{fontenc}
 
-% Links
-\usepackage{hyperref}
+%% Links
+%\usepackage{hyperref}
 
-% Math stuff
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amstext}
-\usepackage{bm}
+%% Math stuff
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%\usepackage{amstext}
+%\usepackage{bm}
 
-% Commands
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
+%% Commands
+%\DeclareMathOperator*{\argmin}{arg\,min}
+%\DeclareMathOperator*{\argmax}{arg\,max}
 
-\begin{document}
+%\begin{document}
 
-\begin{titlepage}
-    \begin{center}
-        \huge\textbf{Thesis Title}
-        \normalsize
+%\begin{titlepage}
+%    \begin{center}
+%        \huge\textbf{Thesis Title}
+%        \normalsize
 
-        \vspace{1cm}
+%        \vspace{1cm}
 
-        \textbf{Tudor Paisa}\\
-        SNR: 2019551 | ANR: 315146\\
-        t.paisa@tilburguniversity.edu\\
+%        \textbf{Tudor Paisa}\\
+%        SNR: 2019551 | ANR: 315146\\
+%        t.paisa@tilburguniversity.edu\\
 
-        \vspace{1cm}
-        \textsc{Thesis submitted in partial fulfillment\\
-        of the requirements for the degree of\\
-        Master of Science in Data Science and Society,\\
-        at the School of Humanities and Digital Sciences\\
-        of Tilburg University\\}
+%        \vspace{1cm}
+%        \textsc{Thesis submitted in partial fulfillment\\
+%        of the requirements for the degree of\\
+%        Master of Science in Data Science and Society,\\
+%        at the School of Humanities and Digital Sciences\\
+%        of Tilburg University\\}
 
-        \vspace{2cm}
-        Thesis Committee:\\
-        Dr. Menno van Zaanen\\
-        Dr. SECOND READER\\
-        \vfill
+%        \vspace{2cm}
+%        Thesis Committee:\\
+%        Dr. Menno van Zaanen\\
+%        Dr. SECOND READER\\
+%        \vfill
 
-        Tilburg University\\
-        School of Humanities and Digital Sciences\\
-        Department of Cognitive Science \& Artificial Intelligence\\
-        Tilburg, The Netherlands\\
-        \today
+%        Tilburg University\\
+%        School of Humanities and Digital Sciences\\
+%        Department of Cognitive Science \& Artificial Intelligence\\
+%        Tilburg, The Netherlands\\
+%        \today
 
-    \end{center}
-\end{titlepage}
+%    \end{center}
+%\end{titlepage}
 
-\chapter*{Abstract}
-    Lorem Ipsum
+%\chapter*{Abstract}
+%    Lorem Ipsum
 
-\tableofcontents
+%\tableofcontents
 
@@
 
-\end{document}
+%\end{document}

### 2019-04-19 src/data.py

@@ -6,9 +6,6 @@ import pandas as pd
 import pretty_midi
 import mido
 import utils
-from flags import FLAGS
-from sklearn.externals import joblib
-from keras.utils import to_categorical
 
 
 def to_seq(path):
@@ -24,9 +21,9 @@ def to_seq(path):
         if hasattr(msg, 'note'):
             # move_by = int(round(msg.time/8, 0)*8)
             # Convert from ticks to miliseconds
-            move_by = int(mido.tick2second(msg.time, tpb, tempo)*1000)
+            move_by = int(mido.tick2second(msg.time, tpb, tempo) * 1000)
             # SORT OF quantizing by 8ms
-            move_by = int(round(move_by/8, 0)*8)
+            move_by = int(round(move_by / 8, 0) * 8)
 
             # If the time between the previous note and
             #  the current note is greater than 1 second,
@@ -41,7 +38,7 @@ def to_seq(path):
             #  velocity values to 32 instead of 128
             #  When decoding velocity must be multiplied by 32
             #  get back the appropriate values
-            sequence.append('velocity_set:{}'.format(msg.velocity//32))
+            sequence.append('velocity_set:{}'.format(msg.velocity // 32))
             # Append `note_on` or `note_off`
             sequence.append(msg.type + ':{}'.format(msg.note))
     return sequence
@@ -49,22 +46,28 @@ def to_seq(path):
 
 def build_sequences(data_paths):
     sequences = []
-    for i in data_paths:
-        if data_paths.index(i) % 50 == 0:
-            progress = str(round(data_paths.index(i)/len(data_paths),
-                                 4)*100)+'%'
-            print('[ ] {} Encoding data: {}'.format(utils.now(), progress))
+    size = len(data_paths)
 
+    # If we passed only one path; i.e., a string with
+    #  path to midi
+    if type(data_paths) == str:
+        return to_seq(data_paths)
+
+    for i in data_paths:
+        if data_paths.index(i) % 100 == 0:
+            loc = data_paths.index(i) + 1
+            print(f'[ ] {utils.now()} Encoding song {loc}/{size}')
         sequences += to_seq(i)
     return sequences
 
 
 def build_vocab():
-    time_move = ['time_move:'+str(i) for i in range(0, 1001, 8)]  # [0:1000]
-    note_on = ['note_on:'+str(i) for i in range(0, 128)]  # [0:127]
-    note_off = ['note_off:'+str(i) for i in range(0, 128)]  # [0:127]
[…]

### 2019-04-19 src/flags.py

@@ -1,8 +1,102 @@
+import argparse
 
-FLAGS = {
-         'raw_datadir': '/home/spacewhisky/projects/thesis/data/',
-         'proc_datadir': '/home/spacewhisky/projects/thesis/data/processed/',
-         'savedir': '/home/spacewhisky/projects/thesis/generated/',
-         'fs': 100,
-         'epochs': 1000
-        }
+
+class Flags:
+    def __init__(self):
+        self.parser = argparse.ArgumentParser(
+            formatter_class=argparse.ArgumentDefaultsHelpFormatter)
+        self.parser.add_argument(
+            '--model-name',
+            help="""name of the model. can be either of the following: 
+            'baseline', 'performance_rnn', 'c_rnn_gan'""",
+            default='baseline',
+            type=str)
+        self.parser.add_argument(
+            '--raw_datadir',
+            help='folder where the midi data is',
+            default='../data/',
+            type=str)
+        self.parser.add_argument(
+            '--proc_datadir',
+            help='folder of the processed midi',
+            default='../data/processed/',
+            type=str)  # DEPRECATED?
+        self.parser.add_argument(
+            '--model-folder',
+            help='folder where the model will be saved',
+            default='../models/',
+            type=str)
+        self.parser.add_argument(
+            '--train-file',
+            help='name of the file as found in the `data` folder',
+            default='CHANGE_ME',
+            type=str)
+        self.parser.add_argument(
+            '--test-file',
+            help='name of the file as found in the `data` folder',
+            default='CHANGE_ME',
+            type=str)
+        self.parser.add_argument(
+            '--layer-size',
+            help='size of the hidden units in the LSTM network',
+            default=256,
+            type=int)
+        self.parser.add_argument(
+            '--dropout',
+            help='rate of dropping out units in a network',
+            default=0.0,
+            type=float)
+        self.parser.add_argument(
+            '--n-classes',
+            help='number of classes per inner task',
+            default=5,
+            type=int)
+        self.parser.add_argument(
+            '--n-shots',
+            help='number of examples per class',
+            default=1,
+            type=int)
+        self.parser.add_argument(
+            '--inner-batch', help='inner batch size', default=64, type=int)
+        self.parser.add_argument(
+            '--inner-iters', help='inner iterations', default=5, type=int)
+        self.parser.add_argument(
+            '--learning-rate',
+            help='Adam step size',
+            default=0.0025,
+            type=float)
[…]

### 2019-04-19 src/lstm_baseline.py

@@ -1,17 +1,273 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
 import utils
+import numpy as np
+from copy import deepcopy
+from tqdm import tqdm
 
-from keras.models import Sequential, Model
-from keras.layers import Input, Flatten, Dropout, LSTM, ReLU, Activation, Dense
-from keras.activations import relu
+import pdb
 
 
-def build_baseline(shape_1, shape_2, n_vocab):
-    model = Sequential()
-    model.add(LSTM(256, input_shape=(shape_1, shape_2),
-                   return_sequences=True))
-    model.add(Dropout(0.3))
-    model.add(Flatten())
-    model.add(Dense(n_vocab))
-    model.add(Activation('softmax'))
-    model.compile(loss='kullback_leibler_divergence', optimizer='rmsprop')
-    return model
+class Baseline(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=415,
+                 dropout=0.0,
+                 num_layers=1,
+                 init_dim=32,
+                 use_bias=False,
+                 is_bidirectional=False):
+        super(Baseline, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = out_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+
+        if is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=use_bias,
[…]

### 2019-04-19 src/old/run.py

@@ -1,0 +1,81 @@
+import os
+import data
+import utils
+import numpy as np
+
+from flags import FLAGS
+from lstm_baseline import build_baseline
+
+# check if the data has been processed
+data_paths = utils.get_file_paths(FLAGS['raw_datadir'], extension='midi')
+
+# Constants
+innerstepsize = 0.02  # stepsize of inner SGD
+innerepochs = 1  # nr. epochs for each inner SGD
+outerstepsize0 = 0.1  # stepsize of outer optimization (meta-optimization)
+n_iterations = 5  # number of outer updates
+n_train = 10  # size of minibatch
+seq_len = 128
+vocab = data.build_vocab()
+vocab_len = len(vocab)
+
+rng = np.random.RandomState(1337)
+
+# sequences = data.build_sequences(data_paths[:5])
+# print('[ ] {} One-Hot Encoding'.format(utils.now()))
+# one_hot = data.one_hot(sequences, vocab)
+
+# TODO: Init model
+# inp, out = data.create_io_sequences(sequences, 128, vocab_len)
+# n_vocab = out.shape[1]
+
+# import sys
+# sys.exit()
+
+lstm = build_baseline(seq_len, 1, vocab_len)
+
+# TODO: Batch from data + load batch (i.e., create task)
+# TODO: Train on batched data
+
+
+def train_on_batch(x, y):
+    weights_before = lstm.get_weights()
+    lstm.train_on_batch(x, y)
+    weights_after = lstm.get_weights()
+    for i in range(len(weights_after)):
+        lstm.weights[i] = (weights_after[i] -
+                          (weights_after[i] - weights_before[i])*innerstepsize)
+
+
+for iteration in range(n_iterations):
+    print('[ ] {} Outer epoch: {}'.format(utils.now(), iteration))
+    weights_before = lstm.get_weights()
+
+    # Generate a new task
+    inp, out = data.gen_task(data_paths, rng, n_samples=5, seq_len=seq_len,
+                             vocab_len=vocab_len)
+
+    # shuffle data... kind of
+    idx = rng.permutation(len(inp))
+
+    for inner in range(innerepochs):
+        print('[ ] {}\tInner epoch: {}'.format(utils.now(), inner))
+        for start in range(0, len(inp), n_train):
+            # minibatch
+            mini_idx = idx[start:start+n_train]
+            train_on_batch(inp[mini_idx], out[mini_idx])
+
[…]

### 2019-04-19 src/performance_rnn.py

@@ -1,0 +1,282 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+
+import os
+import data
+import utils
+import numpy as np
+import pdb
+from copy import deepcopy
+
+
+class PerformanceRNN(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 seq_len,
+                 num_layers=3,
+                 dropout=0.3,
+                 out_dim=414):
+        super(PerformanceRNN, self).__init__()
+        self.in_dim = in_dim
+        self.num_layers = num_layers
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.seq_len = seq_len
+        self.dropout = dropout
+        self.out_dim = out_dim
+
+        self.lstm = nn.LSTM(
+            input_size=self.in_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=self.num_layers,
+            bias=False,
+            batch_first=True,  # Input is provided as (batch, seq, feature)
+            dropout=self.dropout,
+            bidirectional=False)
+
+        self.linear = nn.Linear(self.hidden_dim * self.seq_len, self.out_dim)
+
+    def init_hidden(self):
+        device = torch.device("cuda:0" if torch.cuda.is_available() else "cpu")
+        self.hidden = (torch.zeros(3, self.batch_size,
+                                   self.hidden_dim).to(device),
+                       torch.zeros(3, self.batch_size,
+                                   self.hidden_dim).to(device))
+
+    def forward(self, x):
+        # Input is shape: (batch, seq, feature).
+        out, self.hidden = self.lstm(x.float(), self.hidden)
+
+        y_hat = self.linear(out.contiguous().view(self.batch_size, -1))
+
+        return y_hat
+
+
+def train_model(data_paths,
+                vocab,
+                rng,
+                in_dim=1,
+                hidden_dim=256,
+                seq_len=128,
+                batch_size=32,
+                lr=0.025,
+                dropout=0.3,
+                epochs=1000,
+                save_path='/home/spacewhisky/projects/thesis/models/'):
+    """NORMAL (i.e., not with Reptile) training of
+    PerformanceRNN. Used to test if the model works
+    """
+    device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
+    model = PerformanceRNN(
+        in_dim=in_dim,
[…]

### 2019-04-19 src/run.py

@@ -2,80 +2,130 @@ import os
 import data
 import utils
 import numpy as np
+import pandas as pd
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+import json
+from pathlib import Path
 
-from flags import FLAGS
-from lstm_baseline import build_baseline
+with open('../params/test.json', 'r') as f:
+    FLAGS = json.load(f)
 
-# check if the data has been processed
-data_paths = utils.get_file_paths(FLAGS['raw_datadir'], extension='midi')
+RNG = np.random.RandomState(666)
+VOCAB = data.build_vocab()
+IN_DIM = FLAGS['seq_len']
 
-# Constants
-innerstepsize = 0.02  # stepsize of inner SGD
-innerepochs = 1  # nr. epochs for each inner SGD
-outerstepsize0 = 0.1  # stepsize of outer optimization (meta-optimization)
-n_iterations = 5  # number of outer updates
-n_train = 10  # size of minibatch
-seq_len = 128
-vocab = data.build_vocab()
-vocab_len = len(vocab)
+assert FLAGS['model_name'] in ['baseline', 'performance_rnn',
+                               'c_rnn_gan'], 'Invalid model name.'
 
-rng = np.random.RandomState(1337)
+df = pd.read_csv(os.path.join(FLAGS['raw_datadir'], 'maestro_updated.csv'))
+test_df = df[df['split'] == 'test']
+train_df = df[df['split'] == 'train']
 
-# sequences = data.build_sequences(data_paths[:5])
-# print('[ ] {} One-Hot Encoding'.format(utils.now()))
-# one_hot = data.one_hot(sequences, vocab)
+train_data_paths = [
+    os.path.join(FLAGS['raw_datadir'], i) for i in train_df['midi_filename']
+]
+test_data_paths = [
+    os.path.join(FLAGS['raw_datadir'], i) for i in test_df['midi_filename']
+]
 
-# TODO: Init model
-# inp, out = data.create_io_sequences(sequences, 128, vocab_len)
-# n_vocab = out.shape[1]
+utils.make_folder(FLAGS['model_folder'])
 
-# import sys
-# sys.exit()
+if FLAGS['model_name'] == 'baseline':
+    from lstm_baseline import train_baseline_large
 
-lstm = build_baseline(seq_len, 1, vocab_len)
+    checkpoint = Path(
+        os.path.join(FLAGS['model_folder'],
+                     FLAGS['model_name'] + '_checkpoint'))
 
[…]

### 2019-04-23 src/data.py
some update regarding something

@@ -241,16 +241,23 @@ def pad_song(seq, seq_len=1, vocab_len=415, max_song_len=129126):
 
 
 def batch(seq_lens, batch_size, window_size, stride_size, rng):
+    # Fuck knows what i've done in these lines
     positions = [(i, range(j, j + window_size))
                  for i, seqlen in enumerate(seq_lens)
                  for j in range(0, seqlen - window_size, stride_size)]
 
-    positions = [(sum(seq_lens[:i]) + r.start, sum(seq_lens[:i]) + r.stop)
-                 for i, r in positions]
+    positions = [(sum(seq_lens[:i]) + r.stop - window_size,
+                  sum(seq_lens[:i]) + r.stop) for i, r in positions]
     dummy = [i for i in range(len(positions))]
 
-    indices = rng.choice(
-        dummy, size=(len(positions) // batch_size, batch_size), replace=False)
+    if len(positions) < batch_size:
+        repl = True
+        n_batches = batch_size // len(positions)
+    else:
+        repl = False
+        n_batches = len(positions) // batch_size
+
+    indices = rng.choice(dummy, size=(n_batches, batch_size), replace=repl)
 
     # for i in seq_lens:
     #     start = sum(seq_lens[:seq_lens.index(i)])
@@ -260,6 +267,39 @@ def batch(seq_lens, batch_size, window_size, stride_size, rng):
     #         start = start + stride_size
 
     return [[positions[j] for j in i] for i in indices]
+
+
+def calculate_grad_norm(params, norm_type=2):
+    total_norm = 0
+    for p in params:
+        param_norm = p.grad.data.norm(norm_type)
+        total_norm += param_norm**norm_type
+    total_norm = total_norm**(1.0 / norm_type)
+    return total_norm
+
+
+def midi2piano(path, fs=100):
+    """
+    Transform MIDI file to piano roll
+
+    Parameters
+    ----------
+    path : str
+        Path of MIDI file to open
+    fs : int
+        Sampling frequency of the rows; i.e., each row is spaced
+        apart by 1/fs
+
+    Returns
+    -------
+    piano_roll : np.ndarray, shape=(times, 128)
+        Piano roll of the Instrument
+    """
+    song = pretty_midi.PrettyMIDI(path)
+    inst = song.instruments[0]
+    piano = inst.get_piano_roll(fs=fs)
+    # piano = piano.astype('i')
+    return piano.T
 
 
 #########################################################################
[…]

### 2019-04-23 src/evaluation.py

@@ -1,0 +1,170 @@
+from tqdm import tqdm
+from data import midi2piano
+import mido
+
+
+def open_midi(path):
+    midi = mido.MidiFile(path)
+    tpb = midi.ticks_per_beat
+    tempo = midi.tracks[0][0].tempo
+    melody = midi.tracks[1]
+    return melody, tpb, tempo
+
+
+def polyphony(melody, tpb, tempo, criterion=31.25):
+    # Default criterion is a 64th note in ms
+    # At 120 BPM, a 64th is pretty fast, and two notes
+    #  played at a distance of ~31ms sounds like a
+    #  chord. The same applies for 32nd notes, but
+    #  given that they are more common that the 64s
+    #  we won't be using them as a selection criterion
+
+    poly = 0
+    notes = len([True for i in melody if hasattr(i, 'note')])
+
+    # We loop through the original melody because
+    #  there might be control changes, or something
+    #  of the sort. Nevertheless, the point is they
+    #  have time deltas as well, and it would be silly
+    #  to ignore them
+    for i in range(1, len(melody)):
+        if hasattr(melody[i], 'note') and hasattr(melody[i - 1], 'note'):
+            delta_time = int(
+                mido.tick2second(melody[i].time, tpb, tempo) * 1000)
+
+            if delta_time <= criterion:
+                poly += 1
+
+    return poly / notes
+
+
+def repetitions(melody):
+    count = 0
+    notes = [i.note for i in melody if hasattr(i, 'note')]
+    for i in range(1, len(notes)):
+        if notes[i] == notes[i - 1]:
+            count += 1
+
+    return count
+
+
+def tone_span(melody):
+    notes = [i.note for i in melody if hasattr(i, 'note')]
+
+    return max(notes) - min(notes)
+
+
+def build_interval(scale='major'):
+    scales = [
+        'major', 'har_minor', 'mel_minor', 'whole', 'diminished',
+        'major_penta', 'minor_penta', 'jap_in_sen', 'blues', 'dorian',
+        'phrygian', 'lydian', 'mixolydian', 'aeolian', 'locrian'
+    ]
+
+    assert scale in scales, f'Wrong scale. Must be one of {scales}'
+
+    if scale == 'major':
+        return [2, 2, 1, 2, 2, 2, 1]
+    if scale == 'har_minor':
+        return [2, 1, 2, 2, 1, 3, 1]
[…]

### 2019-04-23 src/lstm_baseline.py

@@ -174,6 +174,7 @@ class Baseline(nn.Module):
 
 def train_baseline_large(vocab,
                          rng,
+                         seq_file,
                          seq_lens,
                          in_dim=415,
                          hidden_dim=256,
@@ -188,9 +189,14 @@ def train_baseline_large(vocab,
                          stride_size=200,
                          use_bias=False,
                          is_bidirectional=False,
+                         clip_grad=True,
+                         clip_norm=3,
+                         max_norm=1.0,
                          save_path='../models/',
                          model_state=None,
-                         optimizer_state=None):
+                         optimizer_state=None,
+                         loss_saved=None,
+                         grad_norm=None):
     # Use GPU if we have one
     device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
 
@@ -215,14 +221,23 @@ def train_baseline_large(vocab,
     if optimizer_state != None:
         optimizer.load_state_dict(optimizer_state)
 
-    loss_saved = []  # for visualization purposes
+    if loss_saved != None:
+        loss_saved = loss_saved
+    else:
+        loss_saved = []  # for visualization purposes
+    if grad_norm != None:
+        grad_norm = grad_norm
+    else:
+        grad_norm = []  # for visualization purposes
 
     vocab = {i: vocab.index(i) for i in vocab}  # list to dict
 
-    sequences = np.load('../data/processed/all_songs.npy')
+    sequences = np.load(seq_file)
     for epoch in range(meta_iters_start, meta_iters):
         batches = data.batch(seq_lens, batch_size, window_size, stride_size,
                              rng)
+        norm_epoch = []
+        loss_epoch = []
 
         for batch in tqdm(batches, desc=f'Epoch {epoch+1}'):
             y = [sequences[i[0]:i[1]] for i in batch]
@@ -235,39 +250,72 @@ def train_baseline_large(vocab,
                 init, window_size, y=y[:-1], output_type='logit')
             # assert outputs.shape[:2] == y.shape[:2]
 
-            loss = F.nll_loss(outputs.contiguous().view(-1, in_dim),
-                              y.contiguous().view(-1))
+            loss = F.cross_entropy(outputs.contiguous().view(-1, in_dim),
+                                   y.contiguous().view(-1))
             model.train()
             model.zero_grad()
             loss.backward()
 
+            norm = data.calculate_grad_norm(model.parameters(), clip_norm)
+            norm_epoch.append(norm.cpu().numpy())
[…]

### 2019-04-23 src/performance_rnn.py

@@ -2,13 +2,17 @@ import torch
 import torch.nn as nn
 import torch.optim as optim
 import torch.nn.functional as F
+from torch.distributions import Categorical
 
 import os
 import data
 import utils
 import numpy as np
-import pdb
+import pandas as pd
 from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
 
 
 class PerformanceRNN(nn.Module):
@@ -16,267 +20,508 @@ class PerformanceRNN(nn.Module):
                  in_dim,
                  hidden_dim,
                  batch_size,
-                 seq_len,
-                 num_layers=3,
-                 dropout=0.3,
-                 out_dim=414):
+                 out_dim=415,
+                 dropout=0.0,
+                 num_layers=1,
+                 init_dim=32,
+                 use_bias=False,
+                 is_bidirectional=False):
         super(PerformanceRNN, self).__init__()
-        self.in_dim = in_dim
-        self.num_layers = num_layers
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
         self.hidden_dim = hidden_dim
         self.batch_size = batch_size
-        self.seq_len = seq_len
-        self.dropout = dropout
         self.out_dim = out_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+
+        if is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
 
         self.lstm = nn.LSTM(
-            input_size=self.in_dim,
+            input_size=self.hidden_dim,
             hidden_size=self.hidden_dim,
-            num_layers=self.num_layers,
-            bias=False,
-            batch_first=True,  # Input is provided as (batch, seq, feature)
-            dropout=self.dropout,
-            bidirectional=False)
+            num_layers=int(self.num_layers),
+            bias=use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=float(dropout),
[…]

### 2019-04-23 src/run.py

@@ -1,4 +1,6 @@
 import os
+from pathlib import Path
+import json
 import data
 import utils
 import numpy as np
@@ -7,29 +9,19 @@ import torch
 import torch.nn as nn
 import torch.optim as optim
 import torch.nn.functional as F
-import json
-from pathlib import Path
 
 with open('../params/test.json', 'r') as f:
     FLAGS = json.load(f)
 
 RNG = np.random.RandomState(666)
 VOCAB = data.build_vocab()
-IN_DIM = FLAGS['seq_len']
 
 assert FLAGS['model_name'] in ['baseline', 'performance_rnn',
                                'c_rnn_gan'], 'Invalid model name.'
 
-df = pd.read_csv(os.path.join(FLAGS['raw_datadir'], 'maestro_updated.csv'))
-test_df = df[df['split'] == 'test']
-train_df = df[df['split'] == 'train']
-
-train_data_paths = [
-    os.path.join(FLAGS['raw_datadir'], i) for i in train_df['midi_filename']
-]
-test_data_paths = [
-    os.path.join(FLAGS['raw_datadir'], i) for i in test_df['midi_filename']
-]
+test_df = pd.read_csv('../data/maestro_test.csv')
+train_df = pd.read_csv('../data/maestro_train.csv')
+val_df = pd.read_csv('../data/maestro_validation.csv')
 
 utils.make_folder(FLAGS['model_folder'])
 
@@ -44,13 +36,15 @@ if FLAGS['model_name'] == 'baseline':
         print(f'[!] {utils.now()} Loading checkpoint')
         saved = torch.load(
             os.path.join(FLAGS['model_folder'],
-                         FLAGS['model_name'] + '_checkpoint'))
+                         FLAGS['model_name'] + '_checkpoint'),
+            map_location='cpu')
         for i in saved.keys():
             FLAGS[i] = saved[i]
-        stats = train_baseline_large(
+        losses, grads = train_baseline_large(
             VOCAB,
             RNG,
-            df['seq_len'].tolist(),
+            '../data/processed/train_songs.npy',
+            train_df['seq_len'].tolist(),
             in_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
@@ -64,16 +58,22 @@ if FLAGS['model_name'] == 'baseline':
             stride_size=FLAGS['stride_size'],
             use_bias=FLAGS['use_bias'],
[…]

### 2019-04-26 paper/main.tex
finished GAN and evaluation metrics

@@
 
-Over the past sixty years, Artificial Neural Networks (ANNs) have experienced a number of cycles in popularity noted by extensive publicity of over-inflated expectations on the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs [CITATION NEEDED]): that a system could learn to do anything it was programmed to do [CITATION NEEDED] \parencite{minsky_perceptrons_1988}.
+Over the past sixty years, Artificial Neural Networks (ANNs) have experienced a number of popularity cycles that were noted by extensive publicity of over-inflated expectations with regards to the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs [CITATION NEEDED]) [CITATION NEEDED].
+More specifically, it promised that a system could learn to do anything it was programmed to do \parencite{minsky_perceptrons_1988}.
 However, each wave of praise was followed by a mass of public disappointment, criticism, and funding cuts [CITATION NEEDED].
@@
 
-The recent revival of ANNs under the scope of Deep Learning (DL) has brought new ways to tackle Machine Learning problems, most notably in the context of classification (but also regression, although less popular; see \cite{lecun_deep_2015} for several examples of DL applications).
+The recent revival of ANNs under the scope of Deep Learning (DL) has brought new ways to tackle Machine Learning problems, most notably in the context of classification (but also regression, although a less popular application; see \cite{lecun_deep_2015} for several DL examples).
 In addition, with the advent of chatbots \parencite{dale_return_2016}, voice-activated personal assistants \parencite{xiong_microsoft_2018}, and general-purpose A.I.
@@
 
-% NOTE: USE ME LATER
-% Currently, two exciting areas of research in the field of Deep Learning (DL) are in designing generative models \parencite{goodfellow_nips_2016} and training NNs on small datasets \parencite{finn_model-agnostic_2017, nichol_first-order_2018}.
-This research will combine both of these tasks by exploring the extent to which generative models can create novel and qualitative samples of music under the constraints posed by the few-shot learning problem.
-In other words, when data is scarce, can a generative model create samples of music that are new (i.e., not in the training set), consistent with common scales, and diverse (tone sequences are not repeated over and over, combines chord structures with single-note sequences, is not restricted to - say - one octave).
-This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under the Reptile learning framework \parencite{nichol_first-order_2018} on small amounts of MIDI data.
-
 Broadly speaking, DL as a field, promises to solve tasks which are easily (intuitively) performed by people, but hard to formalize \parencite[e.g., recognizing faces or spoken word;][]{goodfellow_deep_2016}.
-The solution comprises of enabling computers to learn form experience and understand the world through the discovery of hierarchical relationships between concepts \parencite{lecun_deep_2015}.
+The solution comprises of enabling computers to learn from experience and understand the world through the discovery of hierarchical relationships between concepts \parencite{lecun_deep_2015}.
 This way, there is no need for manual input on the sort of knowledge that the computer needs \parencite{goodfellow_deep_2016}.
@@
 
-The reasons for presenting these facts will be clarified in the following sections however, a terse argumentation is that progress towards general A.I.
+The reasons for presenting these facts will be clarified in the following sections (Section \ref{sec:thesis_goal} provides a concrete explanation) however, a terse argumentation is that progress towards general A.I.
 systems is marked by scant explorations of ANNs outside of the classification task, and even more so when trying to overcome some of their shortcomings, such as being highly reliant on massive amounts of data [CITATION NEEDED].
@@
 
-For the purposes of this paper (see Section \ref{sec:thesis_goal}), generative model are central to our experiments (Chapter \ref{chap:methods}).
+For the purposes of this paper (detailed in Section \ref{sec:thesis_goal}), generative models are a central element in our experiments (Chapter \ref{chap:methods}).
 Borrowing the analogy from \textcite{goodfellow_nips_2016}, one might wonder what is the value of studying generative models.
 After all, in the image domain, such a framework would merely generate more images (something which the Internet as no shortage of).
-Some of the more eloquent answers include using generative models for simulating possible futures to be later used for planning [CITATION NEEDED], providing predictions to missing data, or enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.
+The answer to this lies in the fact that generative models can be used for simulating possible futures for scheduling and planning [CITATION NEEDED], providing predictions to missing data [CITATION NEEDED], or enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.
 
 However, successes in generative models have been scarce, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run.
-Yet, this method requires generating each sample one at a time (ramping up the cost of the process to $O(n)$).
+Yet, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$).
 Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}.
 Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task.
-Their main drawback is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.
+Their main drawback however, is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.
 
@@
 
-Although not as popular as the abovementioned, Recurrent Neural Networks (RNNs) have been successfully implemented for various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}.
-More specifically, a variation of these, Long Short-Term Memory Networks \parencite[LSTMs;][]{hochreiter_long_1997}, are of particular interest to this task.
+In addition, although not as popular as the abovementioned, Recurrent Neural Networks (RNNs) have been successfully implemented for various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}.
+More specifically, a variation of these, the Long Short-Term Memory Networks \parencite[LSTMs;][]{hochreiter_long_1997}, are of particular interest.
 RNNs (and by extension LSTMs) contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}.
-This makes them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc \parencite{fan_tts_2014}.
+This makes them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc.
+\parencite{fan_tts_2014}.
+That being said, the purpose-built memory cell of an LSTM permits better storing and accessing of information \parencite{hochreiter_long_1997, graves_generating_2013} which in turn makes this RNN variation a better candidate for this sort of task.
 
@@
 
-These developments in generative models lie in the field of Computer Vision, but that is not to say that we can generate only images.
-GANs (and others) have been applied - with varying degrees of success - in Natural Language Processing where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}.
+Having said all of these, the majority of developments with generative models lie in the field of Computer Vision.
+But that is not to say that we can generate only images.
+Generative models have been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}.
 Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}.
-This paper will focus on the problem of generating music with a NN-based model.
-Respectively, this paper will evaluate the samples generated by C-RNN-GAN \parencite{mogren_c-rnn-gan_2016} and Performance-RNN \parencite{oore_this_2018} when trained under the few-shot learning problem (detailed below).
+Herein lies the scope of this paper.
+We evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and LSTM \parencite{oore_this_2018} when trained under a meta-learning algorithm (see Section \ref{sec:hungry}).
+To reiterate, a full explanation of the purpose of the paper is provided in Section \ref{sec:thesis_goal}.
 
@@
 
-It is fairly well known that DL models tend to be very "data-hungry", a caveat imposed by its optimization algorithm \parencite{chen_closer_2018, ravi_optimization_2016}.
-DL usually relies on a procedure called stochastic gradient descent (SGC) which implies computing for a given input vector the output and error, followed by calculating the input's average gradient and consequently adjusting the network's weights \parencite{lecun_deep_2015}.
-\textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates: when faced with non-convex optimization problems, DL algorithms do not guarantee convergence within a reasonable amount of time.
-Finally, they add by saying that for each new problem (e.g., new dataset), the model needs to reinitialize with a new set of random parameters which hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}.
+DL models usually rely on a procedure called stochastic gradient descent (SGD) which implies computing an output and an error for a given input vector, followed by calculating the input's average gradient and adjusting the network's parameters in a manner that minimizes the error \parencite{lecun_deep_2015}.
+\textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates over few examples.
+In particular, when faced with non-convex optimization problems, DL algorithms do not guarantee convergence within a reasonable amount of time.
+In addition, they also emphasize that for each new problem (e.g., new dataset), the model needs to reinitialize with a new set of random parameters.
+Ultimately, this hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}.
+Thus they lay out two clear drawbacks of DL models: "data-hunger", and inability to perform separate tasks.
 
-Besides the promise of making DL algorithms easier to train, primary reasons why this issue needs to be addressed are as follows.
-Firstly, humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}, something which DL algorithms do not.
-Secondly, this would alleviate the constant need for data collection in order to provide reasonable performances.
-Finally, there are many fields where the data exhibits a large number of classes, with few examples per class and Bridging the gap between human-type learning and current learning architectures would allow models to properly capture this sort of data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}.
-This is known as the few-shot problem \parencite{chen_closer_2018, vinyals_matching_2016, larochelle_few-shot_2017}.
+A clear outcome from solving these issues would be that DL models are easier to train (less data) and multifunctional however, this list is not exhaustive.
+There is also fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack.
+Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class.
+Bridging the gap between human-type learning and current learning architectures would allow models to properly capture this sort of sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}.
+This has motivated a series of explorations in the direction of "few-shot" learning (i.e., learning a class from a few examples) or more broadly, into the field of "meta-learning" \parencite[learning to learn in hopes of generalizing to new tasks;][]{chen_closer_2018, vinyals_matching_2016, zhang_metagan_2018}.
 
-\section{Making music like a hooman} \label{sec:thesis_goal}
+In spite of its proposed benefits, developments and experiments with few-shot learning are still scarce \parencite{larochelle_few-shot_2017}.
+Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}.
+Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early explorers in this direction.
 
-In spite of its proposed benefits, developments and experiments in few-shot learning are still scarce \parencite{larochelle_few-shot_2017} and evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}.
-Few-shot experiments with generative networks are even less frequent \parencite{zhang_metagan_2018, clouatre_figr_2019}.
-This research aims to explore the viability of this learning approach in the context of generating music data.
-More concretely, in light of recent developments, this research will evaluate two generative networks modelled on MIDI data under the constraints of few-shot learning, and comparing its performance to the baseline proposed by \textcite{larochelle_few-shot_2017} and properties extracted from real music performances.
+\section{Composing like a human} \label{sec:thesis_goal}
 
-\chapter{Related Work}
+At the time of writing, generative models and meta-learning are two exciting areas of research in DL which have been successfully combined.
+However, to the knowledge of the researcher, no such experiments have been conducted in the musical domain.
+Thus, this paper examined the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning.
+In other words, it sought to evaluate the extent to which generative models can create novel and high-fidelity samples when training data is scarce.
+To avoid ambiguity, we define novel as new (i.e., not in the training set), and qualitative as consistent with common scales and diverse (combines chord structures with single-note sequences, is not restricted to only a few notes, etc.).
+This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, on a limited number of piano recordings.
+The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.
+
+In other words, this study sought to answer the following research questions:
+
+\begin{itemize}
+    \item To what extent is the music created by a few-shot generative model comparable to the music of a generative model that is trained on the entire dataset?
+    \item To what extent is the music created by a few-shot generative model comparable to real music?
+\end{itemize}
+
+The following chapters will discuss more about generative models, meta-learning, and some of the models and algorithms involved in this paper (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the setup of the experiments (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}).
+Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} draws some conclusions that answer the above research questions.
+
+\chapter{Related Work}\label{chap:theoretical_framework}
 
@@
 Probably more important than the network architecture, a central piece of the model's performance is due to the final representation of the MIDI data: each time slice is represented as a one-hot encoded vector containing Note-On (128 values), Note-Off (128 values), Time-Shift (125 values), and Velocity (32 values) events \parencite{oore_this_2018}.
-
-C-RNN-GAN and Performance-RNN have not been tested and adapted to the few-shot learning problem, but more general GANs and LSTM networks have been \parencite[see][]{zhang_metagan_2018, vinyals_matching_2016}.
-However, as mentioned previously, no evaluation has been made in the musical domain.
-Therefore, this study will seek to answer the following research questions:
-
-\begin{itemize}
-    \item To what extent is the music created by a few-shot generative model comparable to the music of a generative model that is trained on the entire dataset?
-    \item To what extent is the music created by a few-shot generative model comparable to real music?
-
-\end{itemize}
 
@@
 
-\chapter{Results}
+\chapter{Results}\label{chap:results}
 
-\chapter{Discussion}
+\chapter{Discussion}\label{chap:discussion}
 
-\chapter{Conclusion}
+\chapter{Conclusion}\label{chap:conclusion}
 

### 2019-04-26 src/c_rnn_gan.py

@@ -1,0 +1,237 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
+
+
+class Generator(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=415,
+                 dropout=0.3,
+                 num_layers=2,
+                 init_dim=100,
+                 use_bias=False):
+        super(Generator, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = int(out_dim)
+        self.dropout = dropout
+        self.init_dim = init_dim
+        self.use_bias = use_bias
+        self.num_layers = num_layers
+        self.is_bidirectional = False
+
+        if self.is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=self.use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=self.dropout,
+            bidirectional=False)
+
+        self.out_layer = nn.Linear(self.hidden_dim * self.num_layers,
+                                   self.out_dim)
+        self.out_activation = nn.Softmax(dim=-1)
+
+        self._init_weights()
+
+    def _init_weights(self):
+        nn.init.xavier_normal_(self.event_embedding.weight)
+
+        nn.init.xavier_normal_(self.init_layer.weight)
+        # self.init_layer.bias.fill_(0.0)
+
[…]

### 2019-04-26 src/data.py

@@ -191,14 +191,16 @@ def save_midi(midi, path: str, fname: str):
     midi.write(os.path.join(path, fname))
 
 
-def sample_mini_dataset(dataset: pd.DataFrame, num_classes: int,
-                        num_shots: int, rng):
+def sample_mini_dataset(dataset: pd.DataFrame,
+                        num_classes: int,
+                        num_shots: int,
+                        rng,
+                        labels='genre'):
     shuffled = dataset.sample(frac=1).reset_index(drop=True)
-    sampled_classes = rng.choice(
-        dataset['canonical_composer'].tolist(), size=num_classes)
+    sampled_classes = rng.choice(dataset[labels].tolist(), size=num_classes)
     samples = []
     for cls in sampled_classes:
-        filtered = shuffled[shuffled['canonical_composer'] == cls]
+        filtered = shuffled[shuffled[labels] == cls]
         for sample in filtered['midi_filename'].sample(
                 num_shots, random_state=rng):
             samples.append(sample)
@@ -209,6 +211,9 @@ def sample_mini_dataset(dataset: pd.DataFrame, num_classes: int,
 def mini_batches(samples: list, rng, batch_size: int, replacement=False):
     batches = []
     # samples = np.array(samples)
+    # Originally, there was one guy labelled as
+    #  expressionist, but that meant having a category
+    #  only for him. he was moved to modernism
     idx = np.arange(start=0, stop=len(samples))
     if replacement:
         smp = rng.choice(idx, size=batch_size, replace=replacement)
@@ -252,7 +257,7 @@ def batch(seq_lens, batch_size, window_size, stride_size, rng):
 
     if len(positions) < batch_size:
         repl = True
-        n_batches = batch_size // len(positions)
+        n_batches = 1
     else:
         repl = False
         n_batches = len(positions) // batch_size
@@ -469,6 +474,8 @@ def decode(piano_roll, fs):
 if __name__ == '__main__':
     from tqdm import tqdm
     df = pd.read_csv('../data/maestro-v1.0.0.csv')
+
+    # add genres to DataFrame
     genres = pd.read_csv('../data/composers_genre.csv')
     df['genre'] = [
         genres[genres['canonical_composer'] == i]['genre'].item()

### 2019-04-26 src/evaluation.py

@@ -1,6 +1,13 @@
+from scipy.stats import norm
 from tqdm import tqdm
-from data import midi2piano
+from data import midi2piano, to_seq, build_vocab
+from sklearn.cluster import KMeans
+from sklearn.metrics import pairwise_distances_argmin
+from collections import Counter
+from math import sqrt
 import mido
+import numpy as np
+import pandas as pd
 
 
 def open_midi(path):
@@ -168,3 +175,99 @@ def evaluate_batch(paths, poly_criterion=31.35):
     #     results[key] = sum(results[key]) / len(results[key])
 
     return results
+
+
+def _prepare_data(paths, max_len=None):
+    vocab = build_vocab()
+    vocab = {i: vocab.index(i) for i in vocab}
+    data = []
+
+    for path in tqdm(paths, desc='Building sequences'):
+        dat = to_seq(path)
+        dat = [vocab[i] for i in dat]
+        data.append(np.array(dat))
+
+    length = [len(i) for i in data]
+    if max_len is None:
+        max_len = max(length)
+
+    sequences = []
+    for i in tqdm(data, desc='Padding data'):
+        diff = max_len - i.shape[0]
+        if diff > 0:
+            i = np.pad(i, pad_width=((0, diff)), mode='constant')
+        elif diff < 0:
+            i = i[:max_len]
+
+        sequences.append(i)
+
+    return np.vstack(sequences), max_len
+
+
+def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
+    def assign_counts_to_bins(bins, count):
+        for key in count.keys():
+            bins[key] = count[key]
+        return bins
+
+    def bin_se(key):
+        a = pool_prop[key] * (1 - pool_prop[key])
+        b = (1 / real_data.shape[0]) + (1 / gen_data.shape[0])
+        return sqrt((a * b))
+
+    if rng is None:
+        rng = np.random.RandomState(1337)
+
+    n_samples = real_data.shape[0] + gen_data.shape[0]
+
+    # assign test data into bins
+    clf = KMeans(n_clusters=n_bins, random_state=rng, n_jobs=workers)
+    clf.fit(real_data)
+    r_bins = clf.labels_
+
+    # count how many in bins
+    r_count = {i: 0 for i in range(n_bins)}
+    r_count = assign_counts_to_bins(r_count, Counter(r_bins))
+
+    # assign generated data to closest (L2) centroid
[…]

### 2019-04-26 src/old/c_rnn_gan.py

@@ -1,0 +1,240 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
+
+
+class Generator(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=415,
+                 dropout=0.3,
+                 num_layers=2,
+                 init_dim=100,
+                 use_bias=False):
+        super(Generator, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = int(out_dim)
+        self.dropout = dropout
+        self.init_dim = init_dim
+        self.use_bias = use_bias
+        self.num_layers = num_layers
+        self.is_bidirectional = False
+
+        if self.is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=self.use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=self.dropout,
+            bidirectional=False)
+
+        self.out_layer = nn.Linear(self.hidden_dim * self.num_layers,
+                                   self.out_dim)
+        self.out_activation = nn.Softmax(dim=-1)
+
+        self._init_weights()
+
+    def _init_weights(self):
+        nn.init.xavier_normal_(self.event_embedding.weight)
+
+        nn.init.xavier_normal_(self.init_layer.weight)
+        # self.init_layer.bias.fill_(0.0)
+
[…]

### 2019-04-26 src/old/performance_rnn.py

@@ -1,0 +1,767 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
+
+
+class PerformanceRNN(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=415,
+                 dropout=0.0,
+                 num_layers=1,
+                 init_dim=32,
+                 use_bias=False,
+                 is_bidirectional=False):
+        super(PerformanceRNN, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = out_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+
+        if is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=float(dropout),
+            bidirectional=is_bidirectional)
+
+        self.out_layer = nn.Linear(self.hidden_dim * self.num_layers,
+                                   self.out_dim)
+        self.out_activation = nn.Softmax(dim=-1)
+
+        self._init_weights()
+
+    def _init_weights(self):
+        nn.init.xavier_normal_(self.event_embedding.weight)
+
+        nn.init.xavier_normal_(self.init_layer.weight)
[…]

### 2019-04-26 src/old/train_c_rnn_gan.py

@@ -1,0 +1,404 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+# from lstm_baseline import Baseline
+from c_rnn_gan import Generator, Discriminator
+
+import pdb
+
+
+class ReptileGAN:
+    def __init__(self,
+                 vocab,
+                 rng,
+                 raw_datadir,
+                 split,
+                 in_dim=415,
+                 hidden_dim=512,
+                 init_dim=100,
+                 num_layers=2,
+                 dropout=0.3,
+                 n_classes=5,
+                 n_shots=1,
+                 learning_rate=0.025,
+                 batch_size=64,
+                 inner_iters=5,
+                 meta_step=0.025,
+                 meta_batch=1,
+                 meta_iters=250,
+                 meta_iters_start=0,
+                 use_bias=False,
+                 window_size=200,
+                 stride_size=200,
+                 clip_grad=True,
+                 clip_norm=3,
+                 max_norm=1.0,
+                 g_train=True,
+                 d_train=True,
+                 save_path='../models/',
+                 g_model_state=None,
+                 d_model_state=None,
+                 g_optimizer_state=None,
+                 d_optimizer_state=None,
+                 g_loss_saved=None,
+                 d_loss_saved=None,
+                 g_grad_norm=None,
+                 d_grad_norm=None):
+
+        self.vocab = {i: vocab.index(i) for i in vocab}  # list to dict
+        self.rng = rng
+        self.raw_datadir = raw_datadir
+        self.split = split
+        self.in_dim = in_dim
+        self.hidden_dim = hidden_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+        self.dropout = dropout
+        self.n_classes = n_classes
+        self.n_shots = n_shots
+        self.learning_rate = learning_rate
+        self.batch_size = batch_size
+        self.inner_iters = inner_iters
+        self.meta_step = meta_step
+        self.meta_batch = meta_batch
+        self.meta_iters = meta_iters
+        self.meta_iters_start = meta_iters_start
+        self.use_bias = use_bias
+        self.window_size = window_size
+        self.stride_size = stride_size
+        self.clip_grad = clip_grad
[…]

### 2019-04-26 src/performance_rnn.py

@@ -325,6 +325,12 @@ def reptilian(vocab,
         dropout=dropout,
         out_dim=len(vocab))
 
+    print(model.batch_size)
+    print(model.init_dim)
+    print(model.hidden_dim)
+    print(model.num_layers)
+    print(model.device)
+
     if model_state is not None:
         model.load_state_dict(model_state)
 
@@ -413,7 +419,7 @@ def reptilian(vocab,
                                                     clip_norm)
 
                     norm_epoch.append(norm.cpu().numpy())
-                    pdb.set_trace()
+
                     if clip_grad:
                         nn.utils.clip_grad_norm_(model.parameters(), max_norm,
                                                  clip_norm)
@@ -456,7 +462,7 @@ def reptilian(vocab,
         print(f'[!] {utils.now()} Current Loss: {loss_saved[-1][1]}')
 
         # Save model every 10 epochs
-        if epoch % 10 == 0:
+        if epoch % 10 == 0 and epoch != 0:
             print(f'[!] {utils.now()} Saving checkpoint at epoch {epoch+1}')
             checkpoint = os.path.join(save_path, 'performance_rnn_checkpoint')
             torch.save(
@@ -525,3 +531,237 @@ def reptilian(vocab,
         },
         os.path.join(save_path, f'performance_rnn_{n_shots}_{n_classes}'))
     return loss_saved
+
+
+def figr(vocab,
+         rng,
+         raw_datadir,
+         split,
+         in_dim=415,
+         hidden_dim=512,
+         init_dim=32,
+         num_layers=3,
+         dropout=0.0,
+         n_classes=5,
+         n_shots=1,
+         learning_rate=0.025,
+         batch_size=64,
+         inner_iters=5,
+         meta_step=0.025,
+         meta_step_final=0.0,
+         meta_batch=1,
+         meta_iters=250,
+         meta_iters_start=0,
+         use_bias=False,
+         is_bidirectional=False,
+         window_size=200,
+         stride_size=200,
+         clip_grad=True,
+         clip_norm=3,
+         max_norm=1.0,
+         save_path='../models/',
+         model_state=None,
+         optimizer_state=None,
+         loss_saved=None,
+         grad_norm=None):
+
+    df = pd.read_csv(os.path.join(raw_datadir, f'maestro_{split}.csv'))
+
+    # Use GPU if we have one
+    device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
[…]

### 2019-04-26 src/run.py

@@ -96,7 +96,7 @@ if FLAGS['model_name'] == 'baseline':
                      FLAGS['model_name'] + '_results.csv'))
 
 elif FLAGS['model_name'] == 'performance_rnn':
-    from performance_rnn import reptilian
+    from train_performance_rnn import ReptilePRNN
 
     checkpoint = Path(
         os.path.join(FLAGS['model_folder'],
@@ -112,7 +112,7 @@ elif FLAGS['model_name'] == 'performance_rnn':
         for i in saved.keys():
             FLAGS[i] = saved[i]
 
-        losses, grads = reptilian(
+        reptile = ReptilePRNN(
             VOCAB,
             RNG,
             FLAGS['raw_datadir'],
@@ -128,7 +128,6 @@ elif FLAGS['model_name'] == 'performance_rnn':
             batch_size=FLAGS['batch_size'],
             inner_iters=FLAGS['inner_iters'],
             meta_step=FLAGS['meta_step'],
-            meta_step_final=FLAGS['meta_step_final'],
             meta_batch=FLAGS['meta_batch'],
             meta_iters=FLAGS['meta_iters'],
             meta_iters_start=FLAGS['meta_iters_start'],
@@ -147,7 +146,7 @@ elif FLAGS['model_name'] == 'performance_rnn':
 
     else:
         print(f'[!] {utils.now()} Starting model...')
-        losses, grads = reptilian(
+        reptile = ReptilePRNN(
             VOCAB,
             RNG,
             FLAGS['raw_datadir'],
@@ -163,7 +162,6 @@ elif FLAGS['model_name'] == 'performance_rnn':
             batch_size=FLAGS['batch_size'],
             inner_iters=FLAGS['inner_iters'],
             meta_step=FLAGS['meta_step'],
-            meta_step_final=FLAGS['meta_step_final'],
             meta_batch=FLAGS['meta_batch'],
             meta_iters=FLAGS['meta_iters'],
             use_bias=FLAGS['use_bias'],
@@ -175,7 +173,90 @@ elif FLAGS['model_name'] == 'performance_rnn':
             max_norm=FLAGS['max_norm'],
             save_path=FLAGS['model_folder'])
 
-    results = pd.DataFrame({'loss': losses, 'grad': grads})
-    results.to_csv(
+    reptile.train()
+
+elif FLAGS['model_name'] == 'c_rnn_gan':
+    from train_c_rnn_gan import ReptileGAN
+
+    checkpoint = Path(
[…]

### 2019-04-26 src/train_c_rnn_gan.py

@@ -1,0 +1,435 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+# from lstm_baseline import Baseline
+from c_rnn_gan import Generator, Discriminator
+
+import pdb
+
+
+class ReptileGAN:
+    def __init__(self,
+                 vocab,
+                 rng,
+                 raw_datadir,
+                 split,
+                 in_dim=415,
+                 hidden_dim=512,
+                 init_dim=100,
+                 num_layers=2,
+                 dropout=0.3,
+                 n_classes=5,
+                 n_shots=1,
+                 learning_rate=0.025,
+                 batch_size=64,
+                 inner_iters=5,
+                 meta_step=0.025,
+                 meta_batch=1,
+                 meta_iters=250,
+                 meta_iters_start=0,
+                 use_bias=False,
+                 window_size=200,
+                 stride_size=200,
+                 clip_grad=True,
+                 clip_norm=3,
+                 max_norm=1.0,
+                 g_train=True,
+                 d_train=True,
+                 feature_matching=False,
+                 save_path='../models/',
+                 g_model_state=None,
+                 d_model_state=None,
+                 g_optimizer_state=None,
+                 d_optimizer_state=None,
+                 g_loss_saved=None,
+                 d_loss_saved=None,
+                 g_grad_norm=None,
+                 d_grad_norm=None):
+
+        self.vocab = {i: vocab.index(i) for i in vocab}  # list to dict
+        self.rng = rng
+        self.raw_datadir = raw_datadir
+        self.split = split
+        self.in_dim = in_dim
+        self.hidden_dim = hidden_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+        self.dropout = dropout
+        self.n_classes = n_classes
+        self.n_shots = n_shots
+        self.learning_rate = learning_rate
+        self.batch_size = batch_size
+        self.inner_iters = inner_iters
+        self.meta_step = meta_step
+        self.meta_batch = meta_batch
+        self.meta_iters = meta_iters
+        self.meta_iters_start = meta_iters_start
+        self.use_bias = use_bias
+        self.window_size = window_size
+        self.stride_size = stride_size
+        self.clip_grad = clip_grad
[…]

### 2019-04-26 src/train_performance_rnn.py

@@ -1,0 +1,252 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+# from lstm_baseline import Baseline
+from performance_rnn import PerformanceRNN
+
+import pdb
+
+
+class ReptilePRNN:
+    def __init__(self,
+                 vocab,
+                 rng,
+                 raw_datadir,
+                 split,
+                 in_dim=415,
+                 hidden_dim=512,
+                 init_dim=32,
+                 num_layers=3,
+                 dropout=0.0,
+                 n_classes=5,
+                 n_shots=1,
+                 learning_rate=0.025,
+                 batch_size=64,
+                 inner_iters=5,
+                 meta_step=0.025,
+                 meta_batch=1,
+                 meta_iters=250,
+                 meta_iters_start=0,
+                 use_bias=False,
+                 is_bidirectional=False,
+                 window_size=200,
+                 stride_size=200,
+                 clip_grad=True,
+                 clip_norm=3,
+                 max_norm=1.0,
+                 save_path='../models/',
+                 model_state=None,
+                 optimizer_state=None,
+                 loss_saved=None,
+                 grad_norm=None):
+
+        self.vocab = {i: vocab.index(i) for i in vocab}  # list to dict
+        self.rng = rng
+        self.raw_datadir = raw_datadir
+        self.split = split
+        self.in_dim = in_dim
+        self.hidden_dim = hidden_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+        self.dropout = dropout
+        self.n_classes = n_classes
+        self.n_shots = n_shots
+        self.learning_rate = learning_rate
+        self.batch_size = batch_size
+        self.inner_iters = inner_iters
+        self.meta_step = meta_step
+        self.meta_batch = meta_batch
+        self.meta_iters = meta_iters
+        self.meta_iters_start = meta_iters_start
+        self.use_bias = use_bias
+        self.is_bidirectional = is_bidirectional
+        self.window_size = window_size
+        self.stride_size = stride_size
+        self.clip_grad = clip_grad
+        self.clip_norm = clip_norm
+        self.max_norm = max_norm
+        self.save_path = save_path
+        self.loss_saved = loss_saved
+        self.grad_norm = grad_norm
+
+        self.device = torch.device('cuda:0' if torch.cuda.
[…]

### 2019-04-26 src/c_rnn_gan.py
added code to generate and CC message support

@@ -20,7 +20,7 @@ class Generator(nn.Module):
                  in_dim,
                  hidden_dim,
                  batch_size,
-                 out_dim=415,
+                 out_dim=416,
                  dropout=0.3,
                  num_layers=2,
                  init_dim=100,

### 2019-04-26 src/data.py

@@ -7,11 +7,13 @@ import pretty_midi
 import mido
 import utils
 
+from mido import MetaMessage, Message, MidiTrack, MidiFile
+
 
 def to_seq(path):
     midi = mido.MidiFile(path)
     tpb = midi.ticks_per_beat  # Ticks/beat. between 384-480
-    tempo = midi.tracks[0][0].tempo  # Tempo in ms. All are 500000
+    tempo = midi.tracks[0][0].tempo  # Tempo in microseconds. All are 500000
     # max_tick = mido.second2tick(second=1, ticks_per_beat=tpb, tempo=tempo)
 
     melody = midi.tracks[1]
@@ -22,8 +24,8 @@ def to_seq(path):
             # move_by = int(round(msg.time/8, 0)*8)
             # Convert from ticks to miliseconds
             move_by = int(mido.tick2second(msg.time, tpb, tempo) * 1000)
-            # SORT OF quantizing by 8ms
-            move_by = int(round(move_by / 8, 0) * 8)
+            # sort of quantizing by 8ms
+            move_by = int((move_by // 8) * 8)
 
             # If the time between the previous note and
             #  the current note is greater than 1 second,
@@ -36,12 +38,126 @@ def to_seq(path):
             sequence.append('time_move:{}'.format(move_by))
             # Append velocity; we limit the number of possible
             #  velocity values to 32 instead of 128
-            #  When decoding velocity must be multiplied by 32
+            #  When decoding velocity must be multiplied by 4
             #  get back the appropriate values
-            sequence.append('velocity_set:{}'.format(msg.velocity // 32))
+            sequence.append('velocity_set:{}'.format(msg.velocity // 4))
+
             # Append `note_on` or `note_off`
-            sequence.append(msg.type + ':{}'.format(msg.note))
+            # The data has this weird quirk where it
+            #  does not make use of the note_off
+            #  messsage. What it does, is that it sets
+            #  the velocity of the note to be closed
+            #  to zero. Thus, a messages with 'note_on'
+            #  note=60, velocity=0, actually means
+            #  'note_off', note=60, velocity=0
+            # Thus, we store the message type in a
+            #  variable, and check the velocity. If
+            #  it's zero, then we change the message
+            #  to note_off
+            message_type = msg.type
+            if msg.velocity == 0:
+                message_type = 'note_off'
+            sequence.append(message_type + ':{}'.format(msg.note))
+        elif msg.type == 'control_change':
[…]

### 2019-04-26 src/generate.py

@@ -1,0 +1,55 @@
+import torch
+import os
+import pandas as pd
+import numpy as np
+from tqdm import tqdm
+from data import from_seq
+from utils import make_folder
+
+
+def load_model_data(path):
+    return torch.load(path, map_location='cpu')
+
+
+if __name__ == '__main__':
+    # Use GPU if we have one
+    device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
+    RNG = np.random.RandomState(1337)
+    df = pd.read_csv('../data/maestro_train.csv')
+    median_len = int(df['seq_len'].median())
+
+    greed = False
+    temperature = 1.1
+
+    models = ['baseline']
+
+    for model in models:
+        checkpoint = load_model_data(os.path.join('../models/', model))
+        if model == 'baseline':
+            from lstm_baseline import Baseline
+            net = Baseline(
+                in_dim=checkpoint['in_dim'],
+                hidden_dim=checkpoint['hidden_dim'],
+                batch_size=1,  # we create one song at a time
+                out_dim=checkpoint['out_dim'],
+                dropout=checkpoint['dropout'],
+                num_layers=checkpoint['num_layers'],
+                init_dim=checkpoint['init_dim'],
+                use_bias=checkpoint['use_bias'],
+                is_bidirectional=checkpoint['is_bidirectional'])
+            net.load_state_dict(checkpoint['model_state'])
+        for i in tqdm(range(1, 101), desc=f'Generating from {model}'):
+            init = torch.randn(net.batch_size, net.init_dim).to(device)
+            steps = median_len * (RNG.randint(90, 111) / 100)
+            song = net.generate(
+                init,
+                steps,
+                greedy=greed,
+                temperature=temperature,
+                output_type='index')
+            song = song.tolist()
+            make_folder('../generated')
+            from_seq(
+                song,
+                path=os.path.join('../generated/',
+                                  f'{model}_{i}_{greed}_{temperature}.midi'))

### 2019-04-26 src/lstm_baseline.py

@@ -19,7 +19,7 @@ class Baseline(nn.Module):
                  in_dim,
                  hidden_dim,
                  batch_size,
-                 out_dim=415,
+                 out_dim=416,
                  dropout=0.0,
                  num_layers=1,
                  init_dim=32,
@@ -176,7 +176,8 @@ def train_baseline_large(vocab,
                          rng,
                          seq_file,
                          seq_lens,
-                         in_dim=415,
+                         in_dim=416,
+                         out_dim=416,
                          hidden_dim=256,
                          init_dim=32,
                          num_layers=1,
@@ -276,6 +277,7 @@ def train_baseline_large(vocab,
             torch.save(
                 {
                     'in_dim': in_dim,
+                    'out_dim': out_dim,
                     'hidden_dim': hidden_dim,
                     'init_dim': init_dim,
                     'num_layers': num_layers,
@@ -299,6 +301,7 @@ def train_baseline_large(vocab,
     torch.save(
         {
             'in_dim': in_dim,
+            'out_dim': out_dim,
             'hidden_dim': hidden_dim,
             'init_dim': init_dim,
             'num_layers': num_layers,

### 2019-04-26 src/performance_rnn.py

@@ -20,7 +20,7 @@ class PerformanceRNN(nn.Module):
                  in_dim,
                  hidden_dim,
                  batch_size,
-                 out_dim=415,
+                 out_dim=416,
                  dropout=0.0,
                  num_layers=1,
                  init_dim=32,

### 2019-04-26 src/run.py

@@ -46,6 +46,7 @@ if FLAGS['model_name'] == 'baseline':
             '../data/processed/train_songs.npy',
             train_df['seq_len'].tolist(),
             in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
             num_layers=FLAGS['num_layers'],
@@ -75,6 +76,7 @@ if FLAGS['model_name'] == 'baseline':
             '../data/processed/train_songs.npy',
             train_df['seq_len'].tolist(),
             in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
             num_layers=FLAGS['num_layers'],
@@ -118,6 +120,7 @@ elif FLAGS['model_name'] == 'performance_rnn':
             FLAGS['raw_datadir'],
             FLAGS['split'],
             in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
             num_layers=FLAGS['num_layers'],
@@ -152,6 +155,7 @@ elif FLAGS['model_name'] == 'performance_rnn':
             FLAGS['raw_datadir'],
             FLAGS['split'],
             in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
             num_layers=FLAGS['num_layers'],
@@ -198,6 +202,7 @@ elif FLAGS['model_name'] == 'c_rnn_gan':
             FLAGS['raw_datadir'],
             FLAGS['split'],
             in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
             num_layers=FLAGS['num_layers'],
@@ -236,6 +241,7 @@ elif FLAGS['model_name'] == 'c_rnn_gan':
             FLAGS['raw_datadir'],
             FLAGS['split'],
             in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
             hidden_dim=FLAGS['hidden_dim'],
             init_dim=FLAGS['init_dim'],
             num_layers=FLAGS['num_layers'],

### 2019-04-26 src/train_c_rnn_gan.py

@@ -24,7 +24,8 @@ class ReptileGAN:
                  rng,
                  raw_datadir,
                  split,
-                 in_dim=415,
+                 in_dim=416,
+                 out_dim=416,
                  hidden_dim=512,
                  init_dim=100,
                  num_layers=2,
@@ -62,6 +63,7 @@ class ReptileGAN:
         self.raw_datadir = raw_datadir
         self.split = split
         self.in_dim = in_dim
+        self.out_dim = out_dim
         self.hidden_dim = hidden_dim
         self.init_dim = init_dim
         self.num_layers = num_layers
@@ -123,6 +125,7 @@ class ReptileGAN:
     def init_models(self):
         self.g = Generator(
             in_dim=self.in_dim,
+            out_dim=self.out_dim,
             hidden_dim=self.hidden_dim,
             batch_size=self.batch_size,
             dropout=self.dropout,
@@ -131,6 +134,7 @@ class ReptileGAN:
             use_bias=self.use_bias).to(self.device)
         self.meta_g = Generator(
             in_dim=self.in_dim,
+            out_dim=self.out_dim,
             hidden_dim=self.hidden_dim,
             batch_size=self.batch_size,
             dropout=self.dropout,
@@ -194,6 +198,7 @@ class ReptileGAN:
             {
                 'split': self.split,
                 'in_dim': self.in_dim,
+                'out_dim': self.out_dim,
                 'hidden_dim': self.hidden_dim,
                 'init_dim': self.init_dim,
                 'num_layers': self.num_layers,

### 2019-04-26 src/train_performance_rnn.py

@@ -23,7 +23,8 @@ class ReptilePRNN:
                  rng,
                  raw_datadir,
                  split,
-                 in_dim=415,
+                 in_dim=416,
+                 out_dim=416,
                  hidden_dim=512,
                  init_dim=32,
                  num_layers=3,
@@ -55,6 +56,7 @@ class ReptilePRNN:
         self.raw_datadir = raw_datadir
         self.split = split
         self.in_dim = in_dim
+        self.out_dim = out_dim
         self.hidden_dim = hidden_dim
         self.init_dim = init_dim
         self.num_layers = num_layers
@@ -145,6 +147,7 @@ class ReptilePRNN:
             {
                 'split': self.split,
                 'in_dim': self.in_dim,
+                'out_dim': self.out_dim,
                 'hidden_dim': self.hidden_dim,
                 'init_dim': self.init_dim,
                 'num_layers': self.num_layers,

### 2019-04-29 paper/main.tex
should-be-mandatory backup

@@
 \newcommand{\hsp}{\hspace{10pt}}
-\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{base02}{\thechapter \hsp \vline}\hsp}{0pt}{\Huge\bfseries}
-\titleformat{\section}[hang]{\Large\bfseries}{\textcolor{base01}{\thesection \hsp \vline}\hsp}{0pt}{\Large\bfseries}
-\titleformat{\subsection}[hang]{\large\bfseries}{\textcolor{base01}{\thesubsection \hsp \vline}\hsp}{0pt}{\large\bfseries}
+\titleformat{\chapter}[hang]{\Huge\bfseries}{\textcolor{base02}{\thechapter }\hsp}{0pt}{\Huge\bfseries}
+\titleformat{\section}[hang]{\Large\bfseries}{\textcolor{base01}{\thesection}\hsp}{0pt}{\Large\bfseries}
+\titleformat{\subsection}[hang]{\large\bfseries}{\textcolor{base01}{\thesubsection}\hsp}{0pt}{\large\bfseries}
 
@@
 \usepackage{bm}
+
+% Drawings
+\usepackage{tikz}
+\usetikzlibrary{arrows.meta, fit, positioning}
+
+\tikzset{
+    neuron/.style={
+        circle, draw, thick,
+        inner sep=0pt,
+        minimum size=3.0em,
+        node distance=4em
+    },
+    io/.style={
+        neuron,
+        fill=base01
+    },
+    conn/.style={
+        thick,
+        -{Straight Barb[angle=60:2pt 3]},
+    }
+}
 
@@
 
-\section{"Just... One more bite"}\label{sec:hungry}
+\section{We always want more}\label{sec:hungry}
 
@@
 
+This chapter takes a deep-dive into the fields of meta-learning and generative networks in order to highlight their current state of affairs.
+This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}.
+Section \ref{sec:lstm} will present the principal network architecture of this paper, the Long Short-Term Memory network, and argument for its use as the foundation of a generative network.
+Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning, several approaches to it, and finally narrow down to current state-of-the-art algorithms.
+
+\section{Long Short-Term Memory Networks} \label{sec:lstm}
+
+In a standard ANN, the data traverses the network through the input layer, to the hidden layer, and finally reaching the output layer; setup which is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}.
+Broadly, this type of network is defined by the fact that its connections do not form cycles.
+If however we relax this condition, we arrive at a recurrent neural network \parencite[Figure \ref{fig:rnn_folded};][]{rumelhart_learning_1986}: a family of neural networks specifically designed for processing sequential data \parencite{goodfellow_deep_2016}.
+
+Concretely, information passes through the RNN similarly to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and prior hidden activations from the previous timestep \parencite{graves_supervised_2012}.
+Thus a formal definition of the network would be represented by
+
+\begin{equation}
+    h^t = f(h^{t-1}; x^t; \theta)
+\end{equation}
+
+where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the bias.
+
+\begin{figure}[t]
+    \centering
+    \begin{tikzpicture}
+        \node[io] (x1) {$x_1$};
+        \node[io, above=4em of x1] (x2) {$x_2$};
+        \node[io, above=4em of x2] (x3) {$x_3$};
+
+        \node[neuron, right= of x1] (h1) {$h_1$};
+        \node[neuron, above=4em of h1] (h2) {$h_2$};
+        \node[neuron, above=4em of h2] (h3) {$h_3$};
+
+        \node[io, right= of h2] (y) {$y$};
+
+        \foreach \x in {1, 2, 3}
+            \foreach \h in {1, 2, 3}
+                \draw[conn] (x\x) -- (h\h);
+
+        \foreach \h in {1, 2, 3}
+            \draw[conn] (h\h) -- (y);
+    \end{tikzpicture}
+    \caption{Example of a feedforward neural network with three nodes in the input layer ($x_1$, $x_2$, $x_3$), three nodes in the hidden layer ($h_1$, $h_2$, $h_3$), and an output layer a single node ($y$).}
+    \label{fig:feedforward_net}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \begin{subfigure}{.24\linewidth}
+        \begin{tikzpicture}
+            [align=center]
+            % Draw recurrent network
+            \node[neuron] (h) {$h$};
+            \node[io, below=4em of h] (x) {$x$};
+            \node[io, above=4em of h] (y) {$y$};
+
+            \draw[conn] (h) -- node [left=0.5em, midway, align=center] {$V$} (y);
+            \draw[conn] (x) -- node [left=0.5em, midway, align=center] {$U$} (h);
+            \draw[conn] (h.north) to [out=150, in=180, loop, looseness=4] node[left=0.5em, midway, align=center] {$W$} (h.west);
+
+        \end{tikzpicture}
+        \caption{Folded}
+        \label{fig:rnn_folded}
+    \end{subfigure}
+    \begin{subfigure}{.74\linewidth}
+        \begin{tikzpicture}
+            [align=center]
+            % Draw middle unfold
+            \node[neuron] (ht) {$h^t$};
+            \node[io, below=4em of ht] (xt) {$x^t$};
+            \node[io, above=4em of ht] (yt) {$y^t$};
+
+            \draw[conn] (ht) -- node [left=0.5em, midway, align=center] {$V$} (yt);
+            \draw[conn] (xt) -- node [left=0.5em, midway, align=center] {$U$} (ht);
+
+            % Draw left unfold
+            \node[neuron, left = of ht] (ht-1) {$h^{t-1}$};
+            \node[io, below=4em of ht-1] (xt-1) {$x^{t-1}$};
+            \node[io, above=4em of ht-1] (yt-1) {$y^{t-1}$};
+
+            \node[left= of ht-1] (nullLeft) {};
+
+            \draw[conn] (ht-1) -- node [left=0.5em, midway, align=center] {$V$} (yt-1);
+            \draw[conn] (xt-1) -- node [left=0.5em, midway, align=center] {$U$} (ht-1);
+
+            \draw[conn] (ht-1) -- node [above=0.5em, midway, align=center] {$W$} (ht);
+            \draw[conn] (nullLeft) -- node [above=0.5em, midway, align=center] {$W$} (ht-1);
+            % Draw right unfold
+            \node[neuron, right= of ht] (ht+1) {$h^{t+1}$};
+            \node[io, below=4em of ht+1] (xt+1) {$x^{t+1}$};
+            \node[io, above=4em of ht+1] (yt+1) {$y^{t+1}$};
+
+            \node[right= of ht+1] (nullRight) {};
+
+            \draw[conn] (ht+1) -- node [left=0.5em, midway, align=center] {$V$} (yt+1);
+            \draw[conn] (xt+1) -- node [left=0.5em, midway, align=center] {$U$} (ht+1);
+
+            \draw[conn] (ht) -- node [above=0.5em, midway, align=center] {$W$} (ht+1);
+            \draw[conn] (ht+1) -- node [above=0.5em, midway, align=center] {$W$} (nullRight);
+
+        \end{tikzpicture}
+        \caption{Unfolded}
+        \label{fig:rnn_unfolded}
+    \end{subfigure}
+\end{figure}
+
+\section{Generative Adversarial Networks} \label{sec:generative}
+
+Turning over to the generative side of things, nowadays, the most popular form of generating data is through the use of a Generative Adversarial Network \parencite[GAN;][]{goodfellow_generative_2014}.
+Here a generative model $G$ captures the data distribution and creates samples from it, whereas a discriminative model $D$ estimates the probability that the input is real (i.e., comes from the training data) rather than fake (i.e., comes from $G$).
+Formally, given $x$ sampled from real data, and $z$ being input noise, $D$ tries to keep $D(x)$ near $1$, while making $D(G(z))$ near $0$.
+Conversely, $G$ tries to make $D(G(z))$ near $1$ \parencite{goodfellow_nips_2016}.
+A notable implementation is C-RNN-GAN \parencite{mogren_c-rnn-gan_2016}, a continuous recurrent neural network with adversarial training.
+It consists of a network with two LSTM layers and 350 hidden units per layer, in both $G$ and $D$.
+However, $D$ has a bidirectional layout which allows is to take context from both past and future for its decisions, while $G$ is unidirectional \parencite{mogren_c-rnn-gan_2016}.
+As is the standard practice for GANs, $G$ will create fake samples of music, whereas $D$ will try to classify correctly which sample is real.
+
+Notwithstanding GANs, other generative successes have made use of adaptation on the Long Short-Term Memory (LSTM) network \parencite{hochreiter_long_1997}.
+Here, the cells of a network are designed to allow for information to persist by implementing a recursive state unit, an input gate, a forget gate, and an output gate.
+The input of the network is allowed into the state unit if it is allowed to pass through by the input gate.
+Next, the forget gate controls the self-loop of the state cell.
+Finally, the output gate has the power to shut off the output cell \parencite{goodfellow_deep_2016}.
+Here, a notable implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 layers each.
+Compared to C-RNN-GAN, here the network is completely unidirectional.
+Probably more important than the network architecture, a central piece of the model's performance is due to the final representation of the MIDI data: each time slice is represented as a one-hot encoded vector containing Note-On (128 values), Note-Off (128 values), Time-Shift (125 values), and Velocity (32 values) events \parencite{oore_this_2018}.
+
+\section{Learning to learn} \label{sec:learn2learn}
+
 \textcite{lake_human-level_2015} initially popularized the idea of learning from a few class examples using a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
@@
 This forms a compelling reason to use Reptile as the few-shot learning method for this research.
-
-Turning over to the generative side of things, nowadays, the most popular form of generating data is through the use of a Generative Adversarial Network \parencite[GAN;][]{goodfellow_generative_2014}.
-Here a generative model $G$ captures the data distribution and creates samples from it, whereas a discriminative model $D$ estimates the probability that the input is real (i.e., comes from the training data) rather than fake (i.e., comes from $G$).
-Formally, given $x$ sampled from real data, and $z$ being input noise, $D$ tries to keep $D(x)$ near $1$, while making $D(G(z))$ near $0$.
-Conversely, $G$ tries to make $D(G(z))$ near $1$ \parencite{goodfellow_nips_2016}.
-A notable implementation is C-RNN-GAN \parencite{mogren_c-rnn-gan_2016}, a continuous recurrent neural network with adversarial training.
-It consists of a network with two LSTM layers and 350 hidden units per layer, in both $G$ and $D$.
-However, $D$ has a bidirectional layout which allows is to take context from both past and future for its decisions, while $G$ is unidirectional \parencite{mogren_c-rnn-gan_2016}.
-As is the standard practice for GANs, $G$ will create fake samples of music, whereas $D$ will try to classify correctly which sample is real.
-
-Notwithstanding GANs, other generative successes have made use of adaptation on the Long Short-Term Memory (LSTM) network \parencite{hochreiter_long_1997}.
-Here, the cells of a network are designed to allow for information to persist by implementing a recursive state unit, an input gate, a forget gate, and an output gate.
-The input of the network is allowed into the state unit if it is allowed to pass through by the input gate.
-Next, the forget gate controls the self-loop of the state cell.
-Finally, the output gate has the power to shut off the output cell \parencite{goodfellow_deep_2016}.
-Here, a notable implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 layers each.
-Compared to C-RNN-GAN, here the network is completely unidirectional.
-Probably more important than the network architecture, a central piece of the model's performance is due to the final representation of the MIDI data: each time slice is represented as a one-hot encoded vector containing Note-On (128 values), Note-Off (128 values), Time-Shift (125 values), and Velocity (32 values) events \parencite{oore_this_2018}.
 

### 2019-04-29 src/c_rnn_gan.py

@@ -87,7 +87,7 @@ class Generator(nn.Module):
         return torch.LongTensor([[self.primary_event] * batch_size]).to(
             self.device)
 
-    def _sample_event(self, output, greedy=True, temperature=1.0):
+    def _sample_event(self, output, greedy=False, temperature=1.0):
         if greedy:
             return output.argmax(-1)
 
@@ -136,7 +136,7 @@ class Generator(nn.Module):
                  init,
                  steps,
                  y=None,
-                 greedy=True,
+                 greedy=False,
                  temperature=1.0,
                  output_type='index'):
 

### 2019-04-29 src/data.py

@@ -141,7 +141,7 @@ def from_seq(sequence, path=None):
                     velocity=velocity,
                     time=move_by))
         elif message[0] == 'control_change':
-            value = 127 if int(message[1]) == 1 else 0
+            value = 100 if int(message[1]) == 1 else 1
             track.append(
                 Message(
                     message[0],
@@ -154,7 +154,7 @@ def from_seq(sequence, path=None):
     track.append(MetaMessage('end_of_track', time=1))
 
     if path:
-        with open(path, 'r') as my_midi:
+        with open(path, 'w') as my_midi:
             midi.save(file=my_midi)
 
     return midi

### 2019-04-29 src/generate.py

@@ -19,9 +19,9 @@ if __name__ == '__main__':
     median_len = int(df['seq_len'].median())
 
     greed = False
-    temperature = 1.1
+    temperature = 1.0
 
-    models = ['baseline']
+    models = ['performance_rnn']
 
     for model in models:
         checkpoint = load_model_data(os.path.join('../models/', model))
@@ -38,18 +38,32 @@ if __name__ == '__main__':
                 use_bias=checkpoint['use_bias'],
                 is_bidirectional=checkpoint['is_bidirectional'])
             net.load_state_dict(checkpoint['model_state'])
-        for i in tqdm(range(1, 101), desc=f'Generating from {model}'):
+        elif model == 'performance_rnn':
+            from performance_rnn import PerformanceRNN
+            net = PerformanceRNN(
+                in_dim=checkpoint['in_dim'],
+                hidden_dim=checkpoint['hidden_dim'],
+                batch_size=1,  # we create one song at a time
+                out_dim=checkpoint['out_dim'],
+                dropout=checkpoint['dropout'],
+                num_layers=checkpoint['num_layers'],
+                init_dim=checkpoint['init_dim'],
+                use_bias=checkpoint['use_bias'],
+                is_bidirectional=checkpoint['is_bidirectional'])
+            net.load_state_dict(checkpoint['model_state'])
+        for i in tqdm(range(1, 2), desc=f'Generating from {model}'):
             init = torch.randn(net.batch_size, net.init_dim).to(device)
-            steps = median_len * (RNG.randint(90, 111) / 100)
+            steps = int(median_len * (RNG.randint(90, 111) / 100))
             song = net.generate(
                 init,
                 steps,
                 greedy=greed,
                 temperature=temperature,
                 output_type='index')
-            song = song.tolist()
+            song = song.squeeze().tolist()
             make_folder('../generated')
-            from_seq(
-                song,
-                path=os.path.join('../generated/',
-                                  f'{model}_{i}_{greed}_{temperature}.midi'))
+            song = from_seq(song)
+            song.save(
+                filename=os.path.join(
+                    '../generated/',
+                    f'{model}_{i}_{greed}_{int(temperature*10)}.midi'))

### 2019-04-29 src/lstm_baseline.py

@@ -84,7 +84,7 @@ class Baseline(nn.Module):
         return torch.LongTensor([[self.primary_event] * batch_size]).to(
             self.device)
 
-    def _sample_event(self, output, greedy=True, temperature=1.0):
+    def _sample_event(self, output, greedy=False, temperature=1.0):
         if greedy:
             return output.argmax(-1)
 
@@ -133,7 +133,7 @@ class Baseline(nn.Module):
                  init,
                  steps,
                  y=None,
-                 greedy=True,
+                 greedy=False,
                  temperature=1.0,
                  output_type='index'):
 

### 2019-04-29 src/performance_rnn.py

@@ -85,7 +85,7 @@ class PerformanceRNN(nn.Module):
         return torch.LongTensor([[self.primary_event] * batch_size]).to(
             self.device)
 
-    def _sample_event(self, output, greedy=True, temperature=1.0):
+    def _sample_event(self, output, greedy=False, temperature=1.0):
         if greedy:
             return output.argmax(-1)
 
@@ -134,7 +134,7 @@ class PerformanceRNN(nn.Module):
                  init,
                  steps,
                  y=None,
-                 greedy=True,
+                 greedy=False,
                  temperature=1.0,
                  output_type='index'):
 

### 2019-04-29 src/train_c_rnn_gan.py

@@ -260,7 +260,7 @@ class ReptileGAN:
 
     def train_generator(self,
                         feature_matching=True,
-                        greedy=True,
+                        greedy=False,
                         temperature=1.0):
         # Generator
         self.meta_g_optim.zero_grad()

### 2019-04-30 paper/main.tex
We generate better now

@@
 
-In a standard ANN, the data traverses the network through the input layer, to the hidden layer, and finally reaching the output layer; setup which is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}.
+In a standard ANN, the data traverses the network through the input layer, to the hidden layer, and finally reaches the output layer; setup which is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}.
 Broadly, this type of network is defined by the fact that its connections do not form cycles.
@@
 
-Concretely, information passes through the RNN similarly to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and prior hidden activations from the previous timestep \parencite{graves_supervised_2012}.
-Thus a formal definition of the network would be represented by
+The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and prior hidden activations from the previous timestep \parencite{graves_supervised_2012}.
+Moreover, recurrent networks also benefit from the fact that parameters (weights) are shared across different parts of a model.
+This allows extending and applying the model to data points of different lengths and generalizing across all of them \parencite{goodfellow_deep_2016}.
+In contrast, a feedforward network would require separate parameters for each input feature, at each position.
+Thus, a formal definition of the network's forward pass would be
+\begin{equation}
+    h^t = f(h^{t-1}; x^t; \theta) \label{eq:rnn_hidden}
+\end{equation}
+where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the bias.
 
-\begin{equation}
-    h^t = f(h^{t-1}; x^t; \theta)
-\end{equation}
+The backward pass in a RNN is similar to that of a feedforward network: given the partial derivatives of a differentiable loss function $\mathcal{L}$ we calculate the derivatives with respect to the network's weights \parencite{graves_supervised_2012}.
+In this setting, a popular algorithm is called backpropagation through time
 
-where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the bias.
+Equation \ref{eq:rnn_hidden} can be drawn as in Figure \ref{fig:rnn_folded}: input sequences $x$ are mapped to a hidden layer $h$ and an output $y$, with the state of the hidden layer at time $t$ feeding back into itself with a time delay of $1$.
+In other words, the state of $h_t$ will influence the decision of the recurrent layer at time $t+1$.
+Another way to illustrate this is to unfold the computational graph such that each component of the network is represented by different variables per step, as in Figure \ref{fig:rnn_unfolded}.
 

### 2019-04-30 src/generate.py

@@ -19,11 +19,14 @@ if __name__ == '__main__':
     median_len = int(df['seq_len'].median())
 
     greed = False
-    temperature = 1.0
+    temperatures = [0.8, 0.9, 1.0, 1.1, 1.2]
 
     models = ['performance_rnn']
 
+    make_folder('../generated')
     for model in models:
+        make_folder(f'../generated/{model}')
+
         checkpoint = load_model_data(os.path.join('../models/', model))
         if model == 'baseline':
             from lstm_baseline import Baseline
@@ -51,19 +54,33 @@ if __name__ == '__main__':
                 use_bias=checkpoint['use_bias'],
                 is_bidirectional=checkpoint['is_bidirectional'])
             net.load_state_dict(checkpoint['model_state'])
-        for i in tqdm(range(1, 2), desc=f'Generating from {model}'):
-            init = torch.randn(net.batch_size, net.init_dim).to(device)
-            steps = int(median_len * (RNG.randint(90, 111) / 100))
-            song = net.generate(
-                init,
-                steps,
-                greedy=greed,
-                temperature=temperature,
-                output_type='index')
-            song = song.squeeze().tolist()
-            make_folder('../generated')
-            song = from_seq(song)
-            song.save(
-                filename=os.path.join(
-                    '../generated/',
-                    f'{model}_{i}_{greed}_{int(temperature*10)}.midi'))
+        elif model == 'c_rnn_gan':
+            from c_rnn_gan import Generator
+            net = Generator(
+                in_dim=checkpoint['in_dim'],
+                hidden_dim=checkpoint['hidden_dim'],
+                batch_size=1,  # we create one song at a time
+                out_dim=checkpoint['out_dim'],
+                dropout=checkpoint['dropout'],
+                num_layers=checkpoint['num_layers'],
+                init_dim=checkpoint['init_dim'],
+                use_bias=checkpoint['use_bias'])
+
+        for temp in temperatures:
+            make_folder(f'../generated/{model}/{temp}')
+
+            for i in tqdm(
+                    range(1, 126),
+                    desc=f'Generating from {model} with {temp}'):
+                init = torch.randn(net.batch_size, net.init_dim).to(device)
+                steps = int(median_len * (RNG.randint(90, 111) / 100))
+                song = net.generate(
+                    init,
+                    steps,
+                    greedy=greed,
+                    temperature=temp,
+                    output_type='index')
+                song = song.squeeze().tolist()
[…]

### 2019-05-05 paper/main.tex
done with methods section

@@
 \usepackage{tabularx}
+\usepackage{supertabular}
 \usepackage{graphicx}
@@
 \usepackage{tikz}
-\usetikzlibrary{arrows.meta, fit, positioning}
+\usetikzlibrary{arrows.meta, fit, positioning, shapes}
+\newcommand{\empt}[2]{$#1^{\langle #2 \rangle}$}
 
@@
         -{Straight Barb[angle=60:2pt 3]},
-    }
+    },
+    connl/.style={
+        thick,
+        {Straight Barb[angle=60:2pt 3]}-,
+    },
+    prod/.style={circle, draw, thick, inner sep=0pt},
+    ct/.style={circle, draw, inner sep=5pt, ultra thick,
+		minimum width=10mm},
+    ft/.style={circle, draw, thick, minimum width=8mm, inner sep=1pt},
+    filter/.style={circle, draw, thick, minimum width=7mm, inner sep=1pt,
+		path picture={
+			\draw[thick, rounded corners] (path picture bounding box.center)--++(65:2mm)--++(0:1mm);
+			\draw[thick, rounded corners] (path picture bounding box.center)--++(245:2mm)--++(180:1mm);}},
 }
@@
 This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}.
-Section \ref{sec:lstm} will present the principal network architecture of this paper, the Long Short-Term Memory network, and argument for its use as the foundation of a generative network.
+Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network.
 Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning, several approaches to it, and finally narrow down to current state-of-the-art algorithms.
 
-\section{Long Short-Term Memory Networks} \label{sec:lstm}
+\section{Recurrent Neural Networks} \label{sec:rnn}
 
@@
 If however we relax this condition, we arrive at a recurrent neural network \parencite[Figure \ref{fig:rnn_folded};][]{rumelhart_learning_1986}: a family of neural networks specifically designed for processing sequential data \parencite{goodfellow_deep_2016}.
+It is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution would be the basis for building a time-delay neural network.
+Unfortunately, the output of such a network would be a function of a small number of neighbouring data points \parencite[say, the last three notes;][]{goodfellow_deep_2016} whereas in the contex of music, knowledge of the notes played from the begining of the song is needed.
+As such, RNNs seem to be a more compelling alternative for this task.
 
@@
 \end{equation}
-where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the bias.
+where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the model parameters.
 
 The backward pass in a RNN is similar to that of a feedforward network: given the partial derivatives of a differentiable loss function $\mathcal{L}$ we calculate the derivatives with respect to the network's weights \parencite{graves_supervised_2012}.
-In this setting, a popular algorithm is called backpropagation through time
+In this setting, a popular algorithm is called backpropagation through time \parencite[BPTT;][]{werbos_backpropagation_1990} which consists of repeated applications of the chain rule.
+The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer a the next timestep \parencite{graves_supervised_2012}.
 
@@
 Another way to illustrate this is to unfold the computational graph such that each component of the network is represented by different variables per step, as in Figure \ref{fig:rnn_unfolded}.
+
+The reason why these notions have been introduced is to be able to bring up the following point.
+As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}.
+\textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing bacwards in time tend to either (1) blow up or (2) vanish.
+As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
     \centering
@@
     \end{subfigure}
+    \caption{A typical view of an RNN (left) and an unfolded view of the same network (right).
+In both sub-figures each node is a layer of network units however, in the second representation these are shown for each timestep.
+The weighted connections from the input layer to the hidden one is labelled $U$, those from the hidden layer feeding back to itself are $W$, and those from the hidden layer to the output layer are labelled as $V$.
+Note that these weights are reused at each timestep.
+Bias has been omitted in this representation.
+Figure adapted from \textcite{graves_supervised_2012}}
+\end{figure}
+
+The LSTM is made of a network of recurrently connected subnetworks known as memory blocks, each with one or more self-connected memory cells and three multiplicative units: the input gate, the output gate, and the forget gate \parencite{hochreiter_long_1997}.
+The three gates act as analogies for the read, write, reset computer operations and allow the memory cells to store and access information over long periods of time \parencite{graves_supervised_2012}.
+See Figure \ref{fig:lstm} for an illustration of a LSTM memory block with a single cell.
+
+Aside from being successful at a variety of classification tasks \parencite[e.g.,][]{graves_framewise_2005}, this architecture has been successfully applied to generating new data points \parencite[such as creating sequences of text;][]{graves_generating_2013}.
+In the context of generating musical performances, experiments with this neural architecture go as far as \textcite{eck_first_2002} however, a more noteworthy implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 cells each designed to generate classical compositions whilst maintaining the playing style of expert pianists.
+Considering how successful it was, this model will be re-implemented in this paper's experiments (see Chapter \ref{chap:methods}).
+
+\begin{figure}[t]
+    \centering
+    \begin{tikzpicture}
+        \node[ct, label={Cell}] (ct) {$c^t$};
+        \node[filter, right=of ct] (int1) {};
+        \node[prod, right=of int1] (x1) {$\times$}; 
+        \node[right=of x1] (ht) {$h^t$};
+        \node[prod, left=of ct] (x2) {$\times$}; 
+        \node[filter, left=of x2] (int2) {};
+        \node[prod, below=5mm of ct] (x3) {$\times$}; 
+        \node[ft, below=5mm of x3, label={right:Forget Gate}] (ft) {$f^t$};
+        \node[ft, above=of x2, label={right:Input Gate}] (it) {$i^t$};
+        \node[ft, above=of x1, label={right:Output Gate}] (ot) {$o^t$};
+
+        \foreach \i/\j in {int2/x2, x2/ct, ct/int1, int1/x1,
+                        x1/ht, it/x2, ct/it, ct/ot, ot/x1, ft/x3}
+                            \draw[->, conn] (\i)--(\j);
+
+                            \draw[->, conn] (ct) to[bend right=45] (ft);
+
+                            \draw[->, conn] (ct) to[bend right=30] (x3);
+                            \draw[->, conn] (x3) to[bend right=30] (ct);
+
+                            \node[fit=(int2) (it) (ot) (ft), draw, inner sep=0pt] (fit) {};
+
+                            \draw[<-, connl] (fit.west|-int2) coordinate (aux)--++(180:7mm) node[left]{$x^t$};
+                            \draw[<-, connl] ([yshift=1mm]aux)--++(135:7mm);
+                            \draw[<-, connl] ([yshift=-1mm]aux)--++(-135:7mm);
+
+                            \draw[<-, connl] (fit.north-|it) coordinate (aux)--++(90:7mm) node[above]{$x^t$};
+                            \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
+                            \draw[<-, connl] ([xshift=-1mm]aux)--++(135:7mm);
+
+                            \draw[<-, connl] (fit.north-|ot) coordinate (aux)--++(90:7mm) node[above]{$x^t$};
+                            \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
+                            \draw[<-, connl] ([xshift=-1mm]aux)--++(135:7mm);
+
+                            \draw[<-, connl] (fit.south-|ft) coordinate (aux)--++(-90:7mm) node[below]{$x^t$};
+                            \draw[<-, connl] ([xshift=1mm]aux)--++(-45:7mm);
+                            \draw[<-, connl] ([xshift=-1mm]aux)--++(-135:7mm);
+    \end{tikzpicture}
+    \caption{LSTM memory block with one cell.
+Here the gates are nonlinear summation units that collect activations from the inside and outside of the memory block and control the activation of the cell $c^t$ via multiplications.
+The input and output gates ($i^t$, $o^t$) multiply the block's input and output activations respectively, while the forget gate ($f^t$) multiplies the previous hidden state.
+The input and output activation are typically tanh or sigmoid.
+The only outputs ($h^t$) from the block come from the output gate multiplication.
+Figure adapted from \textcite{graves_supervised_2012}}
+    \label{fig:lstm}
 \end{figure}
@@
 
-Turning over to the generative side of things, nowadays, the most popular form of generating data is through the use of a Generative Adversarial Network \parencite[GAN;][]{goodfellow_generative_2014}.
-Here a generative model $G$ captures the data distribution and creates samples from it, whereas a discriminative model $D$ estimates the probability that the input is real (i.e., comes from the training data) rather than fake (i.e., comes from $G$).
-Formally, given $x$ sampled from real data, and $z$ being input noise, $D$ tries to keep $D(x)$ near $1$, while making $D(G(z))$ near $0$.
-Conversely, $G$ tries to make $D(G(z))$ near $1$ \parencite{goodfellow_nips_2016}.
-A notable implementation is C-RNN-GAN \parencite{mogren_c-rnn-gan_2016}, a continuous recurrent neural network with adversarial training.
-It consists of a network with two LSTM layers and 350 hidden units per layer, in both $G$ and $D$.
-However, $D$ has a bidirectional layout which allows is to take context from both past and future for its decisions, while $G$ is unidirectional \parencite{mogren_c-rnn-gan_2016}.
-As is the standard practice for GANs, $G$ will create fake samples of music, whereas $D$ will try to classify correctly which sample is real.
+Nowadays, the most popular form of generating data is through the use of a Generative Adversarial Network \parencite[GAN;][]{goodfellow_generative_2014}.
+The training procedure is set up as a game between two networks: a generator and a discriminator.
+The purpose of the former is to create samples that come from the same distribution as the training data, whereas the latter aims to determine which samples are real or fake \parencite{goodfellow_nips_2016}.
 
-Notwithstanding GANs, other generative successes have made use of adaptation on the Long Short-Term Memory (LSTM) network \parencite{hochreiter_long_1997}.
-Here, the cells of a network are designed to allow for information to persist by implementing a recursive state unit, an input gate, a forget gate, and an output gate.
-The input of the network is allowed into the state unit if it is allowed to pass through by the input gate.
-Next, the forget gate controls the self-loop of the state cell.
-Finally, the output gate has the power to shut off the output cell \parencite{goodfellow_deep_2016}.
-Here, a notable implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 layers each.
-Compared to C-RNN-GAN, here the network is completely unidirectional.
-Probably more important than the network architecture, a central piece of the model's performance is due to the final representation of the MIDI data: each time slice is represented as a one-hot encoded vector containing Note-On (128 values), Note-Off (128 values), Time-Shift (125 values), and Velocity (32 values) events \parencite{oore_this_2018}.
+To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$.
+The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar where $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}.
+We train $D$ to maximize the probability of correctly assigning a label to the training data and samples from $G$ ("real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, without actually looking at $p_{\text{data}}$.
+In other words, $G$ must minimize $\log (1- D(G(z)))$ \parencite{goodfellow_generative_2014}.
+
+In other words, both models play a minimax game with a value function $V(G, D)$:
+\begin{equation}
+    \min_G \max_D V(D, G) = \mathbb{E}_{x ~ p_{\text{data}}} [ \log D(x) ] + \mathbb{E}_{z ~ p_z (z)} [ \log (1-D(G(z))) ]
+\end{equation}
+where the solution is for $G$ to recover the training data distribution and $D$ to have a classification accuracy of 50\% \parencite{goodfellow_generative_2014}.
+
+Although new, GANs are a significant conceptual and technical innovation \parencite{briot_deep_2017} which can be seen in them revitalizing interest in generative models.
+As mentioned before, they have been proven successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}.
+Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by re-purposing this architecture to generate music.
+The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a unidirectional LSTM network with a similar structure as discriminator.
 
@@
 
-\textcite{lake_human-level_2015} initially popularized the idea of learning from a few class examples using a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
-\textcite{rezende_one-shot_2016} were among the first to provide a solution to the one-shot problem in the context of a neural network (NN) by implementing Bayesian reasoning into a deep NN embedded within hierarchical latent variable models.
-The Reptile algorithm \parencite{nichol_first-order_2018} is a recent approach to this problem.
-It counteracts the shortcomings of gradient-optimization algorithms by learning the model's initial parameters in order to maximize its performance on novel tasks.
-Furthermore, it allows for any type of network to be used and does not place any restrictions on the type of loss function that can be used \parencite{nichol_first-order_2018}.
-The algorithm for Reptile is as follows:
+The rationale for investing time and energy into meta-learning is that we cannot capture the essence of learning simply by relying on a small number of algorithms, but rather a bunch of context-dependent learning strategies that would enable us to properly capture domain specific information \parencite{schmidhuber_evolutionary_1987}.
+Given the complexities of each learning strategy, \textcite{schmidhuber_evolutionary_1987} argues that an obvious way to tackle this initiative is to give a system the ability to learn how to learn.
+A simplistic way to do this would be to train a system on different tasks (e.g, image classification, text classification, etc.).
+However, this approach entails training on a dataset that is much larger that what current hardware can withstand.
+Furthermore, this approach would stray away from the idea of building general-purpose A.I.
+architectures as generalizing over a task would require more data points that what a human would need.
+To be more concrete, a human is able to generalize a task/object after encountering one (or few) example(s) \parencite{ravi_optimization_2016, chen_closer_2018}.
+Therefore, besides having to perform on a variety of tasks, a systems would need to excel at that using sparse amounts of data.
+
+Although ideas of meta-learning can go as far as Schmidhuber's thesis in 1987, it seems that training a network using a handful of examples did not get much traction until \textcite{lake_human-level_2015} successfully designed a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
+For clarity, such a problem is typically defined as a $n$-shot $k$-way problem, where $n$ stands for the number of examples per class, and $k$ is the number of classes to be used in the training process.
+
+Initially, it was thought that such a framework would not work in the context of an ANN since the iterative nature of gradient-based optimization algorithms do work well in the context of a set number of updates \parencite{ravi_optimization_2016} - and even more so when there are only a few examples to work with.
+Having said that, \textcite{rezende_one-shot_2016} were among the first to provide a solution to the one-shot problem in the context of a neural network by implementing Bayesian reasoning into a deep network embedded within hierarchical latent variable models.
+Since that, numerous efforts were made to develop an algorithm that enables an ANN to perform with a constrained number of class examples.
+Probably the most popular algorithm at the time of writing is the Model-Agnostic Meta-Learning (MAML) algorithm \parencite{finn_model-agnostic_2017}.
+One of MAML's core mechanics is the parameter update by calculating second-order derivatives.
+This procedure however implies a substantial computational penalty, which the authors did note.
+Therefore, they developed a first-order variant of MAML (FOMAML) which is able to perform similarly, if only slightly worse \parencite{finn_model-agnostic_2017}.
+
+This however, seemed to be the initial thought, until \textcite{nichol_first-order_2018} showed that FOMAML tends to be somewhat unstable and very sensitive to even the smallest changes in hyper-parameters.
+Therefore, they developed Reptile, a first-order meta-learning algorithm that rivals both MAML and FOMAML \parencite{nichol_first-order_2018}.
+Just like MAML, it counteracts the shortcomings of gradient-optimization algorithms by learning the model's initial parameters in order to maximize performance on novel tasks, it allows for any type of network to be used, and does not place any restrictions on the type of loss function that can be used \parencite{nichol_first-order_2018}.
+The difference between the two however, lies in the way the parameters are updated.
 
@@
 \begin{algorithm}
-Initialize $\phi$, the vector of initial parameters
+Initialize $\theta$, the vector of initial parameters
 
 \For{iteration = $1, 2, \dots$}{
-    Sample task $\tau$, corresponding to loss $L_{\tau}$ on weight vectors $\tilde \phi$
+    Sample task $\tau$, corresponding to loss $L_{\tau}$ on weight vectors $\tilde \theta$
 
-    Compute $\tilde \phi = U_{\tau}^{k} (\phi)$, denoting $k$ steps of SGD or Adam
+    Compute $\tilde \theta = U_{\tau}^{k} (\theta)$, denoting $k$ steps of SGD or Adam
 
-    Update $\phi \leftarrow \phi + \epsilon (\tilde{\phi} - \phi)$
+    Update $\theta \leftarrow \theta + \epsilon (\tilde{\theta} - \theta)$
 }
@@
 
-Although simplistic, this algorithm reaches similar results to another popular few-shot algorithm called MAML \parencite{finn_model-agnostic_2017}.
-Both algorithms are initialization-based learning methods however, a major distinction between the two is that the vanilla MAML implementation requires calculating a second-order derivative, which can significantly slowdown computation \parencite{nichol_first-order_2018}.
-This forms a compelling reason to use Reptile as the few-shot learning method for this research.
+In essence, the way Reptile works (see Algorithm \ref{alg:reptile}) is as follows.
+A model with parameters $\theta$ is initialized, and a meta-loop (or meta-epochs) of $i$ iterations starts.
+Within it, a task $\tau$ (e.g, image classification, text classification, etc.) is sampled, along with its corresponding loss function $L_{\tau}$ (for clarity we assume that each type of task uses a different loss function).
+Then we compute the updated parameters $\tilde \theta$ in an inner-loop with $k$ iterations on task $\tau$ (or in other words, once we sampled our task $\tau$, we train our model $k$ times on it using the $\theta$ as parameters so we can get $\tilde \theta$).
+Finally, we update out initial parameters $\theta$ using this formula: $\theta \leftarrow \theta + \epsilon (\tilde{\theta} - \theta)$.
+Then the algorithm continues to the next iteration of the meta-loop carrying over our updated parameters $\theta$.
+
+The explanation above is fairly verbose however, the algorithm itself is fairly simplistic but achieves similar performances to FOMAML \parencite{nichol_first-order_2018}.
+This forms a compelling reason to use Reptile as the few-shot learning method of choice for this paper.
 
@@
 
-To provide an answer to the above research questions, this study will develop (in Keras) two generative models based on C-RNN-GAN and Performance-RNN, where the learning procedure will incorporate the Reptile algorithm.
-Other than that, the models will not differ from their original implementation.
-These models will firstly be compared to the baseline proposed by \textcite{larochelle_few-shot_2017} (a LSTM network trained on the entire train set) and then with the music from the dataset.
-The evaluation between the models with the baseline will be based on the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018} and domain specific measurements (polyphony, scale consistency, repetitions, tone span).
-The evaluation between the model's generated samples and real music in the dataset will be based on the domain specific metric.
-Refer to the evaluation section for more details on these metrics.
+To reiterate from Section \ref{sec:thesis_goal}, the purpose of this paper is to establish (1) the extent to which the music generated under a few-shot learning algorithm is comparable to that of a model trained on the entire dataset, and (2) the extent to which the music from such a model is comparable to that of real musical performances.
+For these purposes, Performance-RNN and C-RNN-GAN were adapted to the Reptile training procedure (Algorithm \ref{alg:reptile}) and evaluated against the performances generated by the baseline proposed in \textcite{larochelle_few-shot_2017} - a single layer LSTM with 256 units - and actual musical performances from the dataset (Section \ref{sec:dataset}).
+For an in-depth explanation on the steps required to develop a model that is able to generate new data points refer to Section \ref{sec:how_to_gen} and to Section \ref{sec:reptile_train} for details on the training procedure.
+The evaluation of the models' performances will be based on the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018} and several domain specific measurements adapted from \textcite{mogren_c-rnn-gan_2016}: polyphony, scale consistency, repetitions, tone span.
+Refer to Section \ref{sec:evaluation} for more details on these metrics.
 
-\section{Dataset}
+\section{Dataset} \label{sec:dataset}
 
@@
 All performances were done by piano experts on a Yamaha Disklavier, instrument which integrates a highly precise MIDI capture and playback system.
-Recorded MIDI events are of sufficient fidelity to allow judges to remotely listen to the contestant's performance (also on a Disklavier).
+The claim is that the recorded MIDI events are of sufficient fidelity to allow judges to remotely listen to the contestant's performance (also on a Disklavier).
 
@@
 This amounts to 1,184 piano performances, approximately 430 compositions, 6.18 million notes played, and approximately 172 hours of playback.
-There is also a recommended train/validation/test split created on the following criteria \parencite{hawthorne_enabling_2018}:
+There is also a recommended train/validation/test split (954, 105, 125 songs respectively) created on the following criteria \parencite{hawthorne_enabling_2018}:
 
@@
 
-Moreover, each performance comes with additional metadata namely, the name of the composer, the title of the performance, the suggested train, validation, test splits, year of the performance, name of the file, and duration in seconds of the performance.
+Moreover, each performance comes with additional metadata such as, the name of the composer, the title of the performance, the suggested train, validation, test splits, year of the performance, name of the file, and duration in seconds of the performance.
+Having said that, the dataset lacks clearly defined classes.
+The canonical composer could be used as a substitute however, the author strongly believes that there is a better way to categorize music: by genre.
 
-\section{Evaluation}
+Broadly speaking, Western classical music can be segregated into a number of genres based on the year of the composition and style of the piece.
+For this we can use the Concise Oxford Dictionary of Music \parencite{kennedy_concise_2007} which indicates to the following classical genres: baroque (1600 - 1750), classical (1750 - 1820), romanticism (1780 - 1910), modernism (1890 - 1930), and impressionism (circa 1890 - 1925).
+Granted, there are far more genres, but these are the ones that are covered by the dataset.
+As such, given the definitions of \textcite{kennedy_concise_2007} on said genres and artists, each individual piece from our dataset was manually assigned a genre.
+Concretely, the style of music was firstly determined by the period in which the composer has lived (in baroque, classical, etc.) or - if the genres would overlap (i.e., modernism, and impressionism) - the composer's primary style of music would be chosen according to the definition of the composer given by \textcite{kennedy_concise_2007}.
+Said labelling can be found in Appendix \ref{appendix:labelling}.
+
+\section{Data Processing} \label{sec:data_proc}
+
+\begin{figure}[t]
+    \centering
+    \begin{verbatim}
+<note_on note=69 velocity=73 time=33>
+<note_on note=71 velocity=78 time=15>
+<control_change control=64 value=83 time=27>
+<note_off note=69 velocity=0 time=6>
+<note_off note=71 velocity=0 time=3>
+    \end{verbatim}
+    \caption{Example of MIDI messages in the dataset, as printed by the Python library Mido \parencite{bjorndalen_mido_2018}.
+At 33 ticks since the last event, the player pressed an A4 (note 69), followed by a B4 (note 71), 15 ticks later.
+Next, the pedal (Control Change message with a value of 64) is pressed 27 ticks later, and the keys are released 6 and 3 tick later.}
+    \label{fig:midi_examples}
+\end{figure}
+
+The dataset consists of MIDI files which would require some processing in order to be ready for a LSTM network.
+Here, the approach of \textcite{oore_this_2018} was adopted where MIDI midi messages are transformed into a sequence of one-hot encoded events.
+More specifically, MIDI excerpts are represented as a sequence of events from a vocabulary of 416 different events - similar to \textcite{oore_this_2018}, but with three extra events in the vocabulary:
+\begin{itemize}
+    \item 128 Note-On events: starts a new note (one for each of the 128 MIDI pitches)
+    \item 128 Note-Off events: releases a note (one for each of the 128 MIDI pitches)
+    \item 126 Time-Shift events: moves the time step forward somewhere between 0 ms and 1 second, in increments of 8 ms
+    \item 32 Velocity events: changes the velocity for each subsequent note, until the next velocity event
+    \item 2 Pedal events: triggers or releases the pedal
+\end{itemize}
+
+A glance over the MIDI documentation \parencite{midi_association_official_2019} shows how complex and granular the standard is.
+However, we do not need to implement all of the messages into our model in order to make it work.
+We only need three: \texttt{note\_on}, \texttt{note\_off}, \texttt{pedal\_on/off}.
+Figure \ref{fig:midi_examples} show an example of how the midi messages look like.
+It should be noted that in this figure, the time is represented in ticks, MIDI's smallest unit of time, who's value is usually a third of a beat \parencite{midi_association_official_2019}.
+
+As it can be seen in Figure \ref{fig:midi_examples}, each message contains multiple pieces of information such as the message type, the note/value of the message, the delta time (time elapsed since last message and the current message, measured in ticks), and velocity (if applicable).
+To one-hot encode this stream of messages, it is necessary to separate it so it contains only one type of information per entry.
+Thus, the sequence can be further broken down such that the first message that the system encounters is the delta time, followed by the velocity, and the note/pedal event (and their respective values).
+An example of how the MIDI sequences are broken down into singular pieces of information can be found in Figure \ref{fig:representation_midi}.
+
+\begin{figure}[t]
+    \begin{tikzpicture}
+        \draw[step=1cm,base01,very thin] (0,0) grid (6,6);
+        \filldraw[fill=base01, draw=base01] (0,1) rectangle (1,2);
+        \filldraw[fill=base01, draw=base01] (2, 3) rectangle (5, 4);
+        \draw (6.5, 5) node[anchor=south west, color=base09] {1. Move time by 1 tick};
+        \draw (6.5, 4) node[anchor=south west, color=base0A] {2. Set velocity to 80};
+        \draw (6.5, 3) node[anchor=south west, color=base0B] {3. Note 3 on};
+        \draw (6.5, 2) node[anchor=south west, color=base0C] {4. Move time by 3 ticks};
+        \draw (6.5, 1) node[anchor=south west, color=base0D] {5. Note 3 off};
+
+        \draw[ -{Straight Barb[angle=60:2pt 3]}, color=base09, thick] (1, 1) -- (2, 1) node[anchor=north, midway] {1};
+        \draw[color=base0A, thick] (2, 1) -- (2, 3) node[anchor=east, midway] {2};
+        \draw[color=base0B, thick] (2, 3) -> (2, 4) node[anchor=east, midway] {3};
+        \draw[ -{Straight Barb[angle=60:2pt 3]}, color=base0C, thick] (2, 3) -- (5, 3) node[anchor=north, midway] {4};
+        \draw[color=base0D, thick] (5, 3) -- (5, 4) node[anchor=west, midway] {5};
+
+        \draw[thick,-{Straight Barb[angle=60:2pt 3]}] (0,0) -- (6.5,0) node[anchor=north west] {Time (in ticks)};
+        \draw[thick,-{Straight Barb[angle=60:2pt 3]}] (0,0) -- (0,6.5) node[anchor=south east] {Notes};
+        \foreach \x in {0,1,2,3,4,5,6}
+           \draw (\x cm,1pt) -- (\x cm,-1pt) node[anchor=north] {$\x$};
+        \foreach \y in {0,1,2,3,4,5}
+            \draw (1pt,\y.5 cm) -- (-1pt,\y.5 cm) node[anchor=east] {$\y$};
+    \end{tikzpicture}
+    \caption{Example of how the data is represented.
+The progression illustrates how a piano roll of MIDI data is converted into a sequence of commands (right hand side) that belong to our events vocabulary.
+Figure adapted from \textcite{oore_this_2018}.}
+    \label{fig:representation_midi}
+\end{figure}
+
+Having said these, it should be mentioned that the one-hot representation of MIDI messages has undergone some fine quantization, again, in a similar fashion to \textcite{oore_this_2018}.
+The delta times have been modified to correspond to a change of multiples of 8 ms (i.e., 0 ms, 8 ms, 16 ms, etc.) and the velocities have been partitioned into 32 steps as opposed to MIDI's 128.
+With regards to the time movements, \textcite{friberg_overview_2006} note that a noticeable difference in temporally displacing a single tone in a sequence takes place from 10 ms and above.
+When in comes to the note velocities, it should be pointed to the reader that classical music features eight levels of common dynamic marking (from $ppp$ to $fff$) which is why \textcite{oore_this_2018} note that 32 steps are more than enough.
+
+\section{You're a musician Schmidhuber} \label{sec:how_to_gen}
+
+Knowing what we know about RNNs and LSTMs from Section \ref{sec:rnn} one can devise a method where the network can generate a sequence of events.
+All that is needed is to present the system with an event to kickstart the generative process.
+Namely, an input $x^t$ of size 1 is given to the network to output the next likely event $x^{t+1}$, which in turn is fed back to generate $x^{t+2}$.
+This process is repeated until we have a sequence of events of a desired length.
+
+An alternative to this would be to define a noise vector $z^t$ of an arbitrary size (e.g., 100), and a random input $x^t$ selected from our events vocabulary.
+The noise vector is passed through a linear layer that output a tensor that would serve as the network's hidden state $h^{t}$ at time $t$.
+It is then that $x^t$ and $h^{t}$ are fed into the LSTM layer to generate $x^{t+1}$ and $h^{t+1}$.
+The reason for pre-defining the hidden state for the first input is to increase the number of possible outcomes coming from the network.
+
+To be more precise, if we were to present the network with only $x^t$ then the network would mark the input sequence as the beginning of the song.
+With that information, it is possible that the network would affix itself to following a specific pattern.
+For instance, if the first input would be a C\#, the network could learn to always output E as the next note.
+Using a random hidden state is a measure that was included to ensure some diversity in the process of generating samples (granted, it should be admitted the possibility of this scenario being unlikely, and that the regular method described above would still work just as good).
+
+Having said this, diversity is not the only reason for including a noise vector $z$ into the training process.
+The reader might recall that this is part of the generation process in the framework of a GAN respectively, how the generator $G$ would create samples.
+Hence, all three models implemented in this paper feature the architecture of a GAN generator.
+
+In terms of generating the next event, \textcite{oore_this_2018} have defined a parameter $\tau$ called temperature \parencite{google_magenta_performance_2017}.
+It is a value which uses the model's predicted events distribution to control the randomness of the output samples.
+In other words, let $\phi$ be the output of a LSTM network (for clarity, we ignore the hidden state $h$).
+Then, the next sequence $x^{t+1}$ is defined as
+\begin{equation}
+    x^{t+1} = X(\omega)
+\end{equation}
+where
+\begin{equation}
+    X = \text{softmax}(\phi / \tau).
+\end{equation}
+$X$ is the predicted event distribution of the network's output divided by our temperature $\tau$. Thus, $x^{t+1}$ is actually a random sample from our defined categorical distribution $X$.
+
+\section{How to train a Reptile} \label{sec:reptile_train}
+
+With the generative process outlined in the previous section, training the generative networks outlined in this papers is somewhat straightforward.
+A batch of songs is sampled from the training set, the model generates as many songs as there are in the batch, and then compute the loss between the network's output and the real samples.
+This statement is essentially true for the baseline and PerformanceRNN (in spite of the Reptile training procedure, inner-loop training is similar to the standard procedure).
+In contrast, the GAN model requires training two models simultaneously, where the generator's loss function is based on whether the discriminator is fooled by the fake samples (i.e., classifies them as real), and the discriminator's loss is based on whether it can distinguish between real and fake samples (it should be mentioned that both models would use binary cross-entropy as their loss function).
+However, there are still several items that we haven't touched upon.
+
+Firstly, all three networks have been trained with teacher forcing, a procedure in which during training the model receives the ground truth $y^t$ as input at time $t+1$.
+Adopting this measure allows the model to create hidden-to-hidden connections and capture the necessary information from the history of input sequences \parencite{goodfellow_deep_2016}.
+
+Secondly, \textcite{goodfellow_generative_2014} mention a couple issues that can arise when training a GAN.
+One is synchronizing the generator with the discriminator (it's possible that one of the models gets too strong, leaving the other unable to catch up), and the generator collapsing too many values of $z$ to the same value of $x$ (which would ultimately hurt the generator's diversity; this issue is known in the literature as mode collapse).
+
+A solution to the issue of synchronicity would be to freeze the training of one of the models when the loss is below a predefined threshold \parencite{mogren_c-rnn-gan_2016}.
+In this paper, we stop training one of the models when its training loss is less than half the loss of the other's.
+
+To combat mode collapse, \textcite{salimans_improved_2016} propose a technique called feature matching, where we change the objective of the generator to generate real data that matches the statistics of the real data.
+Since in the setting of a GAN showing the real data would be considered "cheating", we use the features of an intermediate layer of the discriminator.
+Specifically, we train the generator to match the expected value of those features \parencite{salimans_improved_2016}.
+Denoting $f(x)$ as the activations on an intermediate layer in the discriminator, the new objective for the generator becomes
+\begin{equation}
+    \|\mathbb{E}_{x~p_{data}} f(x) - \mathbb{E}_{z~p_z (z)} f(G(z))\|_2^2
+\end{equation}
+which is commonly known as the squared error.
+However, to make the re-implementation of C-RNN-GAN true to the original, Mean Square Error (MSE) was used as the loss function when using feature matching.
+
+It should be noted that now, the loss functions of both the generator and discriminator differ in scale, meaning that we can no longer compare them to freeze training of either models: the generator uses MSE, whereas the discriminator use binary cross-entropy.
+A workaround to this is to calculate the both losses, but use MSE for backpropagation and optimization, whilst binary cross-entropy is used to compare our models.
+
+Moving forward, there is an aspect pertaining to the Reptile training framework which needs to be addressed: the training data.
+Normally, Reptile assumes that the training takes place in the $n$-shot $k$-way setting, meaning that our model is trained with $k$ classes and $n$ samples per class.
+Given that we have 954 samples in the training set, we adopted the same training procedure as \textcite{nichol_first-order_2018} and \textcite{clouatre_figr_2019}.
+Respectively, at each meta-loop we sample from our split $n$ examples for each $k$ class.
+For this paper, our $n=1$ and $k=5$ corresponding to a 5-way 1-shot experiment.
+
+Another aspect that needs addressing is the number of epochs.
+All models were trained for 250 epochs each however, when looking at the way training is handled in Reptile (see Algorithm \ref{alg:reptile}), this statement is inconclusive.
+Because the algorithm presents us with a meta-loop and an inner-loop, epochs is defined in this setting as the number of times parameters $\theta$ have been updated.
+Thus, the number of steps we have taken to compute $\tilde \theta$ is ignored.
+If there is skepticism about the fact that the Reptile models are trained much more than the baseline leading to better performances, the reader should be reminded by the fact that the latter is presented with 954 songs per epoch, whereas the former are trained with five.
+
+With these in mind, it should be mentioned that neither of the three models was presented with a full song during training, but rather with each song split into fixed-size chunks.
+This decision was made to better accommodate the memory limits of the hardware (an Nvidia T4 GPU).
+A full explanation of the limitations is found in Appendix \ref{appendix:memory}.
+It suffices to say that our dataset won't fit in memory.
+As such, each song was split into chunks of 200 events (i.e., window size of 200), starting from the beginning and moving towards the end of the song at a rate of 50 events per chunk (i.e, stride size of 50).
+Thus, a song of, say, 800 events would be transformed into 13 chunks, which leads us to the topic of batch size.
+
+At training, the Reptile models were presented with batches of 64 chunks, whereas the baseline was presented with batches of 2048 chunks.
+The rationale for the latter is that training of the baseline under a regimen of 64 batches would take approximately 250 hours until completion.
+
+Finally, another aspect that needs addressing is the learning rate.
+The baseline and Reptile inner-loop training was conducted with an $\alpha = 0.001$, value borrowed from \textcite{oore_this_2018}.
+Since Reptile also dictates a learning rate for the meta-loop, this value was set to $\beta = 0.01$.
+The reason for choosing this value comes from the Reptile experiments done by \textcite{nichol_first-order_2018}, where the meta-loop learning rate tends to be ten times higher than the inner rate.
+
+A summary of the parameters used in the training procedures is found in Appendix \ref{appendix:parameters}.
+
+\section{Evaluation} \label{sec:evaluation}
 
@@
 
-NDB is an evaluation metric specifically designed for generative models to measue the diversity of the generated samples.
+NDB is an evaluation metric specifically designed for generative models to measure the diversity of the generated samples.
 It follows the intuition that given two samples from the same distribution, the number of samples that fall into a bin (read as: class or cluster) should be the same \parencite{richardson_gans_2018}.
@@
 It should be noted that this is a restrictive metric: it can give a low score to music where the two notes start at different times.
-Scale consistency is the fraction of notes that are part of a standard scale (e.g., major, minor, lydian, etc.), and reporting the results for the best matching scale.
-Repetitions counts consecutive subsequences of notes, and tone span is the difference between the lowest note and highest note (counted in semi-tones) \parencite{mogren_c-rnn-gan_2016}.
+Scale consistency is the fraction of notes that are part of a standard scale (e.g., major, minor, lydian, etc.; see Appendix \ref{appendix:scales} for a complete list with examples), and reporting the results for the best matching scale.
+Repetitions counts consecutive subsequences of notes, and tone span is the difference between the lowest note and highest note \parencite[counted in semi-tones;][]{mogren_c-rnn-gan_2016}.
 
-The choice of NDB comes from the fact that it is good at detecting overfitting \parencite{borji_pros_2018}, whereas the domain-specific metrics would favor models that generate high fidelity samples.
+The choice of NDB comes from the fact that it is good at detecting overfitting \parencite{borji_pros_2018}, whereas the domain-specific metrics should favor models that generate high fidelity samples.
 In addition, these evaluation metrics have the benefit of being model-agnostic; they do not require a specific type of generative model.
+
+\section{Utilities}
 
@@
 
+%TC:ignore
+\appendix
+
+\chapter{Composer labelling} \label{appendix:labelling}
+
+\begin{center}
+    \tablefirsthead{\textbf{Canonical Composer} & \textbf{Genre} \\ \hline}
+\tablehead{\multicolumn{2}{l}{(continued)} \\
+\textbf{Canonical Composer} & \textbf{Genre} \\ \hline}
+\bottomcaption{Genres assigned to the unique entries of composers in the MAESTRO dataset.
+Where two composers are present, indicates a collaboration or an adaptation of the latter's work made by the former}
+\begin{supertabular}{p{0.75\linewidth}|p{0.25\linewidth}}
+Alban Berg & modernism \\ 
+Alexander Scriabin & romanticism \\ 
+Antonio Soler & baroque \\ 
+Carl Maria von Weber & classical \\ 
+Charles Gounod / Franz Liszt & classical \\ 
+Claude Debussy & impressionist \\ 
+César Franck & romanticism \\ 
+Domenico Scarlatti & baroque \\ 
+Edvard Grieg & romanticism \\ 
+Felix Mendelssohn & romanticism \\ 
+Felix Mendelssohn / Sergei Rachmaninoff & romanticism \\ 
+Franz Liszt & romanticism \\ 
+Franz Liszt / Camille Saint-Saëns & romanticism \\ 
+Franz Liszt / Vladimir Horowitz & romanticism \\ 
+Franz Schubert & classical \\ 
+Franz Schubert / Franz Liszt & classical \\ 
+Franz Schubert / Leopold Godowsky & classical \\ 
+Fritz Kreisler / Sergei Rachmaninoff & modernism \\ 
+Frédéric Chopin & romanticism \\ 
+George Enescu & modernism \\ 
+George Frideric Handel & baroque \\ 
+Georges Bizet / Ferruccio Busoni & romanticism \\ 
+Georges Bizet / Moritz Moszkowski & romanticism \\ 
+Georges Bizet / Vladimir Horowitz & romanticism \\ 
+Giuseppe Verdi / Franz Liszt & romanticism \\ 
+Henry Purcell & baroque \\ 
+Isaac Albéniz & romanticism \\ 
+Isaac Albéniz / Leopold Godowsky & romanticism \\ 
+Jean-Philippe Rameau & baroque \\ 
+Johann Christian Fischer / Wolfgang Amadeus Mozart & classical \\ 
+Johann Pachelbel & baroque \\ 
+Johann Sebastian Bach & baroque \\ 
+Johann Sebastian Bach / Egon Petri & baroque \\ 
+Johann Sebastian Bach / Ferruccio Busoni & baroque \\ 
+Johann Sebastian Bach / Franz Liszt & baroque \\ 
+Johann Sebastian Bach / Myra Hess & baroque \\ 
+Johann Strauss / Alfred Grünfeld & romanticism \\ 
+Johannes Brahms & romanticism \\ 
+Joseph Haydn & classical \\ 
+Leoš Janáček & romanticism \\ 
+Ludwig van Beethoven & classical \\ 
+Mikhail Glinka / Mily Balakirev & classical \\ 
+Mily Balakirev & romanticism \\ 
+Modest Mussorgsky & romanticism \\ 
+Muzio Clementi & classical \\ 
+Niccolò Paganini / Franz Liszt & classical \\ 
+Nikolai Medtner & modernism \\ 
+Nikolai Rimsky-Korsakov / Sergei Rachmaninoff & romanticism \\ 
+Pyotr Ilyich Tchaikovsky & romanticism \\ 
+Pyotr Ilyich Tchaikovsky / Mikhail Pletnev & romanticism \\ 
+Pyotr Ilyich Tchaikovsky / Sergei Rachmaninoff & romanticism \\ 
+Richard Wagner / Franz Liszt & romanticism \\ 
+Robert Schumann & romanticism \\ 
+Robert Schumann / Franz Liszt & romanticism \\ 
+Sergei Rachmaninoff & romanticism \\ 
+Sergei Rachmaninoff / György Cziffra & romanticism \\ 
+Sergei Rachmaninoff / Vyacheslav Gryaznov & romanticism \\ 
+Wolfgang Amadeus Mozart & classical \\ 
+\end{supertabular}
+\end{center}
+
+\chapter{Scales used in evaluation} \label{appendix:scales}
+
+All of the following examples have the note C as the root of the scale.
+\begin{figure}[ht]
+    \centering
+    \def\svgwidth{\linewidth}
+    \input{images/scales-1.pdf_tex}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \def\svgwidth{\linewidth}
+    \input{images/scales-2.pdf_tex}
+\end{figure}
+
+\chapter{Memory Limitations} \label{appendix:memory}
+
+The length of our songs varies from 2,100 to 202,046 events.
+Coupled with the fact that our model is presented with one-hot encoded vectors, this translates into matrices of sizes 2,100 $\times$ 416 up to 202,046 $\times$ 416.
+A better illustration would be to assume that all training songs have a length equal to our mean length (44,827) and that we are training just the baseline.
+For a batch size of 64, this means we are dealing with a tensor of size 64 $\times$ 44,827 $\times$ 416 elements, where one element is 4 bytes.
+This translates into a tensor that is 4.77 gigabytes.
+To add on top of that, our training dictates generating a second tensor that is equal in the number of elements.
+Furthermore, PyTorch \parencite{paszke_automatic_2017}, the machine learning library which is used in this paper, requires each element in our target tensor to be a double-precision floating point number - i.e., 8 bytes per element.
+By now we have exhausted the entire memory (12.5 gigabytes in Google Colaboratory) of our GPU leaving no space for any of our networks.
+
+Granted, someone could make the argument to lower the batch size of the songs to 32.
+However, due to high variance in the length of our songs, we could risk a situation were all of the songs have a length of 100,000 and up (there are 68 such songs in the training set) leading to a tensor that is 5.32 gigabytes (single-precision floating point).
+
+A batch size of 16 would be a more compelling alternative: 2.66 GB single-precision floating point with a 5.32 GB double-precision floating leaves 4.52 GB of space.
+However, if we account the size of the libraries imported into the training script and the hidden state tensors (batch size $\times$ vocabulary size) puts us dangerously close to the memory limits of the GPU.
+
+\chapter{Parameters} \label{appendix:parameters}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{l|c|c|c}
+                    &\bf Baseline   &\bf Performance RNN&\bf C-RNN-GAN  \\ \hline
+    Num. Layers     &   1           &   3               &   2           \\
+    Num. Units      &   256         &   512             &   350         \\
+    Dropout Rate    &   0.0         &   0.3             &   0.3         \\
+    Learning Rate   &   0.001       &   0.001           &   0.001       \\
+    Meta Rate       &   N/A         &   0.01            &   0.01        \\
+    Batch Size      &   2048        &   64              &   64          \\
+    Meta Epochs     &   250         &   250             &   250         \\
+    Inner Epochs    &   N/A         &   3               &   3           \\
+    Has bias vector &   No          &   Yes             &   Yes         \\
+    Gradient Clipped&   No          &   Yes             &   Yes         \\
+    \end{tabular}
+    \caption{Summary of parameters used to construct and train the three models featured in this paper}
+    \label{tab:params}
+\end{table}
+
+%TC:endignore
 \end{document}

### 2019-05-05 forked_src/c_rnn_gan.py

@@ -1,0 +1,237 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
+
+
+class Generator(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=416,
+                 dropout=0.3,
+                 num_layers=2,
+                 init_dim=100,
+                 use_bias=False):
+        super(Generator, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = int(out_dim)
+        self.dropout = dropout
+        self.init_dim = init_dim
+        self.use_bias = use_bias
+        self.num_layers = num_layers
+        self.is_bidirectional = False
+
+        if self.is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=self.use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=self.dropout,
+            bidirectional=False)
+
+        self.out_layer = nn.Linear(self.hidden_dim * self.num_layers,
+                                   self.out_dim)
+        self.out_activation = nn.Softmax(dim=-1)
+
+        self._init_weights()
+
+    def _init_weights(self):
+        nn.init.xavier_normal_(self.event_embedding.weight)
+
+        nn.init.xavier_normal_(self.init_layer.weight)
+        # self.init_layer.bias.fill_(0.0)
+
[…]

### 2019-05-05 forked_src/data.py

@@ -1,0 +1,618 @@
+import os
+from time import strftime, time
+
+import numpy as np
+import pandas as pd
+import pretty_midi
+import mido
+import utils
+
+from mido import MetaMessage, Message, MidiTrack, MidiFile
+
+
+def to_seq(path):
+    midi = mido.MidiFile(path)
+    tpb = midi.ticks_per_beat  # Ticks/beat. between 384-480
+    tempo = midi.tracks[0][0].tempo  # Tempo in microseconds. All are 500000
+    # max_tick = mido.second2tick(second=1, ticks_per_beat=tpb, tempo=tempo)
+
+    melody = midi.tracks[1]
+
+    sequence = []
+    for msg in melody:
+        if hasattr(msg, 'note'):
+            # move_by = int(round(msg.time/8, 0)*8)
+            # Convert from ticks to miliseconds
+            move_by = int(mido.tick2second(msg.time, tpb, tempo) * 1000)
+            # sort of quantizing by 8ms
+            move_by = int((move_by // 8) * 8)
+
+            # If the time between the previous note and
+            #  the current note is greater than 1 second,
+            #  force time difference to be 1 second.
+            #  1 second at 120 BPM (default) == 2 bars
+            #  Therefore, @var:move_by will still be in time
+            if move_by > 1000:
+                move_by = 1000
+            # Append time since last message
+            sequence.append('time_move:{}'.format(move_by))
+            # Append velocity; we limit the number of possible
+            #  velocity values to 32 instead of 128
+            #  When decoding velocity must be multiplied by 4
+            #  get back the appropriate values
+            sequence.append('velocity_set:{}'.format(msg.velocity // 4))
+
+            # Append `note_on` or `note_off`
+            # The data has this weird quirk where it
+            #  does not make use of the note_off
+            #  messsage. What it does, is that it sets
+            #  the velocity of the note to be closed
+            #  to zero. Thus, a messages with 'note_on'
+            #  note=60, velocity=0, actually means
+            #  'note_off', note=60, velocity=0
+            # Thus, we store the message type in a
+            #  variable, and check the velocity. If
+            #  it's zero, then we change the message
+            #  to note_off
+            message_type = msg.type
+            if msg.velocity == 0:
+                message_type = 'note_off'
+            sequence.append(message_type + ':{}'.format(msg.note))
[…]

### 2019-05-05 forked_src/evaluation.py

@@ -192,22 +192,16 @@ def _prepare_data(paths, max_len=None):
         max_len = max(length)
 
     sequences = []
-    for i in data:
+    for i in tqdm(data, desc='Padding data'):
         diff = max_len - i.shape[0]
         if diff > 0:
             i = np.pad(i, pad_width=((0, diff)), mode='constant')
         elif diff < 0:
             i = i[:max_len]
 
-        try:
-            sequences.append(i)
-        except ValueError:
-            continue
+        sequences.append(i)
 
-    try:
-        return np.vstack(sequences), max_len
-    except ValueError:
-        return None, None
+    return np.vstack(sequences), max_len
 
 
 def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
@@ -277,67 +271,3 @@ def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
     })
     df.index.name = 'bins'
     return df
-
-
-if __name__ == '__main__':
-    import os
-    from glob import glob
-
-    rng = np.random.RandomState(1337)
-
-    GEN_DIR = '../generated/'
-    test_df = pd.read_csv('../data/maestro_test.csv')
-    test_paths = [
-        os.path.join('../data/', i) for i in test_df['midi_filename']
-    ]
-    test_data, max_len = _prepare_data(test_paths)
-
-    models = [i[1] for i in os.walk(GEN_DIR)][0]  # get model folders
-    for model in models:
-        # get parameter folders
-        params = [i[1] for i in os.walk(os.path.join(GEN_DIR, model))][0]
-
-        for param in params:
-            current_folder = os.path.join('../generated', model, param)
-            paths = glob(current_folder + '/*.midi')
-            gen_data, _ = _prepare_data(paths, max_len)
-            if gen_data is None:
-                continue
-
-            results1 = ndb(test_data, gen_data, n_bins=20, rng=rng)
-            results2 = pd.DataFrame(evaluate_batch(paths))
-
-            results1.to_csv(current_folder + '/ndb.csv')
-            results2.to_csv(current_folder + '/domain_evaluation.csv')
-
-    results = {
-        'model': [],
-        'polyphony': [],
-        'repetitions': [],
-        'tone_span': [],
-        'scale_consistency': [],
-        'ndb': []
-    }
-
-    for model in models:
-        # get parameter folders
-        params = [i[1] for i in os.walk(os.path.join(GEN_DIR, model))][0]
-
-        for param in params:
[…]

### 2019-05-05 forked_src/flags.py

@@ -1,0 +1,102 @@
+import argparse
+
+
+class Flags:
+    def __init__(self):
+        self.parser = argparse.ArgumentParser(
+            formatter_class=argparse.ArgumentDefaultsHelpFormatter)
+        self.parser.add_argument(
+            '--model-name',
+            help="""name of the model. can be either of the following: 
+            'baseline', 'performance_rnn', 'c_rnn_gan'""",
+            default='baseline',
+            type=str)
+        self.parser.add_argument(
+            '--raw_datadir',
+            help='folder where the midi data is',
+            default='../data/',
+            type=str)
+        self.parser.add_argument(
+            '--proc_datadir',
+            help='folder of the processed midi',
+            default='../data/processed/',
+            type=str)  # DEPRECATED?
+        self.parser.add_argument(
+            '--model-folder',
+            help='folder where the model will be saved',
+            default='../models/',
+            type=str)
+        self.parser.add_argument(
+            '--train-file',
+            help='name of the file as found in the `data` folder',
+            default='CHANGE_ME',
+            type=str)
+        self.parser.add_argument(
+            '--test-file',
+            help='name of the file as found in the `data` folder',
+            default='CHANGE_ME',
+            type=str)
+        self.parser.add_argument(
+            '--layer-size',
+            help='size of the hidden units in the LSTM network',
+            default=256,
+            type=int)
+        self.parser.add_argument(
+            '--dropout',
+            help='rate of dropping out units in a network',
+            default=0.0,
+            type=float)
+        self.parser.add_argument(
+            '--n-classes',
+            help='number of classes per inner task',
+            default=5,
+            type=int)
+        self.parser.add_argument(
+            '--n-shots',
+            help='number of examples per class',
+            default=1,
+            type=int)
+        self.parser.add_argument(
+            '--inner-batch', help='inner batch size', default=64, type=int)
+        self.parser.add_argument(
+            '--inner-iters', help='inner iterations', default=5, type=int)
+        self.parser.add_argument(
+            '--learning-rate',
+            help='Adam step size',
+            default=0.0025,
+            type=float)
+        self.parser.add_argument(
+            '--meta-step',
+            help='meta-training step size',
+            default=0.025,
+            type=float)
+        self.parser.add_argument(
+            '--meta-step-final',
+            help='meta-training step size',
+            default=0.0,
+            type=float)
+        self.parser.add_argument(
[…]

### 2019-05-05 forked_src/lstm_baseline.py

@@ -1,0 +1,335 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
+
+
+class Baseline(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=416,
+                 dropout=0.0,
+                 num_layers=1,
+                 init_dim=32,
+                 use_bias=False,
+                 is_bidirectional=False):
+        super(Baseline, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = out_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+
+        if is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=float(dropout),
+            bidirectional=is_bidirectional)
+
+        self.out_layer = nn.Linear(self.hidden_dim * self.num_layers,
+                                   self.out_dim)
+        self.out_activation = nn.Softmax(dim=-1)
+
+        self._init_weights()
+
+    def _init_weights(self):
+        nn.init.xavier_normal_(self.event_embedding.weight)
+
+        nn.init.xavier_normal_(self.init_layer.weight)
[…]

### 2019-05-05 forked_src/old/train_c_rnn_gan.py

@@ -2,7 +2,6 @@ import torch
 import torch.nn as nn
 import torch.optim as optim
 import torch.nn.functional as F
-from torch.distributions import Categorical
 
 import os
 import data
@@ -24,8 +23,7 @@ class ReptileGAN:
                  rng,
                  raw_datadir,
                  split,
-                 in_dim=416,
-                 out_dim=416,
+                 in_dim=415,
                  hidden_dim=512,
                  init_dim=100,
                  num_layers=2,
@@ -47,7 +45,6 @@ class ReptileGAN:
                  max_norm=1.0,
                  g_train=True,
                  d_train=True,
-                 feature_matching=False,
                  save_path='../models/',
                  g_model_state=None,
                  d_model_state=None,
@@ -63,7 +60,6 @@ class ReptileGAN:
         self.raw_datadir = raw_datadir
         self.split = split
         self.in_dim = in_dim
-        self.out_dim = out_dim
         self.hidden_dim = hidden_dim
         self.init_dim = init_dim
         self.num_layers = num_layers
@@ -83,9 +79,8 @@ class ReptileGAN:
         self.clip_grad = clip_grad
         self.clip_norm = clip_norm
         self.max_norm = max_norm
-        self.g_train = g_train
-        self.d_train = d_train
-        self.feature_matching = feature_matching
+        self.g_train = g_train,
+        self.d_train = d_train,
         self.save_path = save_path
         self.g_loss_saved = g_loss_saved
         self.d_loss_saved = d_loss_saved
@@ -125,7 +120,6 @@ class ReptileGAN:
     def init_models(self):
         self.g = Generator(
             in_dim=self.in_dim,
-            out_dim=self.out_dim,
             hidden_dim=self.hidden_dim,
             batch_size=self.batch_size,
             dropout=self.dropout,
@@ -134,7 +128,6 @@ class ReptileGAN:
             use_bias=self.use_bias).to(self.device)
         self.meta_g = Generator(
             in_dim=self.in_dim,
-            out_dim=self.out_dim,
             hidden_dim=self.hidden_dim,
             batch_size=self.batch_size,
             dropout=self.dropout,
@@ -198,7 +191,6 @@ class ReptileGAN:
             {
                 'split': self.split,
                 'in_dim': self.in_dim,
-                'out_dim': self.out_dim,
                 'hidden_dim': self.hidden_dim,
                 'init_dim': self.init_dim,
                 'num_layers': self.num_layers,
@@ -221,7 +213,6 @@ class ReptileGAN:
[…]

### 2019-05-05 forked_src/performance_rnn.py

@@ -1,0 +1,767 @@
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.distributions import Categorical
+
+import os
+import data
+import utils
+import numpy as np
+import pandas as pd
+from copy import deepcopy
+from tqdm import tqdm
+
+import pdb
+
+
+class PerformanceRNN(nn.Module):
+    def __init__(self,
+                 in_dim,
+                 hidden_dim,
+                 batch_size,
+                 out_dim=416,
+                 dropout=0.0,
+                 num_layers=1,
+                 init_dim=32,
+                 use_bias=False,
+                 is_bidirectional=False):
+        super(PerformanceRNN, self).__init__()
+
+        self.device = torch.device('cuda:0' if torch.cuda.
+                                   is_available() else 'cpu')
+
+        self.in_dim = in_dim  # dimension of the event
+        self.hidden_dim = hidden_dim
+        self.batch_size = batch_size
+        self.out_dim = out_dim
+        self.init_dim = init_dim
+        self.num_layers = num_layers
+
+        if is_bidirectional:
+            self.num_directions = 2
+        else:
+            self.num_directions = 1
+
+        self.primary_event = self.out_dim - 1
+
+        self.event_embedding = nn.Embedding(self.out_dim, self.out_dim)
+
+        self.init_layer = nn.Linear(
+            self.init_dim,
+            int(self.num_directions * self.num_layers * self.hidden_dim))
+        self.init_activation = nn.Tanh()
+
+        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
+        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+
+        self.lstm = nn.LSTM(
+            input_size=self.hidden_dim,
+            hidden_size=self.hidden_dim,
+            num_layers=int(self.num_layers),
+            bias=use_bias,
+            batch_first=False,  # if true input is (batch, seq, feature)
+            dropout=float(dropout),
+            bidirectional=is_bidirectional)
+
+        self.out_layer = nn.Linear(self.hidden_dim * self.num_layers,
+                                   self.out_dim)
+        self.out_activation = nn.Softmax(dim=-1)
+
+        self._init_weights()
+
+    def _init_weights(self):
+        nn.init.xavier_normal_(self.event_embedding.weight)
+
+        nn.init.xavier_normal_(self.init_layer.weight)
[…]

### 2019-05-05 forked_src/reptile.py

@@ -1,0 +1,160 @@
+from copy import deepcopy
+from utils import read_npy, make_folder, now
+from data import sample_mini_dataset, mini_batches
+from sklearn.preprocessing import LabelEncoder
+from model import CNN
+from torch.autograd import Variable
+
+import numpy as np
+import pandas as pd
+import torch
+import torch.nn.functional as F
+
+import pdb
+
+RNG = np.random.RandomState(1337)
+
+
+def inner_training_step(net, inner_optim, batch):
+
+    x, y = zip(*batch)
+    x = np.vstack([read_npy(i).transpose(0, 3, 1, 2) for i in x])
+    encoder = LabelEncoder()
+    y = encoder.fit_transform(np.array(y).reshape(-1))
+
+    x = torch.from_numpy(x)
+    x.requires_grad_(True)
+    y = torch.from_numpy(y).long()
+    y.requires_grad_(False)
+
+    net.train()
+    inner_optim.zero_grad()
+
+    loss = F.cross_entropy(net(x), y)
+    loss.backward()
+
+    inner_optim.step()
+
+
+def meta_training_step(train_set, model,
+                       optimizer,
+                       num_shots, num_classes, inner_batch_size,
+                       inner_iters, inner_step_size, meta_step_size, meta_batch_size):
+
+    weights_original = deepcopy(model.state_dict())
+    new_weights = []
+
+    for _ in range(meta_batch_size):
+        mini_dataset = sample_mini_dataset(train_set, num_classes,
+                                           num_shots)
+
+        n_batch = 1
+        for j in range(inner_iters):
+            for batch in mini_batches(mini_dataset, inner_batch_size,
+                                      False):
+                print('[ ] {} Batch-wise Inner Epoch: {}'.format(now(), n_batch))
+                # inner_optim = torch.optim.Adam(model.parameters(), lr=inner_step_size)
+                inner_training_step(model, optimizer, batch)
+                n_batch += 1
+
+        new_weights.append(deepcopy(model.state_dict()))
+        model.load_state_dict({name: weights_original[name]
+                               for name in weights_original})
+
+    ws = len(new_weights)
+    fweights = {name: torch.zeros(new_weights[0][name].size())
+                for name in new_weights[0]}
+    print('sanity check {}'.format(ws))
+
+    for i in range(ws):
+        for name in new_weights[i]:
+            fweights[name] += new_weights[i][name]/float(ws)
+            print('diff {}'.format(weights_original[name].sum() - fweights[name].sum()))
+
[…]

### 2019-05-05 forked_src/run.py

@@ -1,0 +1,268 @@
+import os
+from pathlib import Path
+import json
+import data
+import utils
+import numpy as np
+import pandas as pd
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+
+with open('../params/test.json', 'r') as f:
+    FLAGS = json.load(f)
+
+RNG = np.random.RandomState(666)
+VOCAB = data.build_vocab()
+
+assert FLAGS['model_name'] in ['baseline', 'performance_rnn',
+                               'c_rnn_gan'], 'Invalid model name.'
+
+test_df = pd.read_csv('../data/maestro_test.csv')
+train_df = pd.read_csv('../data/maestro_train.csv')
+val_df = pd.read_csv('../data/maestro_validation.csv')
+
+utils.make_folder(FLAGS['model_folder'])
+
+if FLAGS['model_name'] == 'baseline':
+    from lstm_baseline import train_baseline_large
+
+    checkpoint = Path(
+        os.path.join(FLAGS['model_folder'],
+                     FLAGS['model_name'] + '_checkpoint'))
+
+    if checkpoint.is_file():
+        print(f'[!] {utils.now()} Loading checkpoint')
+        saved = torch.load(
+            os.path.join(FLAGS['model_folder'],
+                         FLAGS['model_name'] + '_checkpoint'),
+            map_location='cpu')
+        for i in saved.keys():
+            FLAGS[i] = saved[i]
+        losses, grads = train_baseline_large(
+            VOCAB,
+            RNG,
+            train_df,
+            train_df['seq_len'].tolist(),
+            in_dim=len(VOCAB),
+            out_dim=len(VOCAB),
+            hidden_dim=FLAGS['hidden_dim'],
+            init_dim=FLAGS['init_dim'],
+            num_layers=FLAGS['num_layers'],
+            batch_size=FLAGS['batch_size'],
+            meta_iters=FLAGS['meta_iters'],
+            meta_iters_start=FLAGS['meta_iters_start'],
+            learning_rate=FLAGS['meta_step'],
+            dropout=FLAGS['dropout'],
+            window_size=FLAGS['window_size'],
+            stride_size=FLAGS['stride_size'],
+            use_bias=FLAGS['use_bias'],
+            is_bidirectional=FLAGS['is_bidirectional'],
+            clip_grad=FLAGS['clip_grad'],
[…]

### 2019-05-08 paper/main.tex
disable feature matching + better loss in GAN

@@
 \usepackage{subcaption}
+\usepackage{lscape}
 \usepackage[ruled]{algorithm2e}
@@
 \newcommand{\empt}[2]{$#1^{\langle #2 \rangle}$}
+
+%% Drawing plots from PGFPlots
+\usepackage{pgfplots}
+\pgfplotsset{compat=newest}
+\usepgfplotslibrary{groupplots}
+\usepgfplotslibrary{dateplot}
 
@@
     prod/.style={circle, draw, thick, inner sep=0pt},
-    ct/.style={circle, draw, inner sep=5pt, ultra thick,
-		minimum width=10mm},
+    ct/.style={circle, draw, inner sep=0pt, ultra thick,
+		minimum size=3.0em, node distance=4em},
     ft/.style={circle, draw, thick, minimum width=8mm, inner sep=1pt},
@@
         Dr. Menno van Zaanen\\
-        Dr. SECOND READER\\
+        Dr. Martin Atzmueller\\
         \vfill
@@
 
-\chapter{The return of the Neural Net}
-
-Over the past sixty years, Artificial Neural Networks (ANNs) have experienced a number of popularity cycles that were noted by extensive publicity of over-inflated expectations with regards to the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs [CITATION NEEDED]) [CITATION NEEDED].
+\chapter{A hazy shade of winter}
+%Simon and Garfunkel - A hazy shade of winter
+
+Over the past sixty years, Artificial Neural Networks (ANNs) have experienced several popularity cycles noted in the literature by extensive publicity of over-inflated expectations with regards to the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs) \parencite{minsky_perceptrons_1988, knight_ai_2016, nilsson_speed_2009}.
 More specifically, it promised that a system could learn to do anything it was programmed to do \parencite{minsky_perceptrons_1988}.
-However, each wave of praise was followed by a mass of public disappointment, criticism, and funding cuts [CITATION NEEDED].
+However, each wave of praise was followed by a mass of public disappointment, criticism, and funding cuts \parencite[e.g.,][]{lighthill_artificial_1972}.
 These upturns and downturns of the field of connectionism - or more generally A.I.
 - are now known in the literature as the A.I.
-winters [CITATION NEEDED].
-
-The recent revival of ANNs under the scope of Deep Learning (DL) has brought new ways to tackle Machine Learning problems, most notably in the context of classification (but also regression, although a less popular application; see \cite{lecun_deep_2015} for several DL examples).
+winters \parencite{nilsson_speed_2009}.
+
+The recent revival of ANNs under the brand of Deep Learning (DL) has brought new ways to tackle Machine Learning problems - most notably in the context of classification (but also regression, although less popular; see \cite{lecun_deep_2015} for several examples).
 In addition, with the advent of chatbots \parencite{dale_return_2016}, voice-activated personal assistants \parencite{xiong_microsoft_2018}, and general-purpose A.I.
@@
 
-The reasons for presenting these facts will be clarified in the following sections (Section \ref{sec:thesis_goal} provides a concrete explanation) however, a terse argumentation is that progress towards general A.I.
-systems is marked by scant explorations of ANNs outside of the classification task, and even more so when trying to overcome some of their shortcomings, such as being highly reliant on massive amounts of data [CITATION NEEDED].
+The reasons for presenting these facts will be clarified bit-by-bit throughout this chapter (Section \ref{sec:thesis_goal} provides a concrete explanation) however, a terse argumentation is that progress towards general A.I.
+systems is marked by scant explorations of ANNs outside of the classification task, and even more so when trying to overcome some of their shortcomings, such as being highly reliant on massive amounts of data.
 % NOTE: Maybe I was too harsh?
 
-\section{We do not discriminate}\label{sec:no_disc}
-
-Typically, in a neural network, the data samples go through a set of input units (that might represent pixels, word embeddings, etc.), then through multiple hidden layers, each with a given number of nodes, and reaching the output units where the answer is given \parencite{marcus_deep_2018}.
-An overwhelming majority of experiments with ANNs involve discriminative models \parencite{goodfellow_generative_2014}, where the goal is to assign the high-dimensional inputs to a class label \parencite[such as an animal or piece of furniture;][]{zhang_character-level_2015, krizhevsky_imagenet_2012}.
-However, other possible applications of DL models include (but are not limited to): regression [CITATION NEEDED], data compression \parencite{cheng_deep_2018}, and generating new data points \parencite{graves_generating_2013}.
-
-For the purposes of this paper (detailed in Section \ref{sec:thesis_goal}), generative models are a central element in our experiments (Chapter \ref{chap:methods}).
+\section{Strange Fruit}\label{sec:no_disc}
+% Billie Holiday - Strange Fruit
+
+Typically, in a neural network, data samples go through a set of input units (that might represent pixels, word embeddings, etc.), then through multiple hidden layers, each with a given number of nodes, and reaching the output units where the answer is given \parencite{marcus_deep_2018}.
+An overwhelming majority of experiments with ANNs involve discriminative models \parencite{goodfellow_generative_2014}, where the goal is to assign high-dimensional inputs to a single class label \parencite[such as an animal or piece of furniture;][]{zhang_character-level_2015, krizhevsky_imagenet_2012}.
+However, other possible applications of DL models include (but are not limited to): regression \parencite{qiu_ensemble_2014}, data compression \parencite{cheng_deep_2018}, and generating new data points \parencite{graves_generating_2013}.
+
+The last point mentioned above is particularly interesting not because it has gained a lot of traction in recent years, but because of its potential to assist the scientific community as well as the general population.
 Borrowing the analogy from \textcite{goodfellow_nips_2016}, one might wonder what is the value of studying generative models.
-After all, in the image domain, such a framework would merely generate more images (something which the Internet as no shortage of).
-The answer to this lies in the fact that generative models can be used for simulating possible futures for scheduling and planning [CITATION NEEDED], providing predictions to missing data [CITATION NEEDED], or enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.
-
-However, successes in generative models have been scarce, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run.
-Yet, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$).
+If, for instance, we are dealing with images, such a framework would merely generate more images for us (something which the Internet as no shortage of).
+This sort of worldview, however, would only limit these systems from their full capacity; aside from the obvious use case, generative networks could also be used for simulating possible futures - such as for scheduling and planning \parencite{finn_deep_2016} -, for providing predictions to missing data, or for enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.
+
+Regardless, successes in generative models have been either scarce or unattractive, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run.
+Given the age and - alleged - popularity, a valid question to ask is why things haven't moved that much since then.
+As it turn out, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$).
 Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}.
-Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task.
+This discussion is not exhaustive; for a comprehensive list of various generative approaches and their downsides, refer to \textcite{goodfellow_nips_2016}.
+%Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task.
 Their main drawback however, is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.
@@
 The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this field.
-Besides countering the drawbacks outlined above \parencite{goodfellow_nips_2016}, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.
-
-In addition, although not as popular as the abovementioned, Recurrent Neural Networks (RNNs) have been successfully implemented for various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}.
-More specifically, a variation of these, the Long Short-Term Memory Networks \parencite[LSTMs;][]{hochreiter_long_1997}, are of particular interest.
-RNNs (and by extension LSTMs) contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}.
-This makes them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc.
+Besides countering the drawbacks outlined above, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.
+
+Similarly, and although not as popular, Recurrent Neural Networks (RNNs) have been successful at various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}.
+The reason is because RNNs contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}.
+This made them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc.
 \parencite{fan_tts_2014}.
-That being said, the purpose-built memory cell of an LSTM permits better storing and accessing of information \parencite{hochreiter_long_1997, graves_generating_2013} which in turn makes this RNN variation a better candidate for this sort of task.
+That being said, Long Short-Term Memory networks \parencite[LSTM;][]{hochreiter_long_1997} - a variant of RNNs - have been purpose-built with memory cells which permit better storing and accessing of information.
+This design decision variant of the recurrent network good candidates for generating sequences with long-term dependencies \parencite{graves_generating_2013}.
 
@@
 Herein lies the scope of this paper.
-We evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and LSTM \parencite{oore_this_2018} when trained under a meta-learning algorithm (see Section \ref{sec:hungry}).
+Given such minimal explorations into generating music, we evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and a LSTM network \parencite{oore_this_2018} trained under the constraints imposed by few-shot learning (Section \ref{sec:hungry}) against a baseline model proposed by \textcite{larochelle_few-shot_2017}.
 To reiterate, a full explanation of the purpose of the paper is provided in Section \ref{sec:thesis_goal}.
 
-\section{We always want more}\label{sec:hungry}
+\section{I want it all}\label{sec:hungry}
+% Queen - I want it all
 
@@
 
-A clear outcome from solving these issues would be that DL models are easier to train (less data) and multifunctional however, this list is not exhaustive.
+A clear outcome from solving these issues would be that DL models are easier to train (less data) as well as multifunctional.
+Nonetheless, this list is not exhaustive.
 There is also fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack.
 Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class.
-Bridging the gap between human-type learning and current learning architectures would allow models to properly capture this sort of sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}.
+Bridging the gap between human-type learning and current learning architectures would allow a system to do a proper job at capturing and generalizing information from sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}.
 This has motivated a series of explorations in the direction of "few-shot" learning (i.e., learning a class from a few examples) or more broadly, into the field of "meta-learning" \parencite[learning to learn in hopes of generalizing to new tasks;][]{chen_closer_2018, vinyals_matching_2016, zhang_metagan_2018}.
@@
 Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}.
-Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early explorers in this direction.
+Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early (and few) explorers in this direction.
 
@@
 
-At the time of writing, generative models and meta-learning are two exciting areas of research in DL which have been successfully combined.
-However, to the knowledge of the researcher, no such experiments have been conducted in the musical domain.
+At the time of writing, generative models and meta-learning are two exciting areas of research the intersection of which could be quite promising.
+However, to the knowledge of the researcher, no experiments which combine the both of them have been conducted in the musical domain.
 Thus, this paper examined the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning.
-In other words, it sought to evaluate the extent to which generative models can create novel and high-fidelity samples when training data is scarce.
-To avoid ambiguity, we define novel as new (i.e., not in the training set), and qualitative as consistent with common scales and diverse (combines chord structures with single-note sequences, is not restricted to only a few notes, etc.).
-This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, on a limited number of piano recordings.
+In other words, it sought to evaluate the degree to which generative models can create novel and high-fidelity samples when training data is scarce.
+To avoid any ambiguity, novel is used here as new (i.e., not in the training set), and qualitative as consistent with common scales and structurally diverse (combines chord structures with single-note sequences and is not restricted to only a few notes).
+
+This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, with a limited number of songs.
 The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.
@@
 The following chapters will discuss more about generative models, meta-learning, and some of the models and algorithms involved in this paper (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the setup of the experiments (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}).
-Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} draws some conclusions that answer the above research questions.
+Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} concludes the findings of this paper.
 
@@
 Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network.
-Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning, several approaches to it, and finally narrow down to current state-of-the-art algorithms.
+Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning and some state-of-the-art approaches.
 
@@
 
-In a standard ANN, the data traverses the network through the input layer, to the hidden layer, and finally reaches the output layer; setup which is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}.
-Broadly, this type of network is defined by the fact that its connections do not form cycles.
+In a standard ANN, the data traverses the network through the input layer, to the hidden layers, and finally the output layer; this setup is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}.
+Broadly speaking, this type of network is defined by the fact that its connections do not form cycles.
 If however we relax this condition, we arrive at a recurrent neural network \parencite[Figure \ref{fig:rnn_folded};][]{rumelhart_learning_1986}: a family of neural networks specifically designed for processing sequential data \parencite{goodfellow_deep_2016}.
-It is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution would be the basis for building a time-delay neural network.
-Unfortunately, the output of such a network would be a function of a small number of neighbouring data points \parencite[say, the last three notes;][]{goodfellow_deep_2016} whereas in the contex of music, knowledge of the notes played from the begining of the song is needed.
-As such, RNNs seem to be a more compelling alternative for this task.
-
-The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and prior hidden activations from the previous timestep \parencite{graves_supervised_2012}.
-Moreover, recurrent networks also benefit from the fact that parameters (weights) are shared across different parts of a model.
-This allows extending and applying the model to data points of different lengths and generalizing across all of them \parencite{goodfellow_deep_2016}.
-In contrast, a feedforward network would require separate parameters for each input feature, at each position.
+
+For transparency, it should be noted that it is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution forms the basis of a time-delay neural network.
+Unfortunately, the output of such a system would be a function of a small number of neighbouring data points \parencite[such as the last and the next three notes in a song;][]{goodfellow_deep_2016}.
+Despite that, a system that uses convolution would quickly reach its limitation in the musical domain.
+Here, knowledge of the notes played from the beginning of the song is needed, meaning that we need an architecture that is able to account for long-term dependencies - which brings us back to the RNN.
+
+The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and hidden activations from the previous timestep.
+It should be noted that the parameters of a recurrent network (weights) are shared across different timesteps \parencite{graves_supervised_2012}.
+This allows extending (and applying) the model to data points of different lengths whilst generalizing across all of them \parencite{goodfellow_deep_2016}.
+To place this into perspective, a feedforward network would require separate parameters for each input feature, at each position.
+
 Thus, a formal definition of the network's forward pass would be
@@
 In this setting, a popular algorithm is called backpropagation through time \parencite[BPTT;][]{werbos_backpropagation_1990} which consists of repeated applications of the chain rule.
-The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer a the next timestep \parencite{graves_supervised_2012}.
+The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer at the next timestep \parencite{graves_supervised_2012}.
 
@@
 As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}.
-\textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing bacwards in time tend to either (1) blow up or (2) vanish.
+\textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing backwards in time tend to either (1) blow up or (2) vanish.
 As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.
@@
 In the context of generating musical performances, experiments with this neural architecture go as far as \textcite{eck_first_2002} however, a more noteworthy implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 cells each designed to generate classical compositions whilst maintaining the playing style of expert pianists.
-Considering how successful it was, this model will be re-implemented in this paper's experiments (see Chapter \ref{chap:methods}).
+Considering its success, this model was re-implemented for the purposes of this paper's experiments (see Chapter \ref{chap:methods}).
 
@@
 
-                            \draw[<-, connl] (fit.west|-int2) coordinate (aux)--++(180:7mm) node[left]{$x^t$};
+                            \draw[<-, connl] (fit.west|-int2) coordinate (aux)--++(180:7mm) node[left, io]{$x^t$};
                             \draw[<-, connl] ([yshift=1mm]aux)--++(135:7mm);
@@
 
-                            \draw[<-, connl] (fit.north-|it) coordinate (aux)--++(90:7mm) node[above]{$x^t$};
+                            \draw[<-, connl] (fit.north-|it) coordinate (aux)--++(90:7mm) node[above, io]{$x^t$};
                             \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
@@
 
-                            \draw[<-, connl] (fit.north-|ot) coordinate (aux)--++(90:7mm) node[above]{$x^t$};
+                            \draw[<-, connl] (fit.north-|ot) coordinate (aux)--++(90:7mm) node[above, io]{$x^t$};
                             \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
@@
 
-                            \draw[<-, connl] (fit.south-|ft) coordinate (aux)--++(-90:7mm) node[below]{$x^t$};
+                            \draw[<-, connl] (fit.south-|ft) coordinate (aux)--++(-90:7mm) node[below, io]{$x^t$};
                             \draw[<-, connl] ([xshift=1mm]aux)--++(-45:7mm);
@@
 To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$.
-The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar where $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}.
-We train $D$ to maximize the probability of correctly assigning a label to the training data and samples from $G$ ("real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, without actually looking at $p_{\text{data}}$.
+The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar.
+More concretely, $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}.
+We train $D$ to maximize the probability of correctly classifying the training data and samples from $G$ (i.e., as "real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, \emph{without actually looking at} $p_{\text{data}}$.
 In other words, $G$ must minimize $\log (1- D(G(z)))$ \parencite{goodfellow_generative_2014}.
 
-In other words, both models play a minimax game with a value function $V(G, D)$:
+To formalize, both models play a minimax game with a value function $V(G, D)$:
 \begin{equation}
@@
 As mentioned before, they have been proven successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}.
-Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by re-purposing this architecture to generate music.
+Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by successfully re-purposing this architecture to generate music.
 The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a unidirectional LSTM network with a similar structure as discriminator.
@@
 To be more concrete, a human is able to generalize a task/object after encountering one (or few) example(s) \parencite{ravi_optimization_2016, chen_closer_2018}.
-Therefore, besides having to perform on a variety of tasks, a systems would need to excel at that using sparse amounts of data.
-
-Although ideas of meta-learning can go as far as Schmidhuber's thesis in 1987, it seems that training a network using a handful of examples did not get much traction until \textcite{lake_human-level_2015} successfully designed a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
+Therefore, besides having to perform on a variety of tasks, a system would also need to excel at that using sparse amounts of data.
+
+Although ideas of meta-learning can go as far as Schmidhuber's thesis in 1987, it seems that training a model using a handful of examples did not get much traction until \textcite{lake_human-level_2015} successfully designed a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
 For clarity, such a problem is typically defined as a $n$-shot $k$-way problem, where $n$ stands for the number of examples per class, and $k$ is the number of classes to be used in the training process.
 
-Initially, it was thought that such a framework would not work in the context of an ANN since the iterative nature of gradient-based optimization algorithms do work well in the context of a set number of updates \parencite{ravi_optimization_2016} - and even more so when there are only a few examples to work with.
+Initially, it was thought that ANNs are incompatible with few-shot learning since the iterative nature of gradient-based optimization algorithms do not work well in the context of a set number of updates \parencite{ravi_optimization_2016} - and even more so when there are only a few examples to work with.
 Having said that, \textcite{rezende_one-shot_2016} were among the first to provide a solution to the one-shot problem in the context of a neural network by implementing Bayesian reasoning into a deep network embedded within hierarchical latent variable models.
-Since that, numerous efforts were made to develop an algorithm that enables an ANN to perform with a constrained number of class examples.
+Since that, numerous efforts were made to develop an algorithm that enables an ANN to perform with strictly limited datasets.
 Probably the most popular algorithm at the time of writing is the Model-Agnostic Meta-Learning (MAML) algorithm \parencite{finn_model-agnostic_2017}.
-One of MAML's core mechanics is the parameter update by calculating second-order derivatives.
-This procedure however implies a substantial computational penalty, which the authors did note.
-Therefore, they developed a first-order variant of MAML (FOMAML) which is able to perform similarly, if only slightly worse \parencite{finn_model-agnostic_2017}.
-
-This however, seemed to be the initial thought, until \textcite{nichol_first-order_2018} showed that FOMAML tends to be somewhat unstable and very sensitive to even the smallest changes in hyper-parameters.
-Therefore, they developed Reptile, a first-order meta-learning algorithm that rivals both MAML and FOMAML \parencite{nichol_first-order_2018}.
+One of MAML's core mechanics is that the parameters of the model are trained with a small number of gradient steps on with little training data to arrive at a good generalization on that task.
+However, this procedure requires calculating second-order derivatives which implies a substantial computational penalty - something which the authors did note.
+Therefore, they developed a first-order variant of MAML (FOMAML) which is able to perform similarly, if only with a small performance hit \parencite{finn_model-agnostic_2017}.
+
+This seemed to be the case until \textcite{nichol_first-order_2018} showed that FOMAML tends to be somewhat unstable.
+In response to that, they developed Reptile, a first-order meta-learning algorithm that rivals the performance of both MAML and FOMAML \parencite{nichol_first-order_2018}.
 Just like MAML, it counteracts the shortcomings of gradient-optimization algorithms by learning the model's initial parameters in order to maximize performance on novel tasks, it allows for any type of network to be used, and does not place any restrictions on the type of loss function that can be used \parencite{nichol_first-order_2018}.
-The difference between the two however, lies in the way the parameters are updated.
+The difference between the two, however, lies in the way the parameters are updated.
 
@@
 
-In essence, the way Reptile works (see Algorithm \ref{alg:reptile}) is as follows.
+In essence, Reptile works (see Algorithm \ref{alg:reptile}) as follows.
 A model with parameters $\theta$ is initialized, and a meta-loop (or meta-epochs) of $i$ iterations starts.
-Within it, a task $\tau$ (e.g, image classification, text classification, etc.) is sampled, along with its corresponding loss function $L_{\tau}$ (for clarity we assume that each type of task uses a different loss function).
+Within it, a task $\tau$ (e.g, image classification, text classification, etc.) is sampled, along with its corresponding loss function $L_{\tau}$ (for clarity, we assume that each type of task uses a different loss function).
 Then we compute the updated parameters $\tilde \theta$ in an inner-loop with $k$ iterations on task $\tau$ (or in other words, once we sampled our task $\tau$, we train our model $k$ times on it using the $\theta$ as parameters so we can get $\tilde \theta$).
-Finally, we update out initial parameters $\theta$ using this formula: $\theta \leftarrow \theta + \epsilon (\tilde{\theta} - \theta)$.
+Finally, we update our initial parameters $\theta$ using this formula: $\theta \leftarrow \theta + \epsilon (\tilde{\theta} - \theta)$.
 Then the algorithm continues to the next iteration of the meta-loop carrying over our updated parameters $\theta$.
 
-The explanation above is fairly verbose however, the algorithm itself is fairly simplistic but achieves similar performances to FOMAML \parencite{nichol_first-order_2018}.
+The explanation above is fairly verbose however, the algorithm itself is fairly simplistic and achieves similar performances to FOMAML \parencite{nichol_first-order_2018}.
 This forms a compelling reason to use Reptile as the few-shot learning method of choice for this paper.
@@
 To reiterate from Section \ref{sec:thesis_goal}, the purpose of this paper is to establish (1) the extent to which the music generated under a few-shot learning algorithm is comparable to that of a model trained on the entire dataset, and (2) the extent to which the music from such a model is comparable to that of real musical performances.
-For these purposes, Performance-RNN and C-RNN-GAN were adapted to the Reptile training procedure (Algorithm \ref{alg:reptile}) and evaluated against the performances generated by the baseline proposed in \textcite{larochelle_few-shot_2017} - a single layer LSTM with 256 units - and actual musical performances from the dataset (Section \ref{sec:dataset}).
+For these purposes, Performance-RNN and C-RNN-GAN were adapted to the Reptile training procedure (Algorithm \ref{alg:reptile}) and evaluated against the (1) songs generated by the baseline proposed in \textcite{larochelle_few-shot_2017} - a single layer LSTM with 256 units - and (2) actual musical performances from the dataset (Section \ref{sec:dataset}).
 For an in-depth explanation on the steps required to develop a model that is able to generate new data points refer to Section \ref{sec:how_to_gen} and to Section \ref{sec:reptile_train} for details on the training procedure.
@@
 
-Moreover, each performance comes with additional metadata such as, the name of the composer, the title of the performance, the suggested train, validation, test splits, year of the performance, name of the file, and duration in seconds of the performance.
+Moreover, each performance comes with additional metadata: the name of the composer, the title of the performance, the suggested train, validation, test splits, year of the performance, name of the file, and duration in seconds of the performance.
 Having said that, the dataset lacks clearly defined classes.
@@
 As such, given the definitions of \textcite{kennedy_concise_2007} on said genres and artists, each individual piece from our dataset was manually assigned a genre.
-Concretely, the style of music was firstly determined by the period in which the composer has lived (in baroque, classical, etc.) or - if the genres would overlap (i.e., modernism, and impressionism) - the composer's primary style of music would be chosen according to the definition of the composer given by \textcite{kennedy_concise_2007}.
+Concretely, the style of music was firstly determined by the period in which the composer has lived (in baroque, classical, etc.) or - if the time period of the genres would overlap (the case of modernism and impressionism) - the composer's primary style of music would be chosen according to the definition given by \textcite{kennedy_concise_2007}.
 Said labelling can be found in Appendix \ref{appendix:labelling}.
@@
 Here, the approach of \textcite{oore_this_2018} was adopted where MIDI midi messages are transformed into a sequence of one-hot encoded events.
-More specifically, MIDI excerpts are represented as a sequence of events from a vocabulary of 416 different events - similar to \textcite{oore_this_2018}, but with three extra events in the vocabulary:
+More specifically, MIDI excerpts are represented as a sequence of events from a vocabulary of 416 different events - similar to \textcite{oore_this_2018}, but with one extra events in the vocabulary:
 \begin{itemize}
@@
     \item 128 Note-Off events: releases a note (one for each of the 128 MIDI pitches)
-    \item 126 Time-Shift events: moves the time step forward somewhere between 0 ms and 1 second, in increments of 8 ms
+    \item 126 Time-Shift events: moves the time step forward somewhere between 0 ms and 1 second, in increments of 8 ms (\textcite{oore_this_2018} ignored the 0ms forward step)
     \item 32 Velocity events: changes the velocity for each subsequent note, until the next velocity event
@@
 A glance over the MIDI documentation \parencite{midi_association_official_2019} shows how complex and granular the standard is.
-However, we do not need to implement all of the messages into our model in order to make it work.
-We only need three: \texttt{note\_on}, \texttt{note\_off}, \texttt{pedal\_on/off}.
+Luckily however, we do not need to implement all of the messages into our model in order to make it work.
+We only need three: \texttt{note\_on}, \texttt{note\_off}, and \texttt{pedal\_on/off}.
 Figure \ref{fig:midi_examples} show an example of how the midi messages look like.
@@
 As it can be seen in Figure \ref{fig:midi_examples}, each message contains multiple pieces of information such as the message type, the note/value of the message, the delta time (time elapsed since last message and the current message, measured in ticks), and velocity (if applicable).
-To one-hot encode this stream of messages, it is necessary to separate it so it contains only one type of information per entry.
+To one-hot encode this stream of messages, it is necessary to separate it so it contains only one type of information per input.
 Thus, the sequence can be further broken down such that the first message that the system encounters is the delta time, followed by the velocity, and the note/pedal event (and their respective values).
@@
 With regards to the time movements, \textcite{friberg_overview_2006} note that a noticeable difference in temporally displacing a single tone in a sequence takes place from 10 ms and above.
-When in comes to the note velocities, it should be pointed to the reader that classical music features eight levels of common dynamic marking (from $ppp$ to $fff$) which is why \textcite{oore_this_2018} note that 32 steps are more than enough.
-
-\section{You're a musician Schmidhuber} \label{sec:how_to_gen}
-
-Knowing what we know about RNNs and LSTMs from Section \ref{sec:rnn} one can devise a method where the network can generate a sequence of events.
+Thus, since our movements are quantized in steps of 8 ms, there should be no discernible difference for the human hearing.
+When it comes to the note velocities, it should be pointed to the reader that classical music features eight levels of common dynamic marking (from $ppp$ to $fff$) which is why \textcite{oore_this_2018} note that 32 steps are more than enough.
+
+\section{How to make music} \label{sec:how_to_gen}
+
+Knowing what we know about RNNs and LSTMs from Section \ref{sec:rnn}, one can devise a method where the network can generate a sequence of events.
 All that is needed is to present the system with an event to kickstart the generative process.
-Namely, an input $x^t$ of size 1 is given to the network to output the next likely event $x^{t+1}$, which in turn is fed back to generate $x^{t+2}$.
+Namely, an input $x^t$ is fed to the network in order to output the next likely event $x^{t+1}$, which in turn is fed back to generate $x^{t+2}$.
 This process is repeated until we have a sequence of events of a desired length.
@@
 For instance, if the first input would be a C\#, the network could learn to always output E as the next note.
-Using a random hidden state is a measure that was included to ensure some diversity in the process of generating samples (granted, it should be admitted the possibility of this scenario being unlikely, and that the regular method described above would still work just as good).
+Using a random hidden state is a measure that was placed to ensure some diversity in the process of generating samples (granted, it should be admitted the possibility of this scenario being unlikely, and that the regular method described at the beginning of this section would still work just as good).
 
@@
 The reader might recall that this is part of the generation process in the framework of a GAN respectively, how the generator $G$ would create samples.
-Hence, all three models implemented in this paper feature the architecture of a GAN generator.
-
-In terms of generating the next event, \textcite{oore_this_2018} have defined a parameter $\tau$ called temperature \parencite{google_magenta_performance_2017}.
+As such, all three models implemented in this paper feature the architecture of a GAN generator.
+
+Moreover, in terms of generating the next event, \textcite{oore_this_2018} have defined a parameter $\lambda$ called temperature \parencite{google_magenta_performance_2017}.
 It is a value which uses the model's predicted events distribution to control the randomness of the output samples.
@@
 \begin{equation}
-    X = \text{softmax}(\phi / \tau).
+    X = \text{softmax}(\phi / \lambda).
 \end{equation}
-$X$ is the predicted event distribution of the network's output divided by our temperature $\tau$. Thus, $x^{t+1}$ is actually a random sample from our defined categorical distribution $X$.
-
-\section{How to train a Reptile} \label{sec:reptile_train}
-
-With the generative process outlined in the previous section, training the generative networks outlined in this papers is somewhat straightforward.
+$X$ is the predicted event distribution of the network's output divided by our temperature $\lambda$.
+Thus, $x^{t+1}$ is actually a random sample from our defined categorical distribution $X$.
+In this paper we train our models with a temperature of 1.0, and generate samples with temperatures ranging from 0.8 to 1.2, with an increment of 0.1.
+
+\section{Training procedure} \label{sec:reptile_train}
+
+With the generative process outlined in the previous section, training the generative networks presented in this papers is somewhat straightforward.
 A batch of songs is sampled from the training set, the model generates as many songs as there are in the batch, and then compute the loss between the network's output and the real samples.
-This statement is essentially true for the baseline and PerformanceRNN (in spite of the Reptile training procedure, inner-loop training is similar to the standard procedure).
-In contrast, the GAN model requires training two models simultaneously, where the generator's loss function is based on whether the discriminator is fooled by the fake samples (i.e., classifies them as real), and the discriminator's loss is based on whether it can distinguish between real and fake samples (it should be mentioned that both models would use binary cross-entropy as their loss function).
+This statement is essentially true for the baseline and Performance-RNN (in spite of the Reptile training procedure, inner-loop training is similar to the standard procedure).
+In contrast, the GAN model requires training two models simultaneously, where the generator's loss function is based on whether the discriminator is fooled by the fake samples (i.e., classifies them as real), and the discriminator's loss is based on whether it can distinguish between real and fake samples (it should be mentioned that both components of the GAN would use binary cross-entropy as their loss function; in contrast the baseline and Performance-RNN are based on cross entropy).
 However, there are still several items that we haven't touched upon.
@@
 A solution to the issue of synchronicity would be to freeze the training of one of the models when the loss is below a predefined threshold \parencite{mogren_c-rnn-gan_2016}.
-In this paper, we stop training one of the models when its training loss is less than half the loss of the other's.
+In this paper, we stop training one of the models if its training loss is less than half the loss of the other's.
 
@@
 Thus, the number of steps we have taken to compute $\tilde \theta$ is ignored.
-If there is skepticism about the fact that the Reptile models are trained much more than the baseline leading to better performances, the reader should be reminded by the fact that the latter is presented with 954 songs per epoch, whereas the former are trained with five.
+If this claim causes any skepticism about whether the models trained under Reptile would outperform simply because they "are trained much more", the reader should be reminded by the fact that the baseline is presented with 954 songs per epoch, whereas these models are trained with five.
 
@@
 
-A summary of the parameters used in the training procedures is found in Appendix \ref{appendix:parameters}.
+A summary of the parameters used in the training procedures is found in Table \ref{tab:params}.
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{l|c|c|c}
+                    &\bf Baseline   &\bf Performance RNN&\bf C-RNN-GAN  \\ \hline
+    Num. Layers     &   1           &   3               &   2           \\
+    Num. Units      &   256         &   512             &   350         \\
+    Dropout Rate    &   0.0         &   0.3             &   0.3         \\
+    Learning Rate   &   0.001       &   0.001           &   0.001       \\
+    Meta Rate       &   N/A         &   0.01            &   0.01        \\
+    Batch Size      &   2048        &   64              &   64          \\
+    Meta Epochs     &   250         &   250             &   250         \\
+    Inner Epochs    &   N/A         &   3               &   3           \\
+    Has bias vector &   No          &   Yes             &   Yes         \\
+    Gradient Clipped&   No          &   Yes             &   Yes         \\
+    \end{tabular}
+    \caption{Summary of parameters used to construct and train the three models featured in this paper}
+    \label{tab:params}
+\end{table}
 
@@
 However, as \textcite{borji_pros_2018} notes, NLL is uninformative about the quality of the samples generated and it does not allow to answer whether the generative network is simply memorizing training examples.
-As such, this study will also rely on other evaluation metrics when assessing generative models: the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018}, polyphony, scale consistency, repetitions, and tone span \parencite{mogren_c-rnn-gan_2016}.
+As such, this study will rely on other evaluation metrics when assessing generative models: the number of statistically different bins \parencite[NDB;][]{richardson_gans_2018}, polyphony, scale consistency, repetitions, and tone span \parencite{mogren_c-rnn-gan_2016}.
 
@@
 
+All of the experiments have been conducted in the programming language Python version 3.7.3 \parencite{rossum_python_2019} on a Nvidia T4 GPU as provided by Google Colaboratory \parencite{google_colaboratory_2019}.
+The models have been built entirely with PyTorch version 1.0.1 \parencite{paszke_automatic_2017}, while the dataset has been processed with a combination of libraries: mido version 1.2.9 \parencite{bjorndalen_mido_2018} for MIDI processing, numpy version 1.16.3 \parencite{walt_numpy_2011} and pandas version 0.23.4 \parencite{mckinney_data_2010} for processing arrays and data frames.
+Finally, scikit-learn version 0.20 \parencite{pedregosa_scikit-learn_2011} was used for its implementation of $K$-means clustering.
+The PyTorch reimplementation of Performance-RNN made by \textcite{lee_event-based_2019} was used as an example for developing all three models featured in this paper.
+
 \chapter{Results}\label{chap:results}
+
+To establish the degree to which the music of a Reptile-based model is comparable to (1) a model trained on whole dataset and (2) songs from the dataset, a total of three models have been trained (see Table \ref{tab:params} for a list of training parameters).
+Training of the baseline finished after 27 hours, 9 hours for the Reptile adaptation of Performance-RNN, and 19 hours for the Reptile adaptation of C-RNN-GAN.
+
+Afterwards, the three have been tasked to generate 125 songs of variable lengths - between 90\% and 110\% of the median length (34,222) of songs in the training set - with five different temperature settings ranging from 0.8 to 1.2.
+The generating process took approximately 1 hour per model per temperature.
+The resulting samples were then evaluated against the samples in the test set using the metrics described in Section \ref{sec:evaluation}: polyphony, repetitions, tone span, scale consistency, and number of statistically different bins ($N_{bins} = 20$; arbitrary choice).
+
+The results in Table \ref{tab:results} underline several things about the trained models, the most important one being that all three lack the concept of motif (the presence of a dominant/recurring theme within a piece).
+The clear giveaway sign is the substantial lack of repetition in the generated samples.
+This is enforced by the larger levels of polyphony in the three models when compared to the test set.
+In addition, all models seems to be unrestrained with respect to using the whole gamut of notes they have available, but considering that a piano keyboard has 88 keys and the MIDI standard allows for 128 values (notes), it seems that our models might be heading into the right direction.
+When it comes to playing in a scale, all models show are equal across the board, with extremely minor variations.
+Having said these, the domain-specific evaluation metrics indicate that the models are fairly close to eachother but that should not change the fact that there is still room for improvement when compared to the real samples.
+
+Having said that, the landscape changes a bit when looking at NDB, our other evaluation metric.
+With seven to ten bins statistically different (out of 20), our models seem to display a great deal of disparity form the test samples, even though they somewhat agree between themselves - in other words, they don't seem to be very different from each other.
+
+With these result we have gathered enough information to answer our research questions, where we could claim that our C-RNN-GAN implementation produces some of the most best results (as it scores best on 3/4 domain-specific metrics).
+However, a deep-dive into the performance at training time paints a whole different picture.
+
+\begin{landscape}
+\begin{table}[t]
+    \centering
+\begin{tabular}{l|l|l|l|l|l|l}
+    Model             & Temperature & Polyphony       & Repetitions       & Tone Span        & Scale Consistency & NDB     \\ \hline
+\textit{Test Set} & N/A         & \textit{0.4915} & \textit{814.3280} & \textit{66.9440} & \textit{0.6827}   & N/A     \\
+Baseline          & 0.8         & 0.5707          & \textbf{\color{base0D}388.0160}    & 78.4480          & 0.6770            & 11      \\
+Baseline          & 0.9         & 0.6054          & 373.0560          & 81.3120          & 0.6747            & 10      \\
+Baseline          & 1.0         & 0.6112          & 347.3600          & 83.0560          & 0.6728            & 10      \\
+Baseline          & 1.1         & 0.6292          & 334.4160          & 84.4400          & 0.6722            & \textbf{\color{base0D}7} \\
+Baseline          & 1.2         & 0.6060          & 316.8160          & 85.0480          & 0.6726            & 9       \\
+Performance-RNN   & 0.8         & 0.5684          & 377.8240          & 78.8400          & 0.6760            & 9       \\
+Performance-RNN   & 0.9         & 0.5852          & 364.9680          & 81.4880          & 0.6757            & 8       \\
+Performance-RNN   & 1.0         & 0.6219          & 354.0800          & 83.2400          & 0.6730            & 9       \\
+Performance-RNN   & 1.1         & 0.6342          & 343.2240          & 84.4640          & 0.6721            & 9       \\
+Performance-RNN   & 1.2         & 0.6061          & 312.2480          & 85.2400          & 0.6723            & 10      \\
+C-RNN-GAN         & 0.8         & \textbf{\color{base0D}0.5505}    & 369.5760          & \textbf{\color{base0D}77.7920}    & \textbf{\color{base0D}0.6776}      & 9       \\
+C-RNN-GAN         & 0.9         & 0.5940          & 370.5040          & 81.4320          & 0.6749            & 9       \\
+C-RNN-GAN         & 1.0         & 0.6060          & 348.3520          & 83.2640          & 0.6733            & 8       \\
+C-RNN-GAN         & 1.1         & 0.6343          & 340.4160          & 84.6800          & 0.6730            & 11      \\
+C-RNN-GAN         & 1.2         & 0.6057          & 313.6080          & 85.0080          & 0.6726            & 10     
+\end{tabular}
+\caption{Results of the evaluation process of the generated sample.
+For comparison, relevant statistics of the test samples have been included.
+The values which are closest to the test sample are in bold and highlighted in green.}
+\label{tab:results}
+\end{table}
+\end{landscape}
+
+\begin{figure}[ht]
+    \newlength\figureheight
+    \newlength\figurewidth
+    \setlength\figureheight{6cm}
+    \setlength\figurewidth{6cm}
+
+    \begin{subfigure}{0.5\linewidth}
+        \centering
+        \input{images/baseline.tex}
+        \caption{Baseline loss at training time}
+        \label{fig:loss_baseline}
+    \end{subfigure}
+    \begin{subfigure}{0.5\linewidth}
+        \centering
+        \input{images/performance_rnn_1_5.tex}
+        \caption{Performance-RNN loss at training time}
+        \label{fig:loss_performancernn}
+    \end{subfigure}
+
+    \setlength\figureheight{10cm}
+    \setlength\figurewidth{10cm}
+
+    \begin{subfigure}{\linewidth}
+        \centering
+        \input{images/c_rnn_gan_1_5.tex}
+        \caption{C-RNN-GAN loss at training time}
+        \label{fig:loss_crnngan}
+    \end{subfigure}
+    \caption{Training losses of the three models featured in this paper.
+Note that the C-RNN-GAN reimplementation has a different loss function than the other two.
+Also to note, the loss of C-RNN-GAN is zero for both the generator and discriminator throughout most of the training time.}
+    \label{fig:loss}
+\end{figure}
+
+When looking at Figure \ref{fig:loss} we gain an insight on how well the training has gone.
+First and foremost, the loss of Performance-RNN is as one would expect: the model starts somewhat poorly, and then incrementally tunes itself to perform better (Figure \ref{fig:loss_performancernn}).
+The diminishing up-and-down motions are most likely indicative of the optimizer's carving its path to convergence to a (hopefully) global minima.
+
+Figure \ref{fig:loss_baseline} is somewhat perplexing at first glance.
+The loss drops quickly, only to jump back to its starting position after $\approx$50 epochs.
+A simple - and most likely true - explanation is that the learning rate is too big for the model.
+Considering the granularity with which all models have been trained, it could be that the model has escaped a good local minima due to the high learning rate.
+This should not be cause for any concern as further training would bring back the network to a similar position (it should probably reminded to the reader that the experiments in this papers are done with only 250 iterations, whereas more sophisticated DL experiments are somewhere in the order of thousands of iterations).
+
+Finally, we have the losses of the GAN, a bewildering tableau with more questions than answers (Figure \ref{fig:loss_crnngan}).
+Firstly, it should be mentioned that GANs in and of themselves are somewhat finicky to train.
+Secondly, in a normal game of minimax, the loss of the generator and that of the discriminator should alternate adversarially: when one goes up, the other goes down.
+Here, they move together, to the point where both models reach a loss of zero - meaning that the generator is always able to fake the discriminator and the discriminator is always able to distinguish the fake samples from the real samples.
+This is likely due to the implementation of feature matching where instead of being trained in a competition, they are trained independently.
+However, this is the only explanation that I can come up with, and without other data with respect to the way the training progresses, it is difficult to say with certainty.
+
+Having said that, whichever is the reason, does not change the evidence to the study's research questions.
+Respectively, our data suggests that models trained under Reptile are directly competitive to the baseline (a model trained on the entire dataset).
+However, this does not mean that they are "good enough".
+When compared to our test samples, the data points generated by the Reptile models fail to reach the same level of quality, a limitation likely imposed by the number of training iterations.
 
@@
 
-\chapter{Parameters} \label{appendix:parameters}
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{l|c|c|c}
-                    &\bf Baseline   &\bf Performance RNN&\bf C-RNN-GAN  \\ \hline
-    Num. Layers     &   1           &   3               &   2           \\
-    Num. Units      &   256         &   512             &   350         \\
-    Dropout Rate    &   0.0         &   0.3             &   0.3         \\
-    Learning Rate   &   0.001       &   0.001           &   0.001       \\
-    Meta Rate       &   N/A         &   0.01            &   0.01        \\
-    Batch Size      &   2048        &   64              &   64          \\
-    Meta Epochs     &   250         &   250             &   250         \\
-    Inner Epochs    &   N/A         &   3               &   3           \\
-    Has bias vector &   No          &   Yes             &   Yes         \\
-    Gradient Clipped&   No          &   Yes             &   Yes         \\
-    \end{tabular}
-    \caption{Summary of parameters used to construct and train the three models featured in this paper}
-    \label{tab:params}
-\end{table}
-
 %TC:endignore

### 2019-05-08 forked_src/c_rnn_gan.py

@@ -217,14 +217,22 @@ class Discriminator(nn.Module):
         self.linear = nn.Linear(self.in_dim, 1)
         self.activation = nn.Sigmoid()
 
-    def forward(self, event, hidden=None, output_type='sigmoid'):
+    def forward(self,
+                event,
+                hidden=None,
+                output_type='sigmoid',
+                use_attn=False):
         assert output_type in ['sigmoid', 'logit', 'features']
         # shape events = steps, batch_size
         event = self.event_embedding(event.long())
 
         outputs, _ = self.lstm(event, hidden)
-        weights = (outputs * self.attn).sum(-1, keepdim=True)
-        output = (outputs * weights).mean(0)
+
+        if use_attn:
+            weights = (outputs * self.attn).sum(-1, keepdim=True)
+            output = (outputs * weights).mean(0)
+        else:
+            output = outputs
 
         output = self.features(output)
         if output_type == 'features':

### 2019-05-08 forked_src/evaluation.py

@@ -166,7 +166,7 @@ def evaluate_batch(paths, poly_criterion=31.35):
         'scale_consistency': []
     }
 
-    for path in tqdm(paths):
+    for path in tqdm(paths, desc='Domain-Specific Evaluation'):
         stats = evaluate_song(path, poly_criterion)
         for key in results.keys():
             results[key].append(stats[key])
@@ -192,19 +192,25 @@ def _prepare_data(paths, max_len=None):
         max_len = max(length)
 
     sequences = []
-    for i in tqdm(data, desc='Padding data'):
+    for i in data:
         diff = max_len - i.shape[0]
         if diff > 0:
             i = np.pad(i, pad_width=((0, diff)), mode='constant')
         elif diff < 0:
             i = i[:max_len]
 
-        sequences.append(i)
+        try:
+            sequences.append(i)
+        except ValueError:
+            continue
 
-    return np.vstack(sequences), max_len
+    try:
+        return np.vstack(sequences), max_len
+    except ValueError:
+        return None, None
 
 
-def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
+def ndb(real_data, gen_data, rng=1337, alpha_level=0.05, n_bins=5, workers=4):
     def assign_counts_to_bins(bins, count):
         for key in count.keys():
             bins[key] = count[key]
@@ -214,9 +220,6 @@ def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
         a = pool_prop[key] * (1 - pool_prop[key])
         b = (1 / real_data.shape[0]) + (1 / gen_data.shape[0])
         return sqrt((a * b))
-
-    if rng is None:
-        rng = np.random.RandomState(1337)
 
     n_samples = real_data.shape[0] + gen_data.shape[0]
 
@@ -271,3 +274,114 @@ def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
     })
     df.index.name = 'bins'
     return df
+
+
+if __name__ == '__main__':
+    import os
+    import torch
+    import torch.nn.functional as F
+    from glob import glob
+
+    rng = np.random.RandomState(1337)
+
+    GEN_DIR = '../generated/'
+    N_BINS = 20
+    WORKERS = 4
+
+    test_df = pd.read_csv('../data/maestro_test.csv')
+    test_paths = [
[…]

### 2019-05-08 forked_src/lstm_baseline.py

@@ -201,6 +201,8 @@ def train_baseline_large(vocab,
     # Use GPU if we have one
     device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
 
+    data_paths = ['../data/' + i for i in df['midi_filename']]
+
     model = Baseline(
         in_dim=in_dim,
         hidden_dim=hidden_dim,
@@ -234,32 +236,31 @@ def train_baseline_large(vocab,
     vocab = {i: vocab.index(i) for i in vocab}  # list to dict
 
     for epoch in range(meta_iters_start, meta_iters):
-        batches = data.mini_batches(
-            df['midi_filename'], rng, batch_size, replacement=False)
+        # batches = data.batch(seq_lens, batch_size, window_size, stride_size,
+        # rng)
         norm_epoch = []
         loss_epoch = []
 
+        batches = rng.choice(
+            data_paths, size=(len(data_paths) // batch_size, batch_size))
+        batches = batches.tolist()
+
         for batch in tqdm(batches, desc=f'Epoch {epoch+1}'):
+            # make variable sequences based on batch
+            sequences = data.build_sequences(batch)
+            # import pdb
+            # pdb.set_trace()
 
-            batch = [
-                os.path.join('../data/', df['midi_filename'][i]) for i in batch
-            ]
-
-            y = []
-            for song in batch:
-                sequence = data.build_sequences(song)
-                y += sequence
-
-            y = np.array([vocab[j] for j in y])
-            y = torch.from_numpy(y.T).to(device)
+            # goal is to make y.shape [lens, 1]
+            y = np.array([vocab[i] for i in sequences]).reshape(-1, 1)
+            y = torch.from_numpy(y).to(device)  # [window_size, 1]
             y = y.long()
-            y = y.view(-1, 1)
 
-            init = torch.randn(y.shape[0], model.init_dim).to(device)
-            model.batch_size = y.shape[0]
-
+            # change batch_size to 1
+            model.batch_size = 1
+            init = torch.randn(1, model.init_dim).to(device)
             outputs = model.generate(
-                init, window_size, y=y[:-1], output_type='logit')
+                init, y.shape[0], y=y[:-1], output_type='logit')
             # assert outputs.shape[:2] == y.shape[:2]
 
             loss = F.cross_entropy(outputs.contiguous().view(-1, in_dim),

### 2019-05-08 forked_src/train_c_rnn_gan.py

@@ -375,7 +375,8 @@ class ReptileGAN:
                                  self.stride_size, self.rng)
             for batch in batches:
                 if self.g_train:
-                    g_loss = self.train_generator()
+                    g_loss = self.train_generator(
+                        feature_matching=self.feature_matching)
                     g_loop_losses.append(g_loss)
 
                 if self.d_train:

### 2019-05-08 src/evaluation.py

@@ -166,7 +166,7 @@ def evaluate_batch(paths, poly_criterion=31.35):
         'scale_consistency': []
     }
 
-    for path in tqdm(paths):
+    for path in tqdm(paths, desc='Domain-Specific Evaluation'):
         stats = evaluate_song(path, poly_criterion)
         for key in results.keys():
             results[key].append(stats[key])
@@ -210,7 +210,7 @@ def _prepare_data(paths, max_len=None):
         return None, None
 
 
-def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
+def ndb(real_data, gen_data, rng=1337, alpha_level=0.05, n_bins=5, workers=4):
     def assign_counts_to_bins(bins, count):
         for key in count.keys():
             bins[key] = count[key]
@@ -220,9 +220,6 @@ def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
         a = pool_prop[key] * (1 - pool_prop[key])
         b = (1 / real_data.shape[0]) + (1 / gen_data.shape[0])
         return sqrt((a * b))
-
-    if rng is None:
-        rng = np.random.RandomState(1337)
 
     n_samples = real_data.shape[0] + gen_data.shape[0]
 
@@ -281,16 +278,31 @@ def ndb(real_data, gen_data, n_bins=5, alpha_level=0.05, rng=None, workers=4):
 
 if __name__ == '__main__':
     import os
+    import torch
+    import torch.nn.functional as F
     from glob import glob
 
     rng = np.random.RandomState(1337)
 
     GEN_DIR = '../generated/'
+    N_BINS = 20
+    WORKERS = 4
+
     test_df = pd.read_csv('../data/maestro_test.csv')
     test_paths = [
         os.path.join('../data/', i) for i in test_df['midi_filename']
     ]
     test_data, max_len = _prepare_data(test_paths)
+
+    results2 = pd.DataFrame(evaluate_batch(test_paths))
+    results2.to_csv('../generated/test_set_stats.csv')
+
+    # print('Building classifier')
+    # clf = KMeans(n_clusters=N_BINS, random_state=rng, n_jobs=WORKERS)
+    # clf.fit(test_data)
+
+    # print('Creating Torch arrays')
+    # torch_test = torch.from_numpy(test_data).long()
 
     models = [i[1] for i in os.walk(GEN_DIR)][0]  # get model folders
     for model in models:
[…]

### 2019-05-08 src/old/c_rnn_gan.py

@@ -20,7 +20,7 @@ class Generator(nn.Module):
                  in_dim,
                  hidden_dim,
                  batch_size,
-                 out_dim=415,
+                 out_dim=416,
                  dropout=0.3,
                  num_layers=2,
                  init_dim=100,
@@ -87,7 +87,7 @@ class Generator(nn.Module):
         return torch.LongTensor([[self.primary_event] * batch_size]).to(
             self.device)
 
-    def _sample_event(self, output, greedy=True, temperature=1.0):
+    def _sample_event(self, output, greedy=False, temperature=1.0):
         if greedy:
             return output.argmax(-1)
 
@@ -136,7 +136,7 @@ class Generator(nn.Module):
                  init,
                  steps,
                  y=None,
-                 greedy=True,
+                 greedy=False,
                  temperature=1.0,
                  output_type='index'):
 
@@ -199,10 +199,7 @@ class Discriminator(nn.Module):
         else:
             self.num_directions = 1
 
-        self.event_embedding = nn.Embedding(self.in_dim, self.in_dim)
-
-        self.in_layer = nn.Linear(self.in_dim, self.hidden_dim)
-        self.in_activation = nn.LeakyReLU(0.1, inplace=True)
+        self.event_embedding = nn.Embedding(self.in_dim, self.hidden_dim)
 
         self.lstm = nn.LSTM(
             input_size=self.hidden_dim,
@@ -215,26 +212,26 @@ class Discriminator(nn.Module):
         self.attn = nn.Parameter(
             torch.randn(self.hidden_dim * self.num_directions),
             requires_grad=True)
-        self.linear = nn.Linear(self.hidden_dim * self.num_directions, 1)
+        self.features = nn.Linear(self.hidden_dim * self.num_directions,
+                                  self.in_dim)
+        self.linear = nn.Linear(self.in_dim, 1)
         self.activation = nn.Sigmoid()
 
-    def forward(self,
-                event,
-                hidden=None,
-                input_is_logit=False,
-                output_logits=False):
+    def forward(self, event, hidden=None, output_type='sigmoid'):
+        assert output_type in ['sigmoid', 'logit', 'features']
         # shape events = steps, batch_size
-        if not input_is_logit:
-            event = self.event_embedding(event.long())
+        event = self.event_embedding(event.long())
 
-        x = self.in_layer(event)
-        x = self.in_activation(x)
-
-        outputs, _ = self.lstm(x, hidden)
+        outputs, _ = self.lstm(event, hidden)
[…]

### 2019-05-08 src/plots.py

@@ -1,0 +1,67 @@
+import os
+import torch
+import utils
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import matplotlib2tikz
+
+MODELS_DIR = '../models/'
+PLOTS_DIR = '../plots/'
+MODELS = ['performance_rnn_1_5', 'c_rnn_gan_1_5', 'baseline']
+plt.close('all')
+
+utils.make_folder(PLOTS_DIR)
+
+
+def make_plot(loss: list, title='', ylabel='', legend=[], save='./plot'):
+    mpl.rcParams['font.size'] = 16
+    plt.figure(figsize=(10, 10))
+
+    if len(loss) == 2:
+        plt.plot(loss[0], label=legend[0])
+        plt.plot(loss[1], label=legend[1])
+        plt.legend(loc='upper right')
+    else:
+        plt.plot(loss[0])
+
+    plt.xlabel('Epochs')
+    plt.ylabel(ylabel)
+    plt.title(title)
+
+    plt.savefig(save + '.png')
+    matplotlib2tikz.save(
+        save + '.tex',
+        figureheight='\\figureheight',
+        figurewidth='\\figurewidth')
+    plt.savefig(save + '.pdf', bbox_inches='tight')
+    plt.clf()
+
+
+for model in MODELS:
+    path = os.path.join(MODELS_DIR, model)
+    save_path = os.path.join(PLOTS_DIR, model)
+    saved = torch.load(path, map_location='cpu')
+
+    if not model is 'c_rnn_gan_1_5':
+
+        loss = [i[1] for i in saved['loss_saved']]
+        if model is 'performance_rnn_1_5':
+            title = 'Loss of Reptile-trained Performance RNN'
+        else:
+            title = 'Loss of baseline model'
+
+        make_plot([loss],
+                  title=title,
+                  ylabel='Cross-Entropy Loss',
+                  save=save_path)
+    else:
+        g_loss = [i[1] for i in saved['g_loss_saved']]
+        d_loss = [i[1] for i in saved['d_loss_saved']]
+
+        title = 'Losses of Reptile-trained C-RNN-GAN'
+
+        make_plot([g_loss, d_loss],
+                  title=title,
+                  ylabel='Binary Cross-Entropy Loss',
+                  legend=['Generator', 'Discriminator'],
+                  save=save_path)

### 2019-05-08 src/train_c_rnn_gan.py

@@ -119,7 +119,7 @@ class ReptileGAN:
         y = [sequences[i[0]:i[1]] for i in batch]
         y = np.array([[self.vocab[j] for j in i] for i in y])
         y = torch.from_numpy(y.T).to(self.device)
-        y = y.long()
+        # y = y.long()
         return y
 
     def init_models(self):
@@ -155,13 +155,15 @@ class ReptileGAN:
             num_layers=self.num_layers,
             use_bias=self.use_bias).to(self.device)
 
-        self.g_optim = optim.Adam(self.g.parameters(), lr=self.meta_step)
-        self.meta_g_optim = optim.SGD(
-            self.meta_g.parameters(), lr=self.learning_rate)
+        self.g_optim = optim.Adam(
+            self.g.parameters(), lr=self.meta_step, weight_decay=0.1)
+        self.meta_g_optim = optim.Adam(
+            self.meta_g.parameters(), lr=self.learning_rate, weight_decay=0.1)
 
-        self.d_optim = optim.Adam(self.d.parameters(), lr=self.meta_step)
+        self.d_optim = optim.SGD(
+            self.d.parameters(), lr=self.meta_step, weight_decay=0.1)
         self.meta_d_optim = optim.SGD(
-            self.meta_d.parameters(), lr=self.learning_rate)
+            self.meta_d.parameters(), lr=self.learning_rate, weight_decay=0.1)
 
         self.d_targets = torch.tensor(
             [1] * self.batch_size + [0] * self.batch_size,
@@ -259,7 +261,7 @@ class ReptileGAN:
             self.d_train = True
 
     def train_generator(self,
-                        feature_matching=True,
+                        feature_matching=False,
                         greedy=False,
                         temperature=1.0):
         # Generator
@@ -286,29 +288,40 @@ class ReptileGAN:
             #         torch.cat(generated, 0), output_type='logit')
             #     q_value = q_value.unsqueeze(0)
             #     q_values.append(q_value)
-            q_values = []
+
+            # q_values = []
+            # fake_batch = self.meta_g.generate(
+            #     init, self.window_size, greedy=greedy, temperature=temperature)
+            # for val in fake_batch:
+            #     q_val = self.meta_d(
+            #         val.view((1, val.shape[0])), output_type='logit')
+            #     q_values.append(q_val)
+
+            # q_values = torch.cat(q_values, 0)
+            # # pdb.set_trace()
+            # g_loss = F.mse_loss(
[…]

### 2019-05-10 paper/images/baseline/baseline-acts-transposed.tex
first draft done and added baseline_50

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Events Sequences},
xmin=-0.5, xmax=199.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Activations},
ymin=-0.5, ymax=255.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=255.5, ymax=-0.5] {images/baseline/baseline-acts-transposed-000.png};
\end{axis}

\end{tikzpicture}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5-acts-transposed.tex
first draft done and added baseline_50

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Events Sequences},
xmin=-0.5, xmax=199.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Activations},
ymin=-0.5, ymax=349.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5-acts-transposed-000.png};
\end{axis}

\end{tikzpicture}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5bach-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5-acts-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5-acts-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5bach-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5beethoven-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5bach-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5bach-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5beethoven-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5beethoven.tex
first draft done and added baseline_50

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Activations},
xmin=-0.5, xmax=349.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Events Sequences},
ymin=-0.5, ymax=199.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=349.5, ymin=199.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5beethoven-001.png};
\end{axis}

\end{tikzpicture}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5chopin-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5bach-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5bach-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5chopin-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5debussy-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5bach-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5bach-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5debussy-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5enescu-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5chopin-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5chopin-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5enescu-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-acts-.tex
first draft done and added baseline_50

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Activations},
xmin=-0.5, xmax=511.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Events Sequences},
ymin=-0.5, ymax=199.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-acts--000.png};
\end{axis}

\end{tikzpicture}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-acts-transposed.tex
first draft done and added baseline_50

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Events Sequences},
xmin=-0.5, xmax=199.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Activations},
ymin=-0.5, ymax=511.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-acts-transposed-000.png};
\end{axis}

\end{tikzpicture}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-acts.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-acts-.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-acts--000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-acts-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-gen-.tex
first draft done and added baseline_50

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Vocabulary},
xmin=-0.5, xmax=415.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Event Sequences},
ymin=-0.5, ymax=199.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=415.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-gen--001.png};
\end{axis}

\end{tikzpicture}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-acts-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-acts-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-acts.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-acts-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-001.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5beethoven-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5beethoven-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5beethoven.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-001.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5beethoven-001.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5chopin-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5chopin-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5chopin.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-001.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5chopin-001.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5debussy-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5debussy-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5debussy.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5bach.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5bach-001.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5debussy-001.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5enescu-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5chopin-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5chopin-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5enescu-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5enescu.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5chopin.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5chopin-001.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=511.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5enescu-001.png};
 \end{axis}

### 2019-05-10 paper/main.tex
first draft done and added baseline_50

@@
 
-\usepackage{draftwatermark}
-\SetWatermarkText{Draft}
-\SetWatermarkColor[gray]{0.9}
+%\usepackage{draftwatermark}
+%\SetWatermarkText{Draft}
+%\SetWatermarkColor[gray]{0.9}
 
@@
 
+% Links
+\usepackage{url}
+\def\UrlBreaks{\do\/\do-}
+\setcounter{biburllcpenalty}{7000}
+\setcounter{biburlucpenalty}{8000}
+
+\usepackage[colorlinks=true, citecolor=base08, linkcolor=base08, urlcolor=base08]{hyperref}
+
 % Essentials
@@
 \usepackage{fontenc}
-
-% Links
-\usepackage[colorlinks=true, citecolor=base08, linkcolor=base08, urlcolor=base08]{hyperref}
 
@@
 
+%TC:ignore
 \chapter*{Abstract}
-    Lorem Ipsum
+Experiments in Deep Learning are marked by scant explorations outside of the classification task, and even more so when trying to overcome some of the limitations imposed by gradient-based optimization algorithms (such as the uncertainty of convergence when dealing with non-convexity).
+With the resurgence of generative models and recent advancements in meta-learning we are faced two exciting ares of research, the intersection of which could be a promising area when investing general-purpose A.I.
+systems.
+Inspired by \textcite{clouatre_figr_2019} and \textcite{zhang_metagan_2018}, this paper proposed a new line of inquiry where these two fields could be useful specifically, in the musical domain.
+As such, the paper sought to establish the extent to which meta-learned generative models are comparable to a model trained under normal circumstances, and to real musical performances.
+
+Three models have been trained to generate music akin to the performances of expert musicians participating at the International Piano-e-Competition.
+Two of these models are models come from the literature and followed a 1-shot 5-way training with Reptile, a recent meta-learning algorithm.
+Afterwards they were evaluated against the baseline model (a LSTM network consisting of one layer with 256 units), and the test samples from the MAESTRO dataset.
+
+When compared to the baseline, the two Reptile-based models have show to be quite competitive however, they failed to attain the same level of mastery when compared to the test samples.
+However, it is hypothesized that more time involved into training them could narrow the gap between the models and the real musical performances.
+Thus, the results of these experiments solidify the intuition that meta-learning algorithms are a viable alternative to traditional training methods, in the context of generative models.
+
+%TC:endignore
 
@@
 
-\chapter{A hazy shade of winter}
+\chapter{A hazy shade of winter} \label{chap:intro}
 %Simon and Garfunkel - A hazy shade of winter
@@
 To combat mode collapse, \textcite{salimans_improved_2016} propose a technique called feature matching, where we change the objective of the generator to generate real data that matches the statistics of the real data.
-Since in the setting of a GAN showing the real data would be considered "cheating", we use the features of an intermediate layer of the discriminator.
 Specifically, we train the generator to match the expected value of those features \parencite{salimans_improved_2016}.
-Denoting $f(x)$ as the activations on an intermediate layer in the discriminator, the new objective for the generator becomes
-\begin{equation}
-    \|\mathbb{E}_{x~p_{data}} f(x) - \mathbb{E}_{z~p_z (z)} f(G(z))\|_2^2
-\end{equation}
-which is commonly known as the squared error.
-However, to make the re-implementation of C-RNN-GAN true to the original, Mean Square Error (MSE) was used as the loss function when using feature matching.
+However, this measure was not implemented in this paper as early tests caused both the loss of both the generator and discriminator to drop at zero, causing both models to stop learning.
 
@@
 To establish the degree to which the music of a Reptile-based model is comparable to (1) a model trained on whole dataset and (2) songs from the dataset, a total of three models have been trained (see Table \ref{tab:params} for a list of training parameters).
-Training of the baseline finished after 27 hours, 9 hours for the Reptile adaptation of Performance-RNN, and 19 hours for the Reptile adaptation of C-RNN-GAN.
+Training of the baseline finished after 27 hours, 9 hours for the Reptile adaptation of Performance-RNN, and 13 hours for the Reptile adaptation of C-RNN-GAN.
 
@@
 Having said that, the landscape changes a bit when looking at NDB, our other evaluation metric.
-With seven to ten bins statistically different (out of 20), our models seem to display a great deal of disparity form the test samples, even though they somewhat agree between themselves - in other words, they don't seem to be very different from each other.
-
-With these result we have gathered enough information to answer our research questions, where we could claim that our C-RNN-GAN implementation produces some of the most best results (as it scores best on 3/4 domain-specific metrics).
-However, a deep-dive into the performance at training time paints a whole different picture.
+With seven to ten bins statistically different (out of 20), our models seem to display a great deal of disparity form the test samples, even though they somewhat agree between themselves - i.e., they don't seem to be very different from each other.
+
+With these result we have gathered enough information to answer our research questions, and claim that the baseline still produces some of the best results (as it scores best on 4/5 metrics).
+However, a deep-dive into the performance at training time would shed some more light to these results.
 
@@
 \begin{tabular}{l|l|l|l|l|l|l}
-    Model             & Temperature & Polyphony       & Repetitions       & Tone Span        & Scale Consistency & NDB     \\ \hline
-\textit{Test Set} & N/A         & \textit{0.4915} & \textit{814.3280} & \textit{66.9440} & \textit{0.6827}   & N/A     \\
-Baseline          & 0.8         & 0.5707          & \textbf{\color{base0D}388.0160}    & 78.4480          & 0.6770            & 11      \\
-Baseline          & 0.9         & 0.6054          & 373.0560          & 81.3120          & 0.6747            & 10      \\
-Baseline          & 1.0         & 0.6112          & 347.3600          & 83.0560          & 0.6728            & 10      \\
-Baseline          & 1.1         & 0.6292          & 334.4160          & 84.4400          & 0.6722            & \textbf{\color{base0D}7} \\
-Baseline          & 1.2         & 0.6060          & 316.8160          & 85.0480          & 0.6726            & 9       \\
-Performance-RNN   & 0.8         & 0.5684          & 377.8240          & 78.8400          & 0.6760            & 9       \\
-Performance-RNN   & 0.9         & 0.5852          & 364.9680          & 81.4880          & 0.6757            & 8       \\
-Performance-RNN   & 1.0         & 0.6219          & 354.0800          & 83.2400          & 0.6730            & 9       \\
-Performance-RNN   & 1.1         & 0.6342          & 343.2240          & 84.4640          & 0.6721            & 9       \\
-Performance-RNN   & 1.2         & 0.6061          & 312.2480          & 85.2400          & 0.6723            & 10      \\
-C-RNN-GAN         & 0.8         & \textbf{\color{base0D}0.5505}    & 369.5760          & \textbf{\color{base0D}77.7920}    & \textbf{\color{base0D}0.6776}      & 9       \\
-C-RNN-GAN         & 0.9         & 0.5940          & 370.5040          & 81.4320          & 0.6749            & 9       \\
-C-RNN-GAN         & 1.0         & 0.6060          & 348.3520          & 83.2640          & 0.6733            & 8       \\
-C-RNN-GAN         & 1.1         & 0.6343          & 340.4160          & 84.6800          & 0.6730            & 11      \\
-C-RNN-GAN         & 1.2         & 0.6057          & 313.6080          & 85.0080          & 0.6726            & 10     
+    Model             & Temperature & Polyphony (\%)  & Repetitions       & Tone Span        & Scale Consistency (\%) & NDB     \\ \hline
+\textit{Test Set} & N/A         & \textit{49.15} & \textit{814.33} & \textit{66.9440} & \textit{68.27}   & N/A     \\
+Baseline          & 0.8         & 57.07          & \textbf{\color{base0D}388.02}    & \textbf{\color{base0D}78.45}          & \textbf{\color{base0D}67.70}            & 11      \\
+Baseline          & 0.9         & 60.54          & 373.06          & 81.31          & 67.47            & 10      \\
+Baseline          & 1.0         & 61.12          & 347.36          & 83.06          & 67.28            & 10      \\
+Baseline          & 1.1         & 62.92          & 334.42          & 84.44          & 67.22            & \textbf{\color{base0D}7} \\
+Baseline          & 1.2         & 60.60          & 316.82          & 85.05          & 67.26            & 9       \\
+Performance-RNN   & 0.8         & 56.84          & \textbf{\color{base0A}377.82}          & 78.84          & 67.60            & 9       \\
+Performance-RNN   & 0.9         & 58.52          & 364.97          & 81.49          & 67.57            & 8       \\
+Performance-RNN   & 1.0         & 62.19          & 354.08          & 83.24          & 67.30            & 9       \\
+Performance-RNN   & 1.1         & 63.42          & 343.22          & 84.46          & 67.21            & 9       \\
+Performance-RNN   & 1.2         & 60.61          & 312.25          & 85.24          & 67.23            & 10      \\
+C-RNN-GAN         & 0.8         & \textbf{\color{base0E}56.59}    & 375.25          & \textbf{\color{base0A}78.82}    & \textbf{\color{base0A}67.63}      & 9       \\
+C-RNN-GAN         & 0.9         & 58.41          & 365.19          & 81.22          & 67.50            & 9       \\
+C-RNN-GAN         & 1.0         & 61.25          & 351.02          & 83.00          & 67.40            & \textbf{\color{base0A}8}       \\
+C-RNN-GAN         & 1.1         & 63.49          & 338.39          & 84.66          & 67.29            & 11      \\
+C-RNN-GAN         & 1.2         & 60.57          & 316.35          & 85.30          & 67.19            & 10     
 \end{tabular}
@@
 For comparison, relevant statistics of the test samples have been included.
-The values which are closest to the test sample are in bold and highlighted in green.}
+Important values are in bold.
+The values which are closest to the test sample are in green.
+Values which are closest to the baseline's best score are in red.
+Values which are both close to the baseline's best and the test sample are in purple.}
 \label{tab:results}
@@
 
-    \begin{subfigure}{0.5\linewidth}
-        \centering
-        \input{images/baseline.tex}
-        \caption{Baseline loss at training time}
-        \label{fig:loss_baseline}
-    \end{subfigure}
     \begin{subfigure}{0.5\linewidth}
@@
     \end{subfigure}
-
-    \setlength\figureheight{10cm}
-    \setlength\figurewidth{10cm}
-
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.5\linewidth}
         \centering
@@
     \end{subfigure}
+
+    \setlength\figureheight{10cm}
+    \setlength\figurewidth{10cm}
+
+    \begin{subfigure}{\linewidth}
+        \centering
+        \input{images/baseline.tex}
+        \caption{Baseline loss at training time}
+        \label{fig:loss_baseline}
+    \end{subfigure}
     \caption{Training losses of the three models featured in this paper.
@@
 First and foremost, the loss of Performance-RNN is as one would expect: the model starts somewhat poorly, and then incrementally tunes itself to perform better (Figure \ref{fig:loss_performancernn}).
-The diminishing up-and-down motions are most likely indicative of the optimizer's carving its path to convergence to a (hopefully) global minima.
+Then, we have the losses of the GAN (Figure \ref{fig:loss_crnngan}) which portrays the adversarial training the generator and discriminator.
+Although on a first glance one might think that the loss is subject to high variance, it should be noted that a different function was used in this scenario and that the plot is on a much smaller scale than all others.
+Nevertheless, it does hint that the generator was slowly getting better at tricking the discriminator.
 
@@
 
-Finally, we have the losses of the GAN, a bewildering tableau with more questions than answers (Figure \ref{fig:loss_crnngan}).
-Firstly, it should be mentioned that GANs in and of themselves are somewhat finicky to train.
-Secondly, in a normal game of minimax, the loss of the generator and that of the discriminator should alternate adversarially: when one goes up, the other goes down.
-Here, they move together, to the point where both models reach a loss of zero - meaning that the generator is always able to fake the discriminator and the discriminator is always able to distinguish the fake samples from the real samples.
-This is likely due to the implementation of feature matching where instead of being trained in a competition, they are trained independently.
-However, this is the only explanation that I can come up with, and without other data with respect to the way the training progresses, it is difficult to say with certainty.
-
-Having said that, whichever is the reason, does not change the evidence to the study's research questions.
-Respectively, our data suggests that models trained under Reptile are directly competitive to the baseline (a model trained on the entire dataset).
-However, this does not mean that they are "good enough".
-When compared to our test samples, the data points generated by the Reptile models fail to reach the same level of quality, a limitation likely imposed by the number of training iterations.
+    \newlength\heatheight
+    \newlength\heatwidth
+    \setlength\heatheight{9.5cm}
+    \setlength\heatwidth{11.5cm}
+\begin{figure}
+        \begin{tabular}{c}
+            \input{images/baseline/baseline-gen-.tex} \\
+            \input{images/baseline/baseline-acts-.tex} \\
+    \end{tabular}
+    \caption{Baseline generated sample (top) and neuron activations for the respective event (bottom).
+The generated sample has a size of 200.
+The intensity of the color indicates the strength of the activation.}
+    \label{fig:baseline_activations}
+\end{figure}
+
+\begin{figure}
+        \begin{tabular}{c}
+            \input{images/performance_rnn/performance_rnn_1_5-gen-.tex} \\
+            \input{images/performance_rnn/performance_rnn_1_5-acts-.tex} \\
+    \end{tabular}
+    \caption{Performance-RNN generated sample (top) and neuron activations for the respective event (bottom).
+The generated sample has a size of 200.
+The intensity of the color indicates the strength of the activation.}
+    \label{fig:performancernn_activations}
+\end{figure}
+
+\begin{figure}
+        \begin{tabular}{c}
+            \input{images/c_rnn_gan/c_rnn_gan_1_5-gen-.tex} \\
+            \input{images/c_rnn_gan/c_rnn_gan_1_5-acts-.tex} \\
+    \end{tabular}
+    \caption{C-RNN-GAN generated sample (top) and neuron activations for the respective event (bottom).
+The generated sample has a size of 200.
+The intensity of the color indicates the strength of the activation.}
+    \label{fig:crnngan_activations}
+\end{figure}
+
+\begin{figure}[t]
+    \setlength\heatwidth{6.25cm}
+    \begin{subfigure}[b]{0.5\linewidth}
+        \setlength\heatheight{8cm}
+        \input{images/baseline/baseline-acts-transposed.tex}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.5\linewidth}
+        \setlength\heatheight{14.5cm}
+        \input{images/performance_rnn/performance_rnn_1_5-acts-transposed.tex}
+    \end{subfigure}
+    \caption{At scale neuron activations of the baseline (left) and Performance-RNN (right)}
+    \label{fig:acts_scaled}
+\end{figure}
+
+To add another layer of validation to the results in Table \ref{tab:results}, Softmax activations of the last (or only) LSTM layer for each of our models was plotted on a heatmap.
+These activations take place when the model is tasked with generating a song.
+When accounting for the fact that the scale of the X-axis of the neuron activation plot differs for each model (because each one has a different hidden size), a couple of things become concrete.
+Firstly, that the generating process of Performance-RNN (Figure \ref{fig:performancernn_activations}) is sparse but still similar to that of the baseline (Figure \ref{fig:baseline_activations}).
+To better illustrate this, Figure \ref{fig:acts_scaled} shows the two applications on the same scale.
+Secondly, that C-RNN-GAN (Figure \ref{fig:crnngan_activations}) still has a long way to go until it can be considered adequate at generating a song; in other words, C-RNN-GAN generates songs mostly at random, claim enforced by the fact that almost all neurons are activated and that the pattern of generating a sequence has no order.
+
+With these in mind, the data suggests that models trained under Reptile can be competitive with a model that has been trained under regular circumstances.
+However, this does not mean that they are all "good enough".
+When compared to our test samples, the songs generated by the Reptile models fail to reach the same level of quality, a limitation likely imposed by the number of training iterations.
 
@@
 
+From the lens of mainstream media, we are at a point where we will soon create what is commonly known as "Artificial Intelligence" \parencite{levy_ai_2010}.
+This claim, of course, ignores the popularity cycles of conectionism dating back to the fifties \parencite[e.g.,][]{rosenblatt_perceptron_1957}, or the over-inflated expectations for autonomous human-made systems that were made ever since \parencite{marcus_is_2012}.
+However, even after suffering waves of funding-cuts, masses of criticism, and regressions into obscurity, develpments were still being made \parencite{knight_ai_2016, minsky_perceptrons_1988} - albeit at a smaller pace.
+With the reemergence of ANNs, we are now seeing even stronger developments towards creating general-purpose Artificial Intelligence \parencite[e.g.,][]{vinyals_starcraft_2017}, but that should not cause any relaxation or distress; "we are not there yet".
+
+In Chapter \ref{chap:intro}, I pointed out a couple of shortcomings that current ANN systems have (reliance on massive amounts of data, and inability to perform separate tasks) along with two exciting trends in Deep Learning: the resurgence of generative networks and successfully infusing a system with meta-learning capabilities).
+Acknowledging the initial attempts at intersecting these two fields, this paper proposed expanding those efforts to the musical domain.
+It set to develop two generative models by modelling the training procedure to be more akin to that of a human learning, and evaluated their performances against a baseline proposed by \textcite{larochelle_few-shot_2017} and their resemblance to real samples.
+Concretely, Performance-RNN \parencite{oore_this_2018} and C-RNN-GAN \parencite{mogren_c-rnn-gan_2016} were trained under the Reptile learning framework \parencite{nichol_first-order_2018}, and then evaluated against a LSTM network with one layer and 256 hidden units as well as expert piano performances extracted from the MAESTRO dataset \parencite{hawthorne_enabling_2018}.
+
+In general, the baseline recorded the best performance.
+However, this does not change the fact that it was closely followed by the Reptile models.
+More appropriately is to point that, in spite of being trained to generate music using a handful of examples, the Reptile models closely matched the performance of the baseline.
+This especially holds true for this paper's reimplementation of Performance-RNN whose neuron activations show that it has indeed learned how to generate music, even though its samples scored somewhere in the middle.
+The adversarial model showed that it is still somewhere in the early stages of learning, which would explain why almost all of its neurons were fired but, the fact that it performed reasonably in its evaluation should not be ignored either.
+Having said these, I say with confidence that when tasked with making music, Reptile-based generative models can be quite competitive.
+
+As well as they perform, both few-shot models failed to produce samples that could be passed as real.
+However, neither did the baseline, which would indicate that with more time invested into training, all three models could generate samples that are both novel and high fidelity.
+
+Which leads to one of the first limitations of this paper, namely the training procedure.
+Optimally, all models should be trained for over 1,000 epochs however, time and hardware constrains meant taking some shortcuts.
+This constraint is especially seen in the case of the baseline which has been trained with a batch size 32 times larger than the Reptile models (to drive down the training time from 250 hours to 27).
+In addition, it is likely that the training of the GAN model suffered from the fact that a measure for combating mode-collapse was not implemented.
+Had it been done so, then it's possible that the model would have had better activations and consequently, better samples.
+
+Next, the generating process from which we sample with temperature to generate the next likely event is a somewhat greedy process \parencite{holtzman_curious_2019}.
+Assuming that the probability distribution of how humans make music borrows some of its characteristics from how we create language - which is likely since both tasks share a significant portion of brain activations \parencite{brown_music_2006} -, then greedy sampling or beam search (another sampling method) would completely fail to generate human-like music.
+Specifically, they are not ideal and nor adequate for a high entropy task \parencite{holtzman_curious_2019}.
+This could explain why the models in this paper lacked the concept of motif.
+Curiously enough, two recent solutions for sampling tokens have been shown to be better candidates: top-$k$ sampling \parencite{fan_hierarchical_2018} and nucleus sampling \parencite[also known as top-$p$ sampling;][]{holtzman_curious_2019}.
+Future work could incorporate either of these proposed sampling methods and evaluate them against the more traditional sampling methods in the context of generating music with a meta-learning algorithm.
+
+Which brings me to a final limitation.
+Performance-RNN and C-RNN-GAN are using Reptile, a first-order meta-learning algorithm.
+\textcite{nichol_first-order_2018} have provided clear evidence that this algorithm is more stable than MAML's first-order variant.
+However, their results also point that vanilla MAML is still superior.
+Having said that, these algorithms encompass only one approach to meta-learning (initialisation-based learning) whereas hallucination or distance metric methods could paint a different picture \parencite{chen_closer_2018}.
+Thus, future work could experiment not only with MAML, but also other types of algorithms.
+
+Nevertheless, the results in this paper are intriguing in and of themselves.
+They not only point towards neural networks as a viable infrastructure for music generation, but also towards training such a system in a manner that is (loosely) more akin to the way a human being would.
+By this I exclusively refer to the quantity of examples involved in the generative process.
+More importantly, these results outline the possibility of integrating generative networks and few-shot learning into the production process of a musician.
+To be more precise, \textcite{larochelle_few-shot_2017} indicate that instead of training a model on various types of music to influence the style of the generated samples, a musician could substitute the data from a support set with music of their own choice and obtain a new generative model for that particular style of music (admittedly, after a few more hours of training).
+
 \chapter{Conclusion}\label{chap:conclusion}
 
+This section aims to provide a clear and concise answer to the research questions proposed in this Chapter \ref{chap:intro}. They were formulated as follows:
+
+\begin{itemize}
+    \item To what extent is the music created by a few-shot generative model comparable to the music of a generative model that is trained on the entire dataset?
+    \item To what extent is the music created by a few-shot generative model comparable to real music?
+\end{itemize}
+
+The answers to both questions are summarized in Table \ref{tab:results}.
+Specifically, the evaluation of the Reptile models indicates that samples from these models are not exactly as strong as those of the baseline.
+However, the majority of the Reptile samples seem to lag only slightly, reason for which they could be aptly described as competitive and functionally equivalent.
+
+With regards to the performances of expert pianists, the Reptile samples failed to reach the same levels of mastery.
+However, it should be noted that the baseline showed similar shortcomings.
+It was hypothesized that the reason for this is the short training which the models had undergone.
+In this light, a tentative answer to the second research question is that the Reptile models can be similar to real music.
+Here, a significant area of improvement lies in embedding a sense of motif into the models (since this is what the models have show to lack the most).
+
+\sloppy
 \printbibliography

### 2019-05-10 plots/performance_rnn_1_5beethoven-transposed.tex
first draft done and added baseline_50

--- 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5enescu-transposed.tex
@@
 ]
-\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5enescu-transposed-000.png};
+\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=511.5, ymax=-0.5] {performance_rnn_1_5beethoven-transposed-000.png};
 \end{axis}

### 2019-05-10 paper/images/baseline/fix.sh

@@ -1,0 +1,3 @@
+for i in *.tex; do
+    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/baseline\//g' $i
+done

### 2019-05-10 paper/images/c_rnn_gan/fix.sh

@@ -1,0 +1,3 @@
+for i in *.tex; do
+    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/c_rnn_gan\//g' $i
+done

### 2019-05-10 paper/images/copy_from_plots.sh

@@ -1,0 +1,11 @@
+for i in ../../plots/baseline*; do
+    cp $i ./baseline/
+done
+
+for i in ../../plots/performance_rnn*; do
+    cp $i ./performance_rnn/
+done
+
+for i in ../../plots/c_rnn_gan*; do
+    cp $i ./c_rnn_gan/
+done

### 2019-05-10 paper/images/fix_images.sh

@@ -1,0 +1,9 @@
+cd ./baseline/
+./fix.sh
+cd ../
+cd ./performance_rnn/
+./fix.sh
+cd ../
+cd ./c_rnn_gan/
+./fix.sh
+cd ../

### 2019-05-10 paper/images/onehots/fix.sh

@@ -1,0 +1,3 @@
+for i in *.tex; do
+    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/onehots\//g' $i
+done

### 2019-05-10 paper/images/performance_rnn/fix.sh

@@ -1,0 +1,3 @@
+for i in *.tex; do
+    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/performance_rnn\//g' $i
+done

### 2019-05-10 src/generate.py

@@ -22,7 +22,9 @@ if __name__ == '__main__':
     greed = False
     temperatures = [0.8, 0.9, 1.0, 1.1, 1.2]
 
-    models = ['baseline', 'performance_rnn_1_5', 'c_rnn_gan_1_5']
+    models = [
+        'baseline_50', 'baseline', 'performance_rnn_1_5', 'c_rnn_gan_1_5'
+    ]
 
     make_folder('../generated')
     for model in models:
@@ -30,6 +32,19 @@ if __name__ == '__main__':
 
         checkpoint = load_model_data(os.path.join('../models/', model))
         if model == 'baseline':
+            from lstm_baseline import Baseline
+            net = Baseline(
+                in_dim=checkpoint['in_dim'],
+                hidden_dim=checkpoint['hidden_dim'],
+                batch_size=1,  # we create one song at a time
+                out_dim=checkpoint['out_dim'],
+                dropout=checkpoint['dropout'],
+                num_layers=checkpoint['num_layers'],
+                init_dim=checkpoint['init_dim'],
+                use_bias=checkpoint['use_bias'],
+                is_bidirectional=checkpoint['is_bidirectional'])
+            net.load_state_dict(checkpoint['model_state'])
+        elif model == 'baseline_50':
             from lstm_baseline import Baseline
             net = Baseline(
                 in_dim=checkpoint['in_dim'],
@@ -66,6 +81,7 @@ if __name__ == '__main__':
                 num_layers=checkpoint['num_layers'],
                 init_dim=checkpoint['init_dim'],
                 use_bias=checkpoint['use_bias'])
+            net.load_state_dict(checkpoint['g_model_state'])
 
         net.to(device)
 

### 2019-05-10 src/plots.py

@@ -1,16 +1,12 @@
 import os
 import torch
 import utils
+import data
+import numpy as np
 import matplotlib as mpl
 import matplotlib.pyplot as plt
 import matplotlib2tikz
-
-MODELS_DIR = '../models/'
-PLOTS_DIR = '../plots/'
-MODELS = ['performance_rnn_1_5', 'c_rnn_gan_1_5', 'baseline']
-plt.close('all')
-
-utils.make_folder(PLOTS_DIR)
+from sklearn.manifold import TSNE
 
 
 def make_plot(loss: list, title='', ylabel='', legend=[], save='./plot'):
@@ -37,31 +33,322 @@ def make_plot(loss: list, title='', ylabel='', legend=[], save='./plot'):
     plt.clf()
 
 
-for model in MODELS:
-    path = os.path.join(MODELS_DIR, model)
-    save_path = os.path.join(PLOTS_DIR, model)
-    saved = torch.load(path, map_location='cpu')
+def get_song_activations(model, y, device):
+    model.eval()
+    hidden = (torch.zeros(model.num_layers, 1, model.hidden_dim).to(device),
+              torch.zeros(model.num_layers, 1, model.hidden_dim).to(device))
 
-    if not model is 'c_rnn_gan_1_5':
+    activations = []
+    # import pdb
+    # pdb.set_trace()
+    for i in range(y.shape[0]):
+        event = y[i].view(1, 1)
+        event = model.event_embedding(event)
+        x = model.in_layer(event)
+        x = model.in_activation(x)
 
-        loss = [i[1] for i in saved['loss_saved']]
-        if model is 'performance_rnn_1_5':
-            title = 'Loss of Reptile-trained Performance RNN'
+        out, hidden = model.lstm(x.float(), hidden)
+        output = hidden[0].permute(1, 0, 2).contiguous()
+        output = output.view(1, -1).unsqueeze(0)
+        import pdb
+        # pdb.set_trace()
+        output = model.out_layer(output)
+
+        # activations.append(model.out_activation(output))
+        # activations.append(model.out_activation(hidden[0][-1]))
+        activations.append(model.out_activation(out))
+
+    return torch.cat(activations, 0)
+
+
+def get_generated_activations(model, device, steps=200):
+    model.eval()
+
+    init = torch.randn(model.batch_size, model.init_dim).to(device)
+    event = model.get_primary_event(model.batch_size)
+    hidden = model.init_to_hidden(init)
+    outputs = []
[…]

### 2019-05-10 src/train_c_rnn_gan.py

@@ -165,13 +165,23 @@ class ReptileGAN:
         self.meta_d_optim = optim.SGD(
             self.meta_d.parameters(), lr=self.learning_rate, weight_decay=0.1)
 
-        self.d_targets = torch.tensor(
-            [1] * self.batch_size + [0] * self.batch_size,
-            dtype=torch.float,
-            device=self.device).view(-1).to(self.device)
-        self.g_targets = torch.tensor(
-            [1] * self.batch_size, dtype=torch.float,
-            device=self.device).view(-1).to(self.device)
+    def init_labels(self):
+        rand_ind = np.random.randint(0, 10)
+        target_1 = np.random.uniform(0.9, 1.0, size=self.batch_size)
+        target_2 = np.random.uniform(0.0, 0.1, size=self.batch_size)
+
+        if rand_ind != 1:
+            # Sort of randomly...
+            targets = np.hstack([target_1, target_2])
+        else:
+            # Flip the labels every once in a while
+            targets = np.hstack([target_2, target_1])
+
+        self.d_targets = torch.from_numpy(targets).float().view(-1).to(
+            self.device)
+
+        self.g_targets = torch.from_numpy(target_1).float().view(-1).to(
+            self.device)
 
     def reset_inner_model(self):
         self.meta_g.train()
@@ -309,9 +319,9 @@ class ReptileGAN:
             fake_batch = self.meta_g.generate(
                 init, self.window_size, greedy=True)
             output = self.meta_d(fake_batch, output_type='sigmoid')
-            g_loss = torch.mean(
-                -torch.log(torch.clamp(output, 1e-1000000, 1.0)))
-
+            # g_loss = torch.mean(
+            #     -torch.log(torch.clamp(output, 1e-1000000, 1.0)))
+            g_loss = F.binary_cross_entropy(output, self.g_targets)
             train_loss = g_loss.item()
         else:
             fake_batch = self.meta_g.generate(
@@ -320,8 +330,9 @@ class ReptileGAN:
             # g_loss = F.cross_entropy(output, self.g_targets)
             # g_loss = F.binary_cross_entropy(output, self.g_targets)
             # Use mogren's loss calculation
-            g_loss = torch.mean(
-                -torch.log(torch.clamp(output, 1e-1000000, 1.0)))
+            # g_loss = torch.mean(
[…]

### 2019-05-16 paper/main.tex
thesis == done (i hope)

@@
 
-\usepackage[colorlinks=true, citecolor=base08, linkcolor=base08, urlcolor=base08]{hyperref}
+\usepackage[colorlinks=true, citecolor=black, linkcolor=black, urlcolor=black]{hyperref}
 
@@
 \usepackage{graphicx}
+\usepackage[font=small, labelfont=bf]{caption}
 \usepackage{subcaption}
@@
 
+\newenvironment{acknowledgements}%
+    {\cleardoublepage\thispagestyle{empty}\null\vfill\begin{center}%
+    \bfseries Acknowledgements\end{center}}%
+    {\vfill\null}
+
 \begin{document}
@@
 
+\cleardoublepage
+\begin{center}
+    \vspace*{\fill}
+\begin{figure}[ht]
+    \centering
+    \def\svgwidth{\linewidth}
+    \input{images/the_lick.pdf_tex}
+\end{figure}
+    \vspace{1em}
+\end{center}
+
 %TC:ignore
@@
 systems.
-Inspired by \textcite{clouatre_figr_2019} and \textcite{zhang_metagan_2018}, this paper proposed a new line of inquiry where these two fields could be useful specifically, in the musical domain.
+Inspired by recent literature, this paper proposed a new line of inquiry where these two fields could be useful specifically, in the musical domain.
 As such, the paper sought to establish the extent to which meta-learned generative models are comparable to a model trained under normal circumstances, and to real musical performances.
@@
 
-When compared to the baseline, the two Reptile-based models have show to be quite competitive however, they failed to attain the same level of mastery when compared to the test samples.
+When compared to the baseline, the two Reptile-based models have shown to be quite competitive however, they failed to attain the same level of mastery when compared to the test samples.
 However, it is hypothesized that more time involved into training them could narrow the gap between the models and the real musical performances.
@@
 
-\chapter{A hazy shade of winter} \label{chap:intro}
-%Simon and Garfunkel - A hazy shade of winter
+\chapter{Introduction} \label{chap:intro}
 
@@
 This way, there is no need for manual input on the sort of knowledge that the computer needs \parencite{goodfellow_deep_2016}.
-A more simplistic and concise interpretation of DL would be to see it as a statistical technique for identifying patterns in sample data using large (deep) ANNs \parencite{marcus_deep_2018}.
 
@@
 
-\section{Strange Fruit}\label{sec:no_disc}
-% Billie Holiday - Strange Fruit
+\section{A variety of applications}\label{sec:no_disc}
 
@@
 
-Regardless, successes in generative models have been either scarce or unattractive, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run.
-Given the age and - alleged - popularity, a valid question to ask is why things haven't moved that much since then.
-As it turn out, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$).
+Regardless, successes in generative models have been either scarce or unattractive, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for nearly two decades.
+Given that we have had a solution for all this time, a valid question to ask is why things haven't moved that much since then.
+As it turn out, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$, where $n$ is the number of samples).
 Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}.
@@
 
-The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this field.
-Besides countering the drawbacks outlined above, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.
-
+The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this area.
+Besides countering the drawbacks outlined above, they have been particularly successful at a variety of tasks such as image super-resolution \parencite{ledig_photo-realistic_2016}, or style-based image generation \parencite{karras_style-based_2018}.
 Similarly, and although not as popular, Recurrent Neural Networks (RNNs) have been successful at various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}.
@@
 That being said, Long Short-Term Memory networks \parencite[LSTM;][]{hochreiter_long_1997} - a variant of RNNs - have been purpose-built with memory cells which permit better storing and accessing of information.
-This design decision variant of the recurrent network good candidates for generating sequences with long-term dependencies \parencite{graves_generating_2013}.
-
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}{.24\linewidth}
-        \includegraphics[width=\linewidth]{images/comic_HR.jpg}
-        \caption{original}
-        \label{fig:srgan_orig}
-    \end{subfigure}
-    \begin{subfigure}{.24\linewidth}
-        \includegraphics[width=\linewidth]{images/comic_SRF_4_bicubic.jpg}
-        \caption{bicubic}
-        \label{fig:srgan_bicubic}
-    \end{subfigure}
-    \begin{subfigure}{.24\linewidth}
-        \includegraphics[width=\linewidth]{images/comic_SRResNet-MSE.jpg}
-        \caption{SRResNet}
-        \label{fig:srgan_srresnet}
-    \end{subfigure}
-    \begin{subfigure}{.24\linewidth}
-        \includegraphics[width=\linewidth]{images/comic_SRGAN-VGG54.jpg}
-        \caption{SRGAN}
-        \label{fig:srgan_srgan}
-    \end{subfigure}
-
-    \caption{Example of single-image super-resolution results that highlight the advantages of a GAN.
-The original high-resolution image (\ref{fig:srgan_orig}) has been downsampled to make a low-resolution image.
-The results of bicubic interpolation (\ref{fig:srgan_bicubic}) can be seen in the second image, followed by the results of SRResNet (\ref{fig:srgan_srresnet}), a neural network trained on mean squared error.
-Lastly, the results of SRGAN can be seen in the final image (\ref{fig:srgan_srgan}).
-Figure adapted from \textcite{ledig_photo-realistic_2016}.}
-    \label{fig:srgan}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-    \begin{tabular}{ccc}
-    Input & Ground Truth & Output \\
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_106_AB.jpg} &
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_106_AB.jpg} &
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_106_AB.jpg} \\
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_12_AB.jpg} &
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_12_AB.jpg} &
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_12_AB.jpg} \\
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/input_130_AB.jpg} &
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/gt_130_AB.jpg} &
-    \includegraphics[width=.15\linewidth]{images/handbags_edges_lotsofresults_latex/L1cGAN_130_AB.jpg} \\
-    \end{tabular}
-    \caption{\textcite{isola_image--image_2016} coined the term image-to-image translation where an input image is transformed into a revisualization of the original input.
-This figures illustrates how an outline of a purse (leftmost column) is transformed into a colored photorealistic version of it (rightmost column).
-In the middle we have the ground truth: the real image of the purse.
-Figure adapted from \textcite{isola_image--image_2016}.}
-    \label{fig:pix2pix}
-\end{figure}
-
-Having said all of these, the majority of developments with generative models lie in the field of Computer Vision.
+This variant of the recurrent network makes it an excellent candidate for generating sequences with long-term dependencies \parencite{graves_generating_2013}.
+
+In light of the above, the majority of developments with generative models lie in the field of Computer Vision.
 But that is not to say that we can generate only images.
-Generative models have been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}.
+Generative models have also been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}.
 Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}.
-Herein lies the scope of this paper.
-Given such minimal explorations into generating music, we evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and a LSTM network \parencite{oore_this_2018} trained under the constraints imposed by few-shot learning (Section \ref{sec:hungry}) against a baseline model proposed by \textcite{larochelle_few-shot_2017}.
-To reiterate, a full explanation of the purpose of the paper is provided in Section \ref{sec:thesis_goal}.
-
-\section{I want it all}\label{sec:hungry}
-% Queen - I want it all
-
-DL models usually rely on a procedure called stochastic gradient descent (SGD) which implies computing an output and an error for a given input vector, followed by calculating the input's average gradient and adjusting the network's parameters in a manner that minimizes the error \parencite{lecun_deep_2015}.
+
+\textcite{briot_deep_2017} have conducted one of the most comprehensive survey of DL techniques and models for generating music, work which covers experiments from the last 25 years.
+What is striking, is that the authors admit that in spite of its rich history in computer science, there have been only a few attempts at generating music with ANNs.
+Thus, herein lies the scope of this paper.
+Given such minimal explorations, I evaluated the songs generated by a GAN \parencite{mogren_c-rnn-gan_2016} and a LSTM network \parencite{oore_this_2018} trained with a handful of examples against a baseline model that had access to the entire training set.
+To understand why such an assessment is needed, we need to introduce some of the shortcomings of modern ANNs and how to circumvent them.
+
+\section{Pitfalls in gradient descent}\label{sec:hungry}
+
+The training of DL models usually relies on a procedure called stochastic gradient descent (SGD) which implies computing an output and an error for a given input vector, followed by calculating the input's average gradient and adjusting the network's parameters in a manner that minimizes the error \parencite{lecun_deep_2015}.
 \textcite{ravi_optimization_2016} argue that the iterative nature of gradient-optimization algorithms does not allow them to perform well under the constraint of a set number of updates over few examples.
@@
 Ultimately, this hurts the network's ability to converge to a good solution when constrained by a limited number of updates \parencite{ravi_optimization_2016}.
-Thus they lay out two clear drawbacks of DL models: "data-hunger", and inability to perform separate tasks.
-
-A clear outcome from solving these issues would be that DL models are easier to train (less data) as well as multifunctional.
-Nonetheless, this list is not exhaustive.
-There is also fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack.
+Thus they lay out two clear drawbacks of DL models: "data-hunger", and performing separate tasks is problematic.
+
+A clear outcome from solving these issues would be that DL models are easier to train (less data) as well as being able to carry out multiple tasks.
+But there is also the fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack.
 Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class.
@@
 In spite of its proposed benefits, developments and experiments with few-shot learning are still scarce \parencite{larochelle_few-shot_2017}.
-Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}.
-Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early (and few) explorers in this direction.
-
-\section{Composing like a human} \label{sec:thesis_goal}
-
-At the time of writing, generative models and meta-learning are two exciting areas of research the intersection of which could be quite promising.
-However, to the knowledge of the researcher, no experiments which combine the both of them have been conducted in the musical domain.
-Thus, this paper examined the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning.
-In other words, it sought to evaluate the degree to which generative models can create novel and high-fidelity samples when training data is scarce.
+Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}, whereas those with music are virtually non-existent.
+
+\section{Problem definition} \label{sec:thesis_goal}
+
+At the time of writing, generative models and meta-learning are two exciting areas of research, the intersection of which could be interesting for A.I.
+research.
+Having said that, there are scarcely any few-shot experiments with generative networks; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early (and few) explorers in this direction.
+Moreover, to the knowledge of the researcher, no such experiments have been conducted in the musical domain to this day.
+Thus, to enrich the current literature of DL in music, this paper examines the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning.
+In other words, it seeks to evaluate the degree to which generative models can create novel and high-fidelity samples when trained in a manner which is (quantitatively) more akin to that of an aspiring musician.
 To avoid any ambiguity, novel is used here as new (i.e., not in the training set), and qualitative as consistent with common scales and structurally diverse (combines chord structures with single-note sequences and is not restricted to only a few notes).
 
-This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, with a limited number of songs.
+This research evaluates the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, with a limited number of songs.
 The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.
 
-In other words, this study sought to answer the following research questions:
+In other words, this study seeks to answer the following research questions:
 
@@
 
-The following chapters will discuss more about generative models, meta-learning, and some of the models and algorithms involved in this paper (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the setup of the experiments (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}).
+The following chapters will discuss generative models, meta-learning, and some of the models and algorithms involved in this paper in more depth (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the experimental setup (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}).
 Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} concludes the findings of this paper.
@@
 
-This chapter takes a deep-dive into the fields of meta-learning and generative networks in order to highlight their current state of affairs.
+To highlight the current state of affairs, this chapter takes a deep-dive into the fields of meta-learning and generative networks.
 This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}.
 Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network.
-Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning and some state-of-the-art approaches.
+Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning and recent state-of-the-art approaches.
 
@@
 
-For transparency, it should be noted that it is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution forms the basis of a time-delay neural network.
+For transparency, it should be noted that it is possible to use a feedforward architecture with sequential data.
+This would take the form of a Convolutional Neural Network where the convolution forms the basis of a time-delay system.
 Unfortunately, the output of such a system would be a function of a small number of neighbouring data points \parencite[such as the last and the next three notes in a song;][]{goodfellow_deep_2016}.
-Despite that, a system that uses convolution would quickly reach its limitation in the musical domain.
-Here, knowledge of the notes played from the beginning of the song is needed, meaning that we need an architecture that is able to account for long-term dependencies - which brings us back to the RNN.
+Despite that, a system that uses convolution would quickly reach its limitation.
+Using music (the topic of this paper) as an example, knowledge of the notes played from the beginning of the song is needed.
+This means that we need an architecture that is able to account for long-term dependencies - which brings us back to the RNN.
 
 The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and hidden activations from the previous timestep.
-It should be noted that the parameters of a recurrent network (weights) are shared across different timesteps \parencite{graves_supervised_2012}.
+It should be noted that the parameters (weights) of a recurrent network are shared across different timesteps \parencite{graves_supervised_2012}.
 This allows extending (and applying) the model to data points of different lengths whilst generalizing across all of them \parencite{goodfellow_deep_2016}.
@@
 
-Thus, a formal definition of the network's forward pass would be
-\begin{equation}
-    h^t = f(h^{t-1}; x^t; \theta) \label{eq:rnn_hidden}
-\end{equation}
-where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the model parameters.
-
-The backward pass in a RNN is similar to that of a feedforward network: given the partial derivatives of a differentiable loss function $\mathcal{L}$ we calculate the derivatives with respect to the network's weights \parencite{graves_supervised_2012}.
-In this setting, a popular algorithm is called backpropagation through time \parencite[BPTT;][]{werbos_backpropagation_1990} which consists of repeated applications of the chain rule.
-The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer at the next timestep \parencite{graves_supervised_2012}.
-
-Equation \ref{eq:rnn_hidden} can be drawn as in Figure \ref{fig:rnn_folded}: input sequences $x$ are mapped to a hidden layer $h$ and an output $y$, with the state of the hidden layer at time $t$ feeding back into itself with a time delay of $1$.
-In other words, the state of $h_t$ will influence the decision of the recurrent layer at time $t+1$.
-Another way to illustrate this is to unfold the computational graph such that each component of the network is represented by different variables per step, as in Figure \ref{fig:rnn_unfolded}.
-
-The reason why these notions have been introduced is to be able to bring up the following point.
-As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}.
-\textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing backwards in time tend to either (1) blow up or (2) vanish.
-As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.
-
-\begin{figure}[t]
+\begin{figure}[htbp]
+    \begin{subfigure}{\linewidth}
         \centering
@@
         \end{tikzpicture}
-    \caption{Example of a feedforward neural network with three nodes in the input layer ($x_1$, $x_2$, $x_3$), three nodes in the hidden layer ($h_1$, $h_2$, $h_3$), and an output layer a single node ($y$).}
+        \caption{Feedforward Network}
         \label{fig:feedforward_net}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
+        \vspace{1em}\break
+    \end{subfigure}
     \begin{subfigure}{.24\linewidth}
@@
         \end{tikzpicture}
-        \caption{Folded}
+        \caption{Folded RNN}
         \label{fig:rnn_folded}
@@
         \end{tikzpicture}
-        \caption{Unfolded}
+        \caption{Unfolded RNN}
         \label{fig:rnn_unfolded}
     \end{subfigure}
-    \caption{A typical view of an RNN (left) and an unfolded view of the same network (right).
-In both sub-figures each node is a layer of network units however, in the second representation these are shown for each timestep.
+
+    \caption{Example of the two main ANN structures: a feedforward neural network (top), and a RNN (bottom).
+The feedforward neural network features three nodes in the input layer ($x_1$, $x_2$, $x_3$), three nodes in the hidden layer ($h_1$, $h_2$, $h_3$), and an output layer with a single node ($y$).
+On the bottom left we have a typical view of a RNN followed by an unfolded view of the same network (bottom right).
+In both sub-figures at the bottom each node is a layer of network units however, in the second representation these are shown for each timestep.
 The weighted connections from the input layer to the hidden one is labelled $U$, those from the hidden layer feeding back to itself are $W$, and those from the hidden layer to the output layer are labelled as $V$.
@@
 Bias has been omitted in this representation.
-Figure adapted from \textcite{graves_supervised_2012}}
+RNN figure adapted from \textcite{graves_supervised_2012}}
+    \label{fig:two_nets}
 \end{figure}
 
-The LSTM is made of a network of recurrently connected subnetworks known as memory blocks, each with one or more self-connected memory cells and three multiplicative units: the input gate, the output gate, and the forget gate \parencite{hochreiter_long_1997}.
-The three gates act as analogies for the read, write, reset computer operations and allow the memory cells to store and access information over long periods of time \parencite{graves_supervised_2012}.
-See Figure \ref{fig:lstm} for an illustration of a LSTM memory block with a single cell.
-
-Aside from being successful at a variety of classification tasks \parencite[e.g.,][]{graves_framewise_2005}, this architecture has been successfully applied to generating new data points \parencite[such as creating sequences of text;][]{graves_generating_2013}.
-In the context of generating musical performances, experiments with this neural architecture go as far as \textcite{eck_first_2002} however, a more noteworthy implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 cells each designed to generate classical compositions whilst maintaining the playing style of expert pianists.
-Considering its success, this model was re-implemented for the purposes of this paper's experiments (see Chapter \ref{chap:methods}).
+Thus, a formal definition of the network's forward pass would be
+\begin{equation}
+    h^t = f(h^{t-1}; x^t; \theta) \label{eq:rnn_hidden}
+\end{equation}
+where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the model parameters.
+
+The backward pass in a RNN is similar to that of a feedforward network: given the partial derivatives of a differentiable loss function $\mathcal{L}$ we calculate the derivatives with respect to the network's weights \parencite{graves_supervised_2012}.
+In this setting, a popular algorithm is called backpropagation through time \parencite[BPTT;][]{werbos_backpropagation_1990} which consists of repeated applications of the chain rule.
+The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer at the next timestep \parencite{graves_supervised_2012}.
+
+Equation \ref{eq:rnn_hidden} can be drawn as in Figure \ref{fig:rnn_folded}: input sequences $x$ are mapped to a hidden layer $h$ and an output $y$, with the state of the hidden layer at time $t$ feeding back into itself with a time delay of $1$.
+In other words, the state of $h_t$ will influence the decision of the recurrent layer at time $t+1$.
+Another way to illustrate this is to unfold the computational graph such that each component of the network is represented by different variables per step, as in Figure \ref{fig:rnn_unfolded}.
+
+The reason why these notions have been introduced is to be able to bring up the following point.
+As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}.
+\textcite{hochreiter_long_1997} outline two problems caused by BPTT: the error signals flowing backwards in time tend to either (1) blow up or (2) vanish.
+As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.
 
@@
 
+The LSTM is made of a network of recurrently connected subnetworks known as memory blocks, each with one or more self-connected memory cells and three multiplicative units: the input gate, the output gate, and the forget gate \parencite{hochreiter_long_1997}.
+The three gates act as analogies for the read, write, reset computer operations and allow the memory cells to store and access information over long periods of time \parencite{graves_supervised_2012}.
+See Figure \ref{fig:lstm} for an illustration of a LSTM memory block with a single cell.
+
+Aside from being successful at a variety of classification tasks \parencite[e.g.,][]{graves_framewise_2005}, this architecture has been successfully applied to generating new data points \parencite[such as creating sequences of text;][]{graves_generating_2013}.
+In the context of generating musical performances, experiments with this neural architecture go as far as \textcite{eck_first_2002} however, a more noteworthy implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 cells each designed to generate classical compositions whilst maintaining the playing style of expert pianists.
+Considering its success, this model was re-implemented for the purposes of this paper's experiments (see Chapter \ref{chap:methods}).
+
 \section{Generative Adversarial Networks} \label{sec:generative}
@@
 To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$.
-The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar.
+The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar \parencite{goodfellow_generative_2014}.
 More concretely, $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}.
@@
 Although new, GANs are a significant conceptual and technical innovation \parencite{briot_deep_2017} which can be seen in them revitalizing interest in generative models.
-As mentioned before, they have been proven successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}.
+They have been shown successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}.
 Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by successfully re-purposing this architecture to generate music.
-The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a unidirectional LSTM network with a similar structure as discriminator.
-
-\section{Learning to learn} \label{sec:learn2learn}
+The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a bidirectional LSTM network with a similar structure as discriminator.
+
+\section{Meta-Learning} \label{sec:learn2learn}
 
@@
 A simplistic way to do this would be to train a system on different tasks (e.g, image classification, text classification, etc.).
-However, this approach entails training on a dataset that is much larger that what current hardware can withstand.
+However, this approach entails training on a dataset that is much larger than what current hardware can withstand.
 Furthermore, this approach would stray away from the idea of building general-purpose A.I.
@@
 Although ideas of meta-learning can go as far as Schmidhuber's thesis in 1987, it seems that training a model using a handful of examples did not get much traction until \textcite{lake_human-level_2015} successfully designed a probabilistic learning framework capable of generalizing and learning a large number of concepts from a single class example (i.e., one-shot learning).
-For clarity, such a problem is typically defined as a $n$-shot $k$-way problem, where $n$ stands for the number of examples per class, and $k$ is the number of classes to be used in the training process.
+For clarity, such a problem is typically defined as a $n$-shot $k$-way problem, where $n$ stands for the number of examples per class, and $k$ is the number of classes.
 
 Initially, it was thought that ANNs are incompatible with few-shot learning since the iterative nature of gradient-based optimization algorithms do not work well in the context of a set number of updates \parencite{ravi_optimization_2016} - and even more so when there are only a few examples to work with.
-Having said that, \textcite{rezende_one-shot_2016} were among the first to provide a solution to the one-shot problem in the context of a neural network by implementing Bayesian reasoning into a deep network embedded within hierarchical latent variable models.
+Having said that, \textcite{rezende_one-shot_2016} were among the first to provide here a solution to the one-shot problem by implementing Bayesian reasoning into a deep network embedded within hierarchical latent variable models.
 Since that, numerous efforts were made to develop an algorithm that enables an ANN to perform with strictly limited datasets.
-Probably the most popular algorithm at the time of writing is the Model-Agnostic Meta-Learning (MAML) algorithm \parencite{finn_model-agnostic_2017}.
-One of MAML's core mechanics is that the parameters of the model are trained with a small number of gradient steps on with little training data to arrive at a good generalization on that task.
+Probably the most popular approach at the time of writing is the Model-Agnostic Meta-Learning (MAML) algorithm \parencite{finn_model-agnostic_2017}.
+One of MAML's core mechanics is that the parameters of the model are trained with a small number of gradient steps with little training data to arrive at a good generalization of that task.
 However, this procedure requires calculating second-order derivatives which implies a substantial computational penalty - something which the authors did note.
@@
 
-In essence, Reptile works (see Algorithm \ref{alg:reptile}) as follows.
+In essence, Reptile (see Algorithm \ref{alg:reptile}) works as follows.
 A model with parameters $\theta$ is initialized, and a meta-loop (or meta-epochs) of $i$ iterations starts.
@@
 To reiterate from Section \ref{sec:thesis_goal}, the purpose of this paper is to establish (1) the extent to which the music generated under a few-shot learning algorithm is comparable to that of a model trained on the entire dataset, and (2) the extent to which the music from such a model is comparable to that of real musical performances.
-For these purposes, Performance-RNN and C-RNN-GAN were adapted to the Reptile training procedure (Algorithm \ref{alg:reptile}) and evaluated against the (1) songs generated by the baseline proposed in \textcite{larochelle_few-shot_2017} - a single layer LSTM with 256 units - and (2) actual musical performances from the dataset (Section \ref{sec:dataset}).
+For these purposes, Performance-RNN and C-RNN-GAN were adapted to the Reptile training procedure (Algorithm \ref{alg:reptile}) and evaluated against (1) the songs generated by the baseline proposed in \textcite{larochelle_few-shot_2017} - a single layer LSTM with 256 units - and (2) actual musical performances from the dataset (Section \ref{sec:dataset}), after the data has been transformed into a more suitable format (Section \ref{sec:data_proc}).
 For an in-depth explanation on the steps required to develop a model that is able to generate new data points refer to Section \ref{sec:how_to_gen} and to Section \ref{sec:reptile_train} for details on the training procedure.
@@
 
-The dataset for this research will be similar to that of \parencite{oore_this_2018} namely, the MAESTRO Dataset \parencite{hawthorne_enabling_2018}.
+The dataset for this research will be similar to that of \textcite{oore_this_2018} namely, the MAESTRO Dataset \parencite{hawthorne_enabling_2018}.
 It consists of recorded MIDI data collected from each installment of the International Piano-e-Competition.
@@
 \begin{itemize}
-    \item No composition should appear in more than one split
+    \item No composition should appear in more than one split.
     \item Training set is approx 80\% of the dataset (in time), and the remaining is split equally on between the validation and test sets.
 Where possible, these proportions are true also within each composer.
-    \item Popular compositions are in the training set
-    \item The validation and test splits should maintain a variety of compositions
+    \item Popular compositions are in the training set.
+    \item The validation and test splits should maintain a variety of compositions.
 \end{itemize}
@@
 Having said that, the dataset lacks clearly defined classes.
-The canonical composer could be used as a substitute however, the author strongly believes that there is a better way to categorize music: by genre.
+Although not entirely necessary, having them in the dataset should enable the Reptile models to generate samples that are more analogous to a specific class instead of a mixture.
+Given what the dataset metadata provides, the canonical composer could be used.
+However, the author strongly believes that there is a better way to categorize music: by genre.
 
@@
 \section{Data Processing} \label{sec:data_proc}
-
-\begin{figure}[t]
-    \centering
-    \begin{verbatim}
-<note_on note=69 velocity=73 time=33>
-<note_on note=71 velocity=78 time=15>
-<control_change control=64 value=83 time=27>
-<note_off note=69 velocity=0 time=6>
-<note_off note=71 velocity=0 time=3>
-    \end{verbatim}
-    \caption{Example of MIDI messages in the dataset, as printed by the Python library Mido \parencite{bjorndalen_mido_2018}.
-At 33 ticks since the last event, the player pressed an A4 (note 69), followed by a B4 (note 71), 15 ticks later.
-Next, the pedal (Control Change message with a value of 64) is pressed 27 ticks later, and the keys are released 6 and 3 tick later.}
-    \label{fig:midi_examples}
-\end{figure}
 
@@
 A glance over the MIDI documentation \parencite{midi_association_official_2019} shows how complex and granular the standard is.
-Luckily however, we do not need to implement all of the messages into our model in order to make it work.
+However, we do not need to implement all of the messages into our model in order to make it work.
 We only need three: \texttt{note\_on}, \texttt{note\_off}, and \texttt{pedal\_on/off}.
-Figure \ref{fig:midi_examples} show an example of how the midi messages look like.
-It should be noted that in this figure, the time is represented in ticks, MIDI's smallest unit of time, who's value is usually a third of a beat \parencite{midi_association_official_2019}.
+Figure \ref{fig:midi_examples} show an example of what the MIDI messages look like.
+It should be noted that in this figure, the time is represented in ticks, MIDI's smallest unit of time \parencite{midi_association_official_2019}, which in the dataset is between one 384$^{th}$ and one 480$^{th}$ of a beat.
 
@@
 
-\begin{figure}[t]
+\begin{figure}
+    \centering
+    \begin{verbatim}
+<note_on note=69 velocity=73 time=33>
+<note_on note=71 velocity=78 time=15>
+<control_change control=64 value=83 time=27>
+<note_off note=69 velocity=0 time=6>
+<note_off note=71 velocity=0 time=3>
+    \end{verbatim}
+    \caption{Example of how MIDI messages are printed on the screen.
+At 33 ticks since the last event, the player pressed an A4 (note 69), followed by a B4 (note 71), 15 ticks later.
+Next, the pedal (Control Change message with a value of 64) is pressed 27 ticks later, and the keys are released 6 and 3 tick later.}
+    \label{fig:midi_examples}
+    \vspace{1em}
+%\end{figure}
+
+%\begin{figure}[h]
     \begin{tikzpicture}
@@
 An alternative to this would be to define a noise vector $z^t$ of an arbitrary size (e.g., 100), and a random input $x^t$ selected from our events vocabulary.
-The noise vector is passed through a linear layer that output a tensor that would serve as the network's hidden state $h^{t}$ at time $t$.
+The noise vector is passed through a linear layer that outputs a tensor that would serve as the network's hidden state $h^{t}$ at time $t$.
 It is then that $x^t$ and $h^{t}$ are fed into the LSTM layer to generate $x^{t+1}$ and $h^{t+1}$.
@@
 For instance, if the first input would be a C\#, the network could learn to always output E as the next note.
-Using a random hidden state is a measure that was placed to ensure some diversity in the process of generating samples (granted, it should be admitted the possibility of this scenario being unlikely, and that the regular method described at the beginning of this section would still work just as good).
+Using a random hidden state is a measure that was implemented to ensure some diversity in the process of generating samples (granted, it should be admitted the possibility of this scenario being unlikely, and that the regular method described at the beginning of this section would still work just as well).
 
@@
 \begin{equation}
-    x^{t+1} = X(\omega)
+    x^{t+1} \sim X
 \end{equation}
-where
+where $X$ is the predicted event distribution given by the network's output divided by our temperature $\lambda$:
 \begin{equation}
@@
 \end{equation}
-$X$ is the predicted event distribution of the network's output divided by our temperature $\lambda$.
 Thus, $x^{t+1}$ is actually a random sample from our defined categorical distribution $X$.
@@
 With the generative process outlined in the previous section, training the generative networks presented in this papers is somewhat straightforward.
-A batch of songs is sampled from the training set, the model generates as many songs as there are in the batch, and then compute the loss between the network's output and the real samples.
+A batch of songs is sampled from the training set, the model generates as many songs as there are in the batch, and then computes the loss between the network's output and the real samples.
 This statement is essentially true for the baseline and Performance-RNN (in spite of the Reptile training procedure, inner-loop training is similar to the standard procedure).
-In contrast, the GAN model requires training two models simultaneously, where the generator's loss function is based on whether the discriminator is fooled by the fake samples (i.e., classifies them as real), and the discriminator's loss is based on whether it can distinguish between real and fake samples (it should be mentioned that both components of the GAN would use binary cross-entropy as their loss function; in contrast the baseline and Performance-RNN are based on cross entropy).
-However, there are still several items that we haven't touched upon.
-
-Firstly, all three networks have been trained with teacher forcing, a procedure in which during training the model receives the ground truth $y^t$ as input at time $t+1$.
+In contrast, the GAN model requires training two models simultaneously, where the generator's loss function is based on whether the discriminator is fooled by the fake samples (i.e., classifies them as real), and the discriminator's loss is based on whether it can distinguish between real and fake samples.
+It is worthwhile mentioning that both the baseline and Performance-RNN use cross-entropy as the loss function whereas the GAN components use binary cross-entropy.
+The reason for this difference is because the aim of $D$ is to correctly classify input samples $x$ and $G(z)$ as "real" and "fake", whereas $G$ must minimize the probability of $D$ classifying $G(z)$ as "fake" (and this is done by accessing $D(G(z))$).
+Having said all of these, there are still several items that we haven't touched upon.
+
+Firstly, all three networks have been trained with teacher forcing, a procedure where - during training - the model receives the ground truth $y^t$ as input at time $t+1$.
 Adopting this measure allows the model to create hidden-to-hidden connections and capture the necessary information from the history of input sequences \parencite{goodfellow_deep_2016}.
@@
 Specifically, we train the generator to match the expected value of those features \parencite{salimans_improved_2016}.
-However, this measure was not implemented in this paper as early tests caused both the loss of both the generator and discriminator to drop at zero, causing both models to stop learning.
-
-It should be noted that now, the loss functions of both the generator and discriminator differ in scale, meaning that we can no longer compare them to freeze training of either models: the generator uses MSE, whereas the discriminator use binary cross-entropy.
-A workaround to this is to calculate the both losses, but use MSE for backpropagation and optimization, whilst binary cross-entropy is used to compare our models.
+However, this measure was not implemented in this paper as early tests caused both the loss of both the generator and discriminator to drop at zero, causing either models to stop learning.
 
@@
 Another aspect that needs addressing is the number of epochs.
-All models were trained for 250 epochs each however, when looking at the way training is handled in Reptile (see Algorithm \ref{alg:reptile}), this statement is inconclusive.
+All models were trained for 250 epochs each.
+However, when looking at the way training is handled in Reptile (see Algorithm \ref{alg:reptile}), this statement is inconclusive.
 Because the algorithm presents us with a meta-loop and an inner-loop, epochs is defined in this setting as the number of times parameters $\theta$ have been updated.
@@
 
-\begin{table}[ht]
+\begin{table}[hb]
     \centering
-    \begin{tabular}{l|c|c|c}
+    \begin{tabular}{l|r|r|r}
                     &\bf Baseline   &\bf Performance RNN&\bf C-RNN-GAN  \\ \hline
@@
 The more domain-specific evaluation measures are adapted from \textcite{mogren_c-rnn-gan_2016}.
-Polyphony measures how often a minimum of two tones are played simultaneously (when the start time is the same).
+Polyphony measures how often a minimum of two tones are played simultaneously.
+Because playing two notes at the exact same time would be rare occurrence, this metric would consider as played simultaneously notes which are at most 20.7 milliseconds apart; which is the mean time needed to construct a three-notes chord, as measured by \textcite{sandnes_human_2015}.
 It should be noted that this is a restrictive metric: it can give a low score to music where the two notes start at different times.
 Scale consistency is the fraction of notes that are part of a standard scale (e.g., major, minor, lydian, etc.; see Appendix \ref{appendix:scales} for a complete list with examples), and reporting the results for the best matching scale.
-Repetitions counts consecutive subsequences of notes, and tone span is the difference between the lowest note and highest note \parencite[counted in semi-tones;][]{mogren_c-rnn-gan_2016}.
-
-The choice of NDB comes from the fact that it is good at detecting overfitting \parencite{borji_pros_2018}, whereas the domain-specific metrics should favor models that generate high fidelity samples.
+Repetitions counts when the same note is played consecutively, and tone span is the difference between the lowest note and highest note \parencite[counted in semi-tones;][]{mogren_c-rnn-gan_2016}.
+
+The choice of NDB comes from the fact that it is good at detecting overfitting \parencite[i.e., when NDB is low;][]{borji_pros_2018}, whereas the domain-specific metrics should favor models that generate high fidelity samples.
 In addition, these evaluation metrics have the benefit of being model-agnostic; they do not require a specific type of generative model.
@@
 
-The results in Table \ref{tab:results} underline several things about the trained models, the most important one being that all three lack the concept of motif (the presence of a dominant/recurring theme within a piece).
+The results in Table \ref{tab:results} underline several aspects about the trained models, the most important one being that all three produce somewhat disorganized music.
 The clear giveaway sign is the substantial lack of repetition in the generated samples.
 This is enforced by the larger levels of polyphony in the three models when compared to the test set.
-In addition, all models seems to be unrestrained with respect to using the whole gamut of notes they have available, but considering that a piano keyboard has 88 keys and the MIDI standard allows for 128 values (notes), it seems that our models might be heading into the right direction.
-When it comes to playing in a scale, all models show are equal across the board, with extremely minor variations.
-Having said these, the domain-specific evaluation metrics indicate that the models are fairly close to eachother but that should not change the fact that there is still room for improvement when compared to the real samples.
-
+In addition, all models seems to be unrestrained with respect to using the whole gamut of notes available, but considering that a piano keyboard has 88 keys and the MIDI standard allows for 128 values (notes), it seems that our models might be heading into the right direction.
+When it comes to playing in a scale, all models show that they are equal across the board, with extremely minor variations in their adherence.
+In light of these, the domain-specific evaluation metrics indicate that the models are fairly close to eachother but that should not change the fact that there is still room for improvement when compared to the real samples.
 Having said that, the landscape changes a bit when looking at NDB, our other evaluation metric.
-With seven to ten bins statistically different (out of 20), our models seem to display a great deal of disparity form the test samples, even though they somewhat agree between themselves - i.e., they don't seem to be very different from each other.
+With seven to eleven bins statistically different (out of 20), our models seem to display a great deal of disparity from the test samples.
 
 With these result we have gathered enough information to answer our research questions, and claim that the baseline still produces some of the best results (as it scores best on 4/5 metrics).
-However, a deep-dive into the performance at training time would shed some more light to these results.
+However, a deep-dive into the performance at training time would shed some light as to why this is happening.
 
@@
     \centering
-\begin{tabular}{l|l|l|l|l|l|l}
+\begin{tabular}{l|r|r|r|r|r|r}
     Model             & Temperature & Polyphony (\%)  & Repetitions       & Tone Span        & Scale Consistency (\%) & NDB     \\ \hline
-\textit{Test Set} & N/A         & \textit{49.15} & \textit{814.33} & \textit{66.9440} & \textit{68.27}   & N/A     \\
-Baseline          & 0.8         & 57.07          & \textbf{\color{base0D}388.02}    & \textbf{\color{base0D}78.45}          & \textbf{\color{base0D}67.70}            & 11      \\
-Baseline          & 0.9         & 60.54          & 373.06          & 81.31          & 67.47            & 10      \\
-Baseline          & 1.0         & 61.12          & 347.36          & 83.06          & 67.28            & 10      \\
-Baseline          & 1.1         & 62.92          & 334.42          & 84.44          & 67.22            & \textbf{\color{base0D}7} \\
-Baseline          & 1.2         & 60.60          & 316.82          & 85.05          & 67.26            & 9       \\
-Performance-RNN   & 0.8         & 56.84          & \textbf{\color{base0A}377.82}          & 78.84          & 67.60            & 9       \\
-Performance-RNN   & 0.9         & 58.52          & 364.97          & 81.49          & 67.57            & 8       \\
-Performance-RNN   & 1.0         & 62.19          & 354.08          & 83.24          & 67.30            & 9       \\
-Performance-RNN   & 1.1         & 63.42          & 343.22          & 84.46          & 67.21            & 9       \\
-Performance-RNN   & 1.2         & 60.61          & 312.25          & 85.24          & 67.23            & 10      \\
-C-RNN-GAN         & 0.8         & \textbf{\color{base0E}56.59}    & 375.25          & \textbf{\color{base0A}78.82}    & \textbf{\color{base0A}67.63}      & 9       \\
-C-RNN-GAN         & 0.9         & 58.41          & 365.19          & 81.22          & 67.50            & 9       \\
-C-RNN-GAN         & 1.0         & 61.25          & 351.02          & 83.00          & 67.40            & \textbf{\color{base0A}8}       \\
-C-RNN-GAN         & 1.1         & 63.49          & 338.39          & 84.66          & 67.29            & 11      \\
-C-RNN-GAN         & 1.2         & 60.57          & 316.35          & 85.30          & 67.19            & 10     
+\textit{Test Set} & N/A         & \textit{41.94} & \textit{814.33} & \textit{66.9440} & \textit{68.27}   & N/A     \\
+Baseline          & 0.8         & 49.65          & \textbf{\color{base0D}388.02}    & \textbf{\color{base0D}78.45}          & \textbf{\color{base0D}67.70}            & 11      \\
+Baseline          & 0.9         & 52.19          & 373.06          & 81.31          & 67.47            & 10      \\
+Baseline          & 1.0         & 51.91          & 347.36          & 83.06          & 67.28            & 10      \\
+Baseline          & 1.1         & 52.59          & 334.42          & 84.44          & 67.22            & \textbf{\color{base0D}7} \\
+Baseline          & 1.2         & 49.71          & 316.82          & 85.05          & 67.26            & 9       \\
+Performance-RNN   & 0.8         & 49.59          & \textbf{\color{base0A}377.82}          & 78.84          & 67.60            & 9       \\
+Performance-RNN   & 0.9         & 50.24          & 364.97          & 81.49          & 67.57            & 8       \\
+Performance-RNN   & 1.0         & 52.86          & 354.08          & 83.24          & 67.30            & 9       \\
+Performance-RNN   & 1.1         & 53.05          & 343.22          & 84.46          & 67.21            & 9       \\
+Performance-RNN   & 1.2         & 49.73          & 312.25          & 85.24          & 67.23            & 10      \\
+C-RNN-GAN         & 0.8         & \textbf{\color{base0E}49.17}    & 375.25          & \textbf{\color{base0A}78.82}    & \textbf{\color{base0A}67.63}      & 9       \\
+C-RNN-GAN         & 0.9         & 50.23          & 365.19          & 81.22          & 67.50            & 9       \\
+C-RNN-GAN         & 1.0         & 51.99          & 351.02          & 83.00          & 67.40            & \textbf{\color{base0A}8}       \\
+C-RNN-GAN         & 1.1         & 53.12          & 338.39          & 84.66          & 67.29            & 11      \\
+C-RNN-GAN         & 1.2         & 49.73          & 316.35          & 85.30          & 67.19            & 10     
 \end{tabular}
-\caption{Results of the evaluation process of the generated sample.
-For comparison, relevant statistics of the test samples have been included.
-Important values are in bold.
-The values which are closest to the test sample are in green.
-Values which are closest to the baseline's best score are in red.
-Values which are both close to the baseline's best and the test sample are in purple.}
+\caption{Evaluation results of the generated samples.
+Test set statistics are in italics.
+Values closest to the test set are in green.
+Values closest to the baseline's best score are in red.
+Values which match both the baseline's best and the test sample are in purple.}
 \label{tab:results}
+\end{table}
+
+\begin{table}[b]
+    \centering
+\begin{tabular}{l|r|r|r|r|r|r}
+    Model             & Temperature & Polyphony (\%)  & Repetitions       & Tone Span        & Scale Consistency (\%) & NDB     \\ \hline
+Baseline (50) & 0.8 & 37.75 & 599.94 & \textbf{\color{base0D}73.72} & \textbf{\color{base0D}68.06} & 10 \\
+Baseline (50) & 0.9 & 46.02 & 897.23 & 77.43 & 67.69 & \textbf{\color{base0D}7}  \\
+Baseline (50) & 1.0 & 43.63 & \textbf{\color{base0D}790.02} & 79.67 & 67.54 & 10 \\
+Baseline (50) & 1.1 & 43.79 & 630.93 & 80.95 & 67.48 & 9  \\
+Baseline (50) & 1.2 & \textbf{\color{base0D}40.33} & 536.90 & 82.12 & 67.32 & 10
+\end{tabular}
+\caption{Results of the baseline model after 50 epochs. Values in green closely match the the test set (Table \ref{tab:results})}
+\label{tab:base50}
 \end{table}
@@
 
-When looking at Figure \ref{fig:loss} we gain an insight on how well the training has gone.
-First and foremost, the loss of Performance-RNN is as one would expect: the model starts somewhat poorly, and then incrementally tunes itself to perform better (Figure \ref{fig:loss_performancernn}).
-Then, we have the losses of the GAN (Figure \ref{fig:loss_crnngan}) which portrays the adversarial training the generator and discriminator.
+When looking at Figure \ref{fig:loss} we gain an insight into the training.
+First and foremost, the loss of Performance-RNN is as expected: the model starts somewhat poorly, and then incrementally tunes itself to perform better (Figure \ref{fig:loss_performancernn}).
+Then, we have the losses of the GAN (Figure \ref{fig:loss_crnngan}) which portrays the adversarial training of the generator and discriminator.
 Although on a first glance one might think that the loss is subject to high variance, it should be noted that a different function was used in this scenario and that the plot is on a much smaller scale than all others.
@@
 The loss drops quickly, only to jump back to its starting position after $\approx$50 epochs.
-A simple - and most likely true - explanation is that the learning rate is too big for the model.
-Considering the granularity with which all models have been trained, it could be that the model has escaped a good local minima due to the high learning rate.
+A simple - and most likely true - explanation is that the learning rate is too big for the model; it could be that the model has escaped a good local minima due to the high learning rate.
 This should not be cause for any concern as further training would bring back the network to a similar position (it should probably reminded to the reader that the experiments in this papers are done with only 250 iterations, whereas more sophisticated DL experiments are somewhere in the order of thousands of iterations).
+But, to be sure that the model did not miss a good local minima, the baseline was retrained for 50 epochs - the results of which can be found in Table \ref{tab:base50}.
+It seems that the 50 epochs baseline performs much better now, significant gains being in polyphony and repetitions.
 
@@
 To better illustrate this, Figure \ref{fig:acts_scaled} shows the two applications on the same scale.
-Secondly, that C-RNN-GAN (Figure \ref{fig:crnngan_activations}) still has a long way to go until it can be considered adequate at generating a song; in other words, C-RNN-GAN generates songs mostly at random, claim enforced by the fact that almost all neurons are activated and that the pattern of generating a sequence has no order.
+Secondly, that C-RNN-GAN (Figure \ref{fig:crnngan_activations}) still has a long way to go until it can be considered adequate at generating a song; to be more precise, C-RNN-GAN generates songs mostly at random, claim enforced by the fact that almost all neurons are activated and that the pattern of generating a sequence has no order.
 
@@
 From the lens of mainstream media, we are at a point where we will soon create what is commonly known as "Artificial Intelligence" \parencite{levy_ai_2010}.
-This claim, of course, ignores the popularity cycles of conectionism dating back to the fifties \parencite[e.g.,][]{rosenblatt_perceptron_1957}, or the over-inflated expectations for autonomous human-made systems that were made ever since \parencite{marcus_is_2012}.
-However, even after suffering waves of funding-cuts, masses of criticism, and regressions into obscurity, develpments were still being made \parencite{knight_ai_2016, minsky_perceptrons_1988} - albeit at a smaller pace.
-With the reemergence of ANNs, we are now seeing even stronger developments towards creating general-purpose Artificial Intelligence \parencite[e.g.,][]{vinyals_starcraft_2017}, but that should not cause any relaxation or distress; "we are not there yet".
-
-In Chapter \ref{chap:intro}, I pointed out a couple of shortcomings that current ANN systems have (reliance on massive amounts of data, and inability to perform separate tasks) along with two exciting trends in Deep Learning: the resurgence of generative networks and successfully infusing a system with meta-learning capabilities).
+This claim, of course, ignores the popularity cycles of conectionism dating back to the fifties, or the over-inflated expectations for autonomous human-made systems that were made ever since \parencite{marcus_is_2012}.
+However, even after suffering waves of funding-cuts, masses of criticism, and regressions into obscurity, significant developments in A.I.
+were still being made \parencite{knight_ai_2016, minsky_perceptrons_1988}.
+With the reemergence of ANNs, we are now taking even bigger steps towards creating general-purpose Artificial Intelligence \parencite[e.g.,][]{vinyals_starcraft_2017}.
+But that should not cause any relaxation or distress as we are not "there" yet.
+
+In Chapter \ref{chap:intro}, I pointed out a couple of shortcomings that current ANN systems have (reliance on massive amounts of data, and inability to perform separate tasks) along with two exciting trends in Deep Learning: the resurgence of generative networks and successfully infusing a system with meta-learning capabilities.
 Acknowledging the initial attempts at intersecting these two fields, this paper proposed expanding those efforts to the musical domain.
-It set to develop two generative models by modelling the training procedure to be more akin to that of a human learning, and evaluated their performances against a baseline proposed by \textcite{larochelle_few-shot_2017} and their resemblance to real samples.
+It set to develop two generative models by modelling the training procedure to be more akin to that of human learning, and evaluated their performances against a baseline proposed by \textcite{larochelle_few-shot_2017} and their resemblance to real samples.
 Concretely, Performance-RNN \parencite{oore_this_2018} and C-RNN-GAN \parencite{mogren_c-rnn-gan_2016} were trained under the Reptile learning framework \parencite{nichol_first-order_2018}, and then evaluated against a LSTM network with one layer and 256 hidden units as well as expert piano performances extracted from the MAESTRO dataset \parencite{hawthorne_enabling_2018}.
@@
 However, this does not change the fact that it was closely followed by the Reptile models.
-More appropriately is to point that, in spite of being trained to generate music using a handful of examples, the Reptile models closely matched the performance of the baseline.
-This especially holds true for this paper's reimplementation of Performance-RNN whose neuron activations show that it has indeed learned how to generate music, even though its samples scored somewhere in the middle.
-The adversarial model showed that it is still somewhere in the early stages of learning, which would explain why almost all of its neurons were fired but, the fact that it performed reasonably in its evaluation should not be ignored either.
-Having said these, I say with confidence that when tasked with making music, Reptile-based generative models can be quite competitive.
+More appropriately is to point that in spite of being trained to generate music using a handful of examples, the Reptile models closely matched the performance of the baseline.
+This especially holds true for this paper's reimplementation of Performance-RNN whose neuron activations show that it has indeed learned how to generate samples, even though it scored somewhere in the middle.
+The adversarial model showed that it is still somewhere in the early stages of learning, which would explain why almost all of its neurons fired but, the fact that it performed reasonably well should not be ignored either.
+Having said these, I say with confidence that when tasked with making music, the Reptile-based generative models can be quite competitive.
 
 As well as they perform, both few-shot models failed to produce samples that could be passed as real.
-However, neither did the baseline, which would indicate that with more time invested into training, all three models could generate samples that are both novel and high fidelity.
+However, neither did the original baseline, whereas the adjusted baseline has shown impressive boosts in performance.
+This finding indicates that with proper hyperparameter tuning, either models could generate songs which are both novel and high fidelity.
 
@@
 Optimally, all models should be trained for over 1,000 epochs however, time and hardware constrains meant taking some shortcuts.
-This constraint is especially seen in the case of the baseline which has been trained with a batch size 32 times larger than the Reptile models (to drive down the training time from 250 hours to 27).
+This is especially visible in the baseline which has been trained with a batch size 32 times larger than the Reptile models (to drive down the training time from 250 hours to 27, at the risk of making the models incomparable).
 In addition, it is likely that the training of the GAN model suffered from the fact that a measure for combating mode-collapse was not implemented.
-Had it been done so, then it's possible that the model would have had better activations and consequently, better samples.
-
-Next, the generating process from which we sample with temperature to generate the next likely event is a somewhat greedy process \parencite{holtzman_curious_2019}.
+Had it been done so, then it's possible that the model would have had more thorough activations and consequently, better samples.
+
+Secondly, this paper adopts a quantitative evaluation of the generated samples.
+Considering that this is a highly debated topic \parencite{borji_pros_2018} it is likely that the metrics in this paper are not sufficient.
+Moreover, given the nature of the topic, a qualitative appraisal of the generated samples is crucial to a better understanding of the quality of the samples.
+
+Next, the generating process from which we sample with temperature to generate the next likely event is a greedy procedure \parencite{holtzman_curious_2019}.
 Assuming that the probability distribution of how humans make music borrows some of its characteristics from how we create language - which is likely since both tasks share a significant portion of brain activations \parencite{brown_music_2006} -, then greedy sampling or beam search (another sampling method) would completely fail to generate human-like music.
-Specifically, they are not ideal and nor adequate for a high entropy task \parencite{holtzman_curious_2019}.
-This could explain why the models in this paper lacked the concept of motif.
+Specifically, they are not ideal nor adequate for a high entropy task \parencite{holtzman_curious_2019}.
+This could explain why the models in this paper scored low in repetitions.
 Curiously enough, two recent solutions for sampling tokens have been shown to be better candidates: top-$k$ sampling \parencite{fan_hierarchical_2018} and nucleus sampling \parencite[also known as top-$p$ sampling;][]{holtzman_curious_2019}.
-Future work could incorporate either of these proposed sampling methods and evaluate them against the more traditional sampling methods in the context of generating music with a meta-learning algorithm.
+Future work could incorporate either of these proposed sampling methods and evaluate them against the more traditional ones in the context of generating music with a meta-learning algorithm.
 
@@
 Nevertheless, the results in this paper are intriguing in and of themselves.
-They not only point towards neural networks as a viable infrastructure for music generation, but also towards training such a system in a manner that is (loosely) more akin to the way a human being would.
-By this I exclusively refer to the quantity of examples involved in the generative process.
-More importantly, these results outline the possibility of integrating generative networks and few-shot learning into the production process of a musician.
+They not only point towards neural networks as a viable infrastructure for music generation, but also towards the viability of training such a system in a manner that is (loosely) more akin to the way a human being would.
+By this I exclusively refer to the quantity of examples involved in the learning process.
+Most importantly, these results outline the possibility of integrating generative networks and few-shot learning into the production process of a musician.
 To be more precise, \textcite{larochelle_few-shot_2017} indicate that instead of training a model on various types of music to influence the style of the generated samples, a musician could substitute the data from a support set with music of their own choice and obtain a new generative model for that particular style of music (admittedly, after a few more hours of training).
@@
 In this light, a tentative answer to the second research question is that the Reptile models can be similar to real music.
-Here, a significant area of improvement lies in embedding a sense of motif into the models (since this is what the models have show to lack the most).
+Here, a significant area of improvement lies in embedding a sense of musical structure into the models (since this is what the models have shown to lack the most).
 

### 2019-05-16 demo/main.py

@@ -1,0 +1,37 @@
+import os
+import wx
+from glob import glob
+
+
+class MainWindow(wx.Frame):
+    def __init__(self, parent, title):
+        wx.Frame.__init__(self, parent, title=title, size=(1248, 800))
+
+        self.init_params()
+        self.get_models()
+        self.create_controls()
+
+        self.Show(True)
+
+    def init_params(self):
+        self.models_dir = '../models/'
+
+    def get_models(self):
+        models = [i for i in os.walk(self.models_dir)][0][2]
+        self.models = [
+            i for i in models if not ('checkpoint' in i or 'csv' in i)
+        ]
+
+    def create_controls(self):
+        self.model_choice = wx.RadioBox(
+            self,
+            label='Choose a model',
+            choices=self.models,
+            majorDimension=4)
+        self.gen_button = wx.Button(self, label='Generate')
+
+
+if __name__ == '__main__':
+    app = wx.App(False)
+    frame = MainWindow(None, 'Demo Music Generator')
+    app.MainLoop()

### 2019-05-16 saved_models/unzip_all.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+unzip baseline.zip -d ../models/
+unzip baseline_50.zip -d ../models/
+
+cat performance_rnn_part_* > performance_rnn.zip
+unzip performance_rnn.zip -d ../models/
+
+cat c_rnn_gan_part_* > c_rnn_gan.zip
+unzip c_rnn_gan.zip -d ../models/
+
+rm performance_rnn.zip
+rm c_rnn_gan.zip

### 2019-05-16 src/evaluation.py

@@ -18,7 +18,7 @@ def open_midi(path):
     return melody, tpb, tempo
 
 
-def polyphony(melody, tpb, tempo, criterion=31.25):
+def polyphony(melody, tpb, tempo, criterion=20.7):
     # Default criterion is a 64th note in ms
     # At 120 BPM, a 64th is pretty fast, and two notes
     #  played at a distance of ~31ms sounds like a
@@ -144,7 +144,7 @@ def scale_consistency(melody, verbose=False):
         return max(scores)
 
 
-def evaluate_song(path, poly_criterion=31.25):
+def evaluate_song(path, poly_criterion=20.7):
     melody, tpb, tempo = open_midi(path)
     poly = polyphony(melody, tpb, tempo, criterion=poly_criterion)
     rep = repetitions(melody)
@@ -158,7 +158,7 @@ def evaluate_song(path, poly_criterion=31.25):
     }
 
 
-def evaluate_batch(paths, poly_criterion=31.35):
+def evaluate_batch(paths, poly_criterion=20.7):
     results = {
         'polyphony': [],
         'repetitions': [],

### 2019-05-16 paper/images/baseline/fix.sh
moved paper to a different repo

@@ -1,3 +1,0 @@
-for i in *.tex; do
-    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/baseline\//g' $i
-done

### 2019-05-16 paper/images/c_rnn_gan/fix.sh

@@ -1,3 +1,0 @@
-for i in *.tex; do
-    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/c_rnn_gan\//g' $i
-done

### 2019-05-16 paper/images/copy_from_plots.sh

@@ -1,11 +1,0 @@
-for i in ../../plots/baseline*; do
-    cp $i ./baseline/
-done
-
-for i in ../../plots/performance_rnn*; do
-    cp $i ./performance_rnn/
-done
-
-for i in ../../plots/c_rnn_gan*; do
-    cp $i ./c_rnn_gan/
-done

### 2019-05-16 paper/images/fix_images.sh

@@ -1,9 +1,0 @@
-cd ./baseline/
-./fix.sh
-cd ../
-cd ./performance_rnn/
-./fix.sh
-cd ../
-cd ./c_rnn_gan/
-./fix.sh
-cd ../

### 2019-05-16 paper/images/onehots/fix.sh

@@ -1,3 +1,0 @@
-for i in *.tex; do
-    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/onehots\//g' $i
-done

### 2019-05-16 paper/images/performance_rnn/fix.sh

@@ -1,3 +1,0 @@
-for i in *.tex; do
-    sed -i 's/ymax=-0.5] {/ymax=-0\.5] {images\/performance_rnn\//g' $i
-done

### 2019-05-17 demo/backend.py
finished demo app

@@ -1,0 +1,69 @@
+import os
+import torch
+import sys
+sys.path.insert(0, '../src')
+from generate import load_model_data
+from data import from_seq
+from utils import make_folder
+
+
+def generate_song(name, temperature=1.0, length=500):
+    checkpoint = load_model_data(os.path.join('../models/', name))
+    if 'baseline' in name:
+        from lstm_baseline import Baseline
+        net = Baseline(
+            in_dim=checkpoint['in_dim'],
+            hidden_dim=checkpoint['hidden_dim'],
+            batch_size=1,  # we create one song at a time
+            out_dim=checkpoint['out_dim'],
+            dropout=checkpoint['dropout'],
+            num_layers=checkpoint['num_layers'],
+            init_dim=checkpoint['init_dim'],
+            use_bias=checkpoint['use_bias'],
+            is_bidirectional=checkpoint['is_bidirectional'])
+        net.load_state_dict(checkpoint['model_state'])
+    elif 'performance_rnn' in name:
+        from performance_rnn import PerformanceRNN
+        net = PerformanceRNN(
+            in_dim=checkpoint['in_dim'],
+            hidden_dim=checkpoint['hidden_dim'],
+            batch_size=1,  # we create one song at a time
+            out_dim=checkpoint['out_dim'],
+            dropout=checkpoint['dropout'],
+            num_layers=checkpoint['num_layers'],
+            init_dim=checkpoint['init_dim'],
+            use_bias=checkpoint['use_bias'],
+            is_bidirectional=checkpoint['is_bidirectional'])
+        net.load_state_dict(checkpoint['model_state'])
+    elif 'c_rnn_gan' in name:
+        from c_rnn_gan import Generator
+        net = Generator(
+            in_dim=checkpoint['in_dim'],
+            hidden_dim=checkpoint['hidden_dim'],
+            batch_size=1,  # we create one song at a time
+            out_dim=checkpoint['out_dim'],
+            dropout=checkpoint['dropout'],
+            num_layers=checkpoint['num_layers'],
+            init_dim=checkpoint['init_dim'],
+            use_bias=checkpoint['use_bias'])
+        net.load_state_dict(checkpoint['g_model_state'])
+
+    device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
+    net.to(device)
+    init = torch.randn(net.batch_size, net.init_dim).to(device)
+    song = net.generate(
[…]

### 2019-05-17 demo/main.py

@@ -1,6 +1,8 @@
 import os
 import wx
+import wx.lib.agw.floatspin as FS
 from glob import glob
+import backend
 
 
 class MainWindow(wx.Frame):
@@ -9,7 +11,14 @@ class MainWindow(wx.Frame):
 
         self.init_params()
         self.get_models()
+        # self.pnl = wx.Panel(self)
+        self.sizer = wx.BoxSizer(wx.VERTICAL)
+
         self.create_controls()
+
+        self.bind_events()
+
+        self.set_layout()
 
         self.Show(True)
 
@@ -23,12 +32,85 @@ class MainWindow(wx.Frame):
         ]
 
     def create_controls(self):
+
+        self.load_sf_button = wx.Button(self, label='Load Soundfont')
+
         self.model_choice = wx.RadioBox(
+            self, label='Model', choices=self.models, majorDimension=4)
+
+        self.temperature = FS.FloatSpin(
             self,
-            label='Choose a model',
-            choices=self.models,
-            majorDimension=4)
+            value=1.0,
+            min_val=0.7,
+            max_val=2.0,
+            increment=0.1,
+            style=wx.SL_HORIZONTAL | wx.SL_LABELS)
+        self.temperature.SetFormat('%f')
+        self.temperature.SetDigits(2)
+
+        self.length = wx.Slider(
+            self,
+            value=300,
+            minValue=100,
+            maxValue=400,
+            style=wx.SL_HORIZONTAL | wx.SL_LABELS)
+
         self.gen_button = wx.Button(self, label='Generate')
+
+        self.play_button = wx.Button(self, label='Play Song')
+
+        self.txt_section = wx.StaticText(
+            self, label='Options', style=wx.ALIGN_CENTER)
+
+        self.txt_temp = wx.StaticText(
+            self, label='Temperature', style=wx.ALIGN_LEFT)
+
+        self.txt_len = wx.StaticText(
+            self, label='Song Length', style=wx.ALIGN_LEFT)
+
+    def bind_events(self):
+        self.Bind(wx.EVT_BUTTON, self.generate, self.gen_button)
+        self.Bind(wx.EVT_BUTTON, self.load_soundfont, self.load_sf_button)
+        self.Bind(wx.EVT_BUTTON, self.play_song, self.play_button)
+
+    def load_soundfont(self, e):
[…]

### 2019-05-17 demo/utils.py

@@ -1,0 +1,69 @@
+import os
+import torch
+import sys
+sys.path.insert(0, '../src')
+from generate import load_model_data
+from data import from_seq
+from utils import make_folder
+
+
+def generate_song(name, temperature=1.0):
+    checkpoint = load_model_data(os.path.join('../models/', name))
+    if 'baseline' in name:
+        from lstm_baseline import Baseline
+        net = Baseline(
+            in_dim=checkpoint['in_dim'],
+            hidden_dim=checkpoint['hidden_dim'],
+            batch_size=1,  # we create one song at a time
+            out_dim=checkpoint['out_dim'],
+            dropout=checkpoint['dropout'],
+            num_layers=checkpoint['num_layers'],
+            init_dim=checkpoint['init_dim'],
+            use_bias=checkpoint['use_bias'],
+            is_bidirectional=checkpoint['is_bidirectional'])
+        net.load_state_dict(checkpoint['model_state'])
+    elif 'performance_rnn' in name:
+        from performance_rnn import PerformanceRNN
+        net = PerformanceRNN(
+            in_dim=checkpoint['in_dim'],
+            hidden_dim=checkpoint['hidden_dim'],
+            batch_size=1,  # we create one song at a time
+            out_dim=checkpoint['out_dim'],
+            dropout=checkpoint['dropout'],
+            num_layers=checkpoint['num_layers'],
+            init_dim=checkpoint['init_dim'],
+            use_bias=checkpoint['use_bias'],
+            is_bidirectional=checkpoint['is_bidirectional'])
+        net.load_state_dict(checkpoint['model_state'])
+    elif 'c_rnn_gan' in name:
+        from c_rnn_gan import Generator
+        net = Generator(
+            in_dim=checkpoint['in_dim'],
+            hidden_dim=checkpoint['hidden_dim'],
+            batch_size=1,  # we create one song at a time
+            out_dim=checkpoint['out_dim'],
+            dropout=checkpoint['dropout'],
+            num_layers=checkpoint['num_layers'],
+            init_dim=checkpoint['init_dim'],
+            use_bias=checkpoint['use_bias'])
+        net.load_state_dict(checkpoint['g_model_state'])
+
+    device = torch.device('cuda:0' if torch.cuda.is_available() else 'cpu')
+    net.to(device)
+    init = torch.randn(net.batch_size, net.init_dim).to(device)
+    steps = 500
+    song = net.generate(
+        init,
+        steps,
[…]
